### 2017-09-04 honors-thesis/usenix.tex
Initial commit of MD4 research
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the
% usepackage command, below.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes,listings,amsmath,xcolor,courier,float}
\usepackage{hyperref,supertabular}

\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Motivating a Constraint-Based Approach to Colliding Hash Functions }

\author{
{\rm Alexander Scheel}\\
\href{mailto:alexander.m.scheel@gmail.com}{alexander.m.scheel@gmail.com} \\
Iowa State University
\and
{\rm Eric Rozier}\\
\href{mailto:erozier@iastate.edu}{erozier@iastate.edu} \\
Iowa State University
}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
% \thispagestyle{empty}

\subsection*{Abstract}

TODO

\section{Introduction}
\subsection{History of SAT in Cryptography}

\begin{itemize}
    \item CryptoSAT
    \item cryptominisat
    \item SHA-1
\end{itemize}

\subsection{Motivating Logical Cryptanalysis on Hash Functions}

We wish to provide the motivations driving our work. To begin, while collisions
in cryptographic hash functions have a wide reach beyond the academic community,
most of the discovered collisions are driven by academics. As such, most papers
begin by introducing a single hash function, the author's preferred notation,
the methods used in whole or part used to discover the collision, the constraints
on the differential and the differential path, and usually end with an example of a
found collision. However, while community's goal is to improve techniques by
better differentials as measured by higher likelihoods of occurrence, little work
is done to determine the utility of the collision. We seek to push the
community in a new direction, towards measuring, and later improving, the utility
of collisions.

Thus we seek to answer the following questions: What does the input space of
a collision look like? Are blocks only of high entropy, likely reflecting
complex dependencies on bits of the input? Or, can fixed strings be required at
positions of an input? Can these blocks be constrained to only inputs of ASCII
or other character sets? Is the collision fixed to a single input differential,
or can many different input differentials be used? Are there constraints on
possible starting states?

By answering these questions, we begin to develop a sense of utility of
collisions. We can see whether they work well at the beginning or end of
longer messages; we can see how much flexibility we have to fit existing
data (towards second-preimage attacks matching existing collisions); and, we
can see how much structure we can enforce for members of a collision. However,
what we cannot answer are the relationships between these facets of utility,
and whether there are more facets we are overlooking.

Ultimately, we seek to develop a framework for pushing collisions from an
academic exercise to a practical exercise in a constrained environment. The
use is two fold: to develop better hash functions with highly-fractured collision
spaces, and to allow users of hash functions to understand the weaknesses
better. Our belief is that, while there should be an abundance of possible
collisions, hash functions can be designed so that possible collisions do not
pose security risks in all use cases. Further, we seek to demonstrate the
usefulness of discussing collisions in a common language of SAT. This will
allow properties of collisions to be easily verified, and find collision
examples targeting a specific use cases.

We leave it as an open challenge to find a set of collisions which lie outside
of our proposed methodology but which can be demonstrated to have high utility.

\subsection{Terminology of Utility}

We seek to extend the literature about the utility of collisions and thus
propose the following new terminology.

The \textbf{constraints} of a collision are the necessary and/or sufficient
boolean formulas for generating a collision, typically as a function of the
state blocks.

A \textbf{state transition constraint} is a constraint on the intermediate
state variables, i.e., round variables. A constraint is \textbf{weak} if
it is a any difference, and a constraint is \textbf{strong} if it is a
signed difference. A equality constraint can either be weak or strong. This
is roughly equivalent to a differential path.

\textbf{Input block constraints}, \textbf{input state constraints}, and
\textbf{output state constraints} are all analogously defined as state
transition constraints.

A \textbf{property} of a collision is any constraint placed upon the
input block which exists independent from the constraints of a collision.

An attack is \textbf{constraint-based} if there exists a SAT encoding of the
attack which can generate multiple collisions.

A collision is \textbf{extensible} if one example of a collision will lead to
other examples under the same differential and weak state transition constraints.
Note that an extensible collision is inherently constraint-based.

A collision is \textbf{differential-independent} if there are multiple
differentials which satisfy the intermediate state constr
A collision is \textbf{malleable} if collisions can be found with different
properties such as ASCII or JSON input blocks, or with fixed internal text.

A collision is \textbf{state-independent} if there exist alternative
starting states which produce collisions. A state-independent collision
is termed strong if one collision can collide under two or more starting
states.

Thus, a collision is said to have \textbf{high utility} if it is
differential-independent, malleable, and state-independent.

\subsection{Common Notation for Constraint-Based Collisions}
A SAT model, $S = (E, A)$, is an ordered tuple consisting of a set, $E$, of
boolean equations of the form $var := expression$, and a set, $A$, of expressions
which must be asserted to true. It should be obvious that this is equivalent
to $3-CNF-SAT$. This notation is used as it is most similar to the ``circuits''
language \cite{circuits}; further, unline $3-CNF-SAT$, this notation is composable
in the sense that adding new constraints does not impact other existing constraints.

We now introduce a series of notations to ease construction of cryptographic
SAT models. $b := AND_{i = 0}^{32}(a)$. Needs to be

Discuss encodings of hash functions? No, leave to section in

Use notation in other sections. Use Overview subsection to discuss
the particular quirks of encoding of hash function into SAT.

\section{MD4}
\subsection{Brief History and Overview}

The MD4 hashing algorithm was developed by Ron Rivest and first released as RFC 1186
in 1990 \cite{rfc1186}. This was later superseded by RFC 1320 in 1992 because the
reference implementation was ``more portable'' \cite{rfc1320}. As early as 2004,
Wang et al, demonstrated the first full collision in MD4 and MD5 \cite{cryptoeprint:2004:199}.
However, it wasn't until 2005 that details of the differential and constraints
were published \cite{Wang2005}. It is believed by some that these differentials
and constraints were found "by hand" \cite{cryptoeprint:2007:206}. Others thus
strove to automate these attacks with algorithmic approaches \cite{Schlaffer2006}
\cite{cryptoeprint:2007:206}, and \cite{Sasaki2007}.

In this section, we will analyze Wang's 2005 and Sasaki's 2007 differentials to MD4.

\subsection{SAT Model of MD4}
Base model.

\subsection{Wang's MD4 Collisions}
\subsubsection{Overview and Motivation}
While many have studied Wang's attack from the point of view of improving the
probability of collision or validating constraints, we seek to extend Wang's work
by showing properties about colliding blocks. In particular, we wish to show that
Wang's attack is constraint-based, extensible, highly malleable,
differential-independent, and state-independent. Thus, we venture that Wang's
attack is a strong attack against MD4, and that MD4 has no remaining collision
resistance.

\subsubsection{Extensible}
We note that Wang's attack is extensible. With a loose fitting on the original
example as referenced by `md4-wang-extensible', we generated an additional
40,000 new collisions. We note that, while not fully explored, the loose
fitting on the original hash produces valid collisions satisfying the closer
reconstruction of the original constraints; that is, no further differential
paths were discovered. By analyzing the new collisions for fixed properties,
an alternative construction of Wang's constraints can be created. Further,
by then negating individual constraints and testing for UNSAT, we can validate
that they are indeed necessary. See table \ref{table:1} for the set of differential
path constraints and table \ref{table:2} for the set of fixed-value constraints.
Together these form an equivalent constraint set as what is seen in \cite{Wang2005}.

\subsubsection{Constraint-Based}
\begin{itemize}
    \item Simple encoding
    \item Advanced constraints
    \item Time difference
\end{itemize}
testing
\subsubsection{Differential-Independent}
\subsubsection{State-Independent}
\subsubsection{Malleable}
\begin{itemize}
    \item ASCII
    \item JSON
    \item substring
\end{itemize}
Testing
\subsubsection{Additional Notes}

\subsection{Sasaki's MD4 Collisions}
\subsubsection{Overview and Motivation}
We study Sasaki's collisions as a means of analyzing a more complete attack,
and as a point of comparison to Wang's \cite{Sasaki2007}. In particular, Sasaki boasts
of substantial speed improvement over Wang's and an increased probability of
collision. The latter should thus translate to improved results with respect to
malleability.

\subsubsection{Extensible}
We note that Sasaki's MD4 Collisions is an extensible collision. With a loose
fitting on the original example as referenced by `md4-sasaki-extensible',
we generated an additional 40,000 new collisions. Further, the constraints
published on page 16, while sufficient, are not necessary; many of the negated
forms have valid collisions as well \cite{Sasaki2007}.

\subsubsection{Differential-Independent}
We note that Sasaki's MD4 collisions are differential independent, finding
examples of all 64 theoretical differentials.

\section{MD5}
\subsection{Brief History and Overview}
\subsection{Wang's MD5 Collisions}
\subsection{Stevens's MD5 Collisions}
\subsection{Xie's MD5 Collisions}

\section{SHA1}
\subsection{Brief History and Overview}
\subsection{Wang's SHA1 Partial Collision}
\subsection{Stevens's SHA1 Full Collision}

\newpage
\section{Collisions}
\subsection{Notation}
We propose the convention of using both a hexadecimal and binary representation.
Hex in a "standard order", i.e.,
\begin{lstlisting}
echo ``<hex>'' | xxd -r -p | openssl md4
\end{lstlisting}

\subsection{MD4}
\subsubsection{Wang's Collisions}

\newpage
\section{Tables}
\subsection{Notation}
Tabular notation notes go here.
Meaning of +, -, T, F, and *.
Note that missing numbers are replaced
with ... when all the same.

\subsection{MD4}
\subsubsection{Wang's Collision}

\begin{table}[H]
\centering
\begin{tiny}
\texttt{
\begin{tabular}{|c c|}
\hline
h1i & constraints \\
\hline
0 & ................................ \\
1 & .........................+...... \\
2 & .....................+..-....... \\
3 & ......+......................... \\
4 & ................................ \\
5 & ..................+............. \\
6 & ..........+-++.................. \\
7 & .................+--............ \\
8 & ...............+................ \\
9 & ......-..+--+................... \\
10 & ..-............................. \\
11 & +............................... \\
12 & ......+..+...................... \\
13 & ..+-.-.......................... \\
14 & ................................ \\
15 & .............+.................. \\
16 & -..-.+-......................... \\
17 & ................................ \\
18 & ................................ \\
19 & +.-............................. \\
20 & -.+-............................ \\
21 & ................................ \\
... & ................................ \\
47 & ................................ \\
\hline
\end{tabular}
}
\end{tiny}
\caption{Discovered differential path constraints in MD4 under Wang's Attack}
\label{table:1}
\end{table}

\begin{table}[H]
\centering
\begin{tiny}
\texttt{
\begin{tabular}{|c c|}
\hline
h1i & constraints \\
\hline
0 & .........................F...... \\
1 & .........................F...... \\
2 & .....................F..TT...... \\
3 & ......F..............F..FT...... \\
4 & ......F..............T..T....... \\
5 & ......T...........F............. \\
6 & ..........FTFF....F............. \\
7 & ..........FFFF...FTT............ \\
8 & ..........TFFF.F.TTT............ \\
9 & ......T..FTTF..F.TTT............ \\
10 & ..T...F..FFFF..T................ \\
11 & F.F...T..FTTF................... \\
12 & F.T...F..F...................... \\
13 & T.FT.TF..F...................... \\
14 & ..FF.FT..T...................... \\
15 & ..FT.TT......F.................. \\
16 & T..T.FT......................... \\
17 & ...T.TT......................... \\
18 & ...T.TT......................... \\
19 & F.TT............................ \\
20 & T.FT............................ \\
21 & ...T............................ \\
22 & ...T............................ \\
23 & ................................ \\
... & ................................ \\
34 & ................................ \\
35 & T............................... \\
36 & T............................... \\
37 & ................................ \\
... & ................................ \\
47 & ................................ \\
\hline
\end{tabular}
}
\end{tiny}
\caption{Discovered fixed-value constraints in MD4 under Wang's Attack}
\label{table:2}
\end{table}

\subsubsection{Sasaki's Collision}

\begin{table}[H]
\centering
\begin{tiny}
\texttt{
\begin{tabular}{|c c|}
\hline
h1i & constraints \\
\hline
0 & -.............................+- \\
0 & +.............................-+ \\
1 & ......................-.+....... \\
1 & ......................+.-....... \\
2 & ....................-........... \\
2 & ....................+........... \\
3 & ..........-++................... \\
3 & ..........+--................... \\
4 & .......-+-............-++++++... \\
4 & .......-++............-++++++... \\
4 & .......+--............+------... \\
4 & .......+-+............+------... \\
5 & -..........-++.................. \\
5 & -..........+--.................. \\
5 & +..........-++.................. \\
5 & +..........+--.................. \\
6 & -......-++............-++++++++. \\
6 & -......-++............+--------. \\
6 & -......+--............-++++++++. \\
6 & -......+--............+--------. \\
6 & +......-++............-++++++++. \\
6 & +......-++............+--------. \\
6 & +......+--............-++++++++. \\
6 & +......+--............+--------. \\
7 & ................................ \\
8 & ......-.-+...................... \\
8 & ......-.+-...................... \\
8 & ......+.-+...................... \\
8 & ......+.+-...................... \\
9 & ---.+--............-............ \\
9 & -++.-++............+............ \\
9 & -++.+--............+............ \\
9 & +--.-++............-............ \\
9 & +--.+--............-............ \\
9 & +++.-++............+............ \\
10 & ...................-............ \\
10 & ...................+............ \\
11 & ...................-............ \\
11 & ...................+............ \\
12 & ..-+..-......................... \\
12 & ..-+..+......................... \\
12 & ..+-..-......................... \\
12 & ..+-..+......................... \\
13 & ................................ \\
14 & ................................ \\
15 & -............................... \\
15 & +............................... \\
16 & ...-............................ \\
16 & ...+............................ \\
17 & ................................ \\
18 & ................................ \\
19 & ................................ \\
20 & -............................... \\
20 & +............................... \\
21 & ................................ \\
... & ................................ \\
31 & ................................ \\
32 & -............................... \\
32 & +............................... \\
33 & ................................ \\
... & ................................ \\
47 & ................................ \\
\hline
\end{tabular}
}
\end{tiny}
\caption{Discovered differential paths' constraints in MD4 under Sasaki's Attack.
Includes duplicate paths.}
\label{table:3}
\end{table}

\begin{table}[H]
\centering
\begin{tiny}
\texttt{
\begin{tabular}{|c c|}
\hline
h1i & constraints \\
\hline
0 & ......................T.T....... \\
1 & T..............................T \\
2 & T.....................F.F......F \\
3 & ....................F.T.T....... \\
4 & ..........FFF.......F........... \\
5 & .........T............FFFFFFF... \\
6 & ...........FTF.................. \\
7 & T......F.F.TFT.....F..FFFFFFTFF. \\
8 & F..................F..TTTTFTTTT. \\
9 & ........TT...................... \\
10 & FFF..TF.TT...................... \\
11 & FTT..FT......................... \\
12 & ...................F............ \\
13 & ..FF..F............F............ \\
14 & ..TT..T......................... \\
15 & ...T............................ \\
16 & ................................ \\
17 & ...T............................ \\
18 & ...T............................ \\
19 & ................................ \\
... & ................................ \\
47 & ................................ \\
\hline
\end{tabular}
}
\end{tiny}
\caption{Discovered fixed-value constraints in MD4 under Sasaki's Attack}
\label{table:4}
\end{table}

{\footnotesize \bibliographystyle{acm}

\bibliography{thesis}}

% \theendnotes

\end{document}

### 2017-09-04 honors-thesis/build.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+pdflatex ./usenix.tex ; bibtex ./usenix  ; pdflatex ./usenix.tex ; pdflatex ./usenix.tex

### 2017-09-23 md4-wangs-neighborhood/usenix.tex

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the
% usepackage command, below.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes,listings,amsmath,xcolor,courier,float}
\usepackage{hyperref,supertabular,amssymb}

\providecommand{\lxor}{\oplus}

\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf The Neighborhood of a MD4 Collision }

\author{
{\rm Alexander Scheel}\\
\href{mailto:alexander.m.scheel@gmail.com}{alexander.m.scheel@gmail.com} \\
Iowa State University
\and
{\rm Eric Rozier}\\
\href{mailto:erozier@iastate.edu}{erozier@iastate.edu} \\
Iowa State University
}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
% \thispagestyle{empty}

\subsection*{Abstract}

In 2005, Wang et al. published the details of a collision attack against the
MD4 hash function. We extend their work by showing that not
only is the attack state independent and differential independent, but also
that the neighborhood of the attack is populated with other classes of
collisions. We describe an efficient encoding of this problem into SAT,
and present example classes in the neighborhood of Wang's.

\section{Introduction}
The MD4 algorithm was developed by Ron Rivest around 1990 and is standardized
in RFC 1320 \cite{rfc1320}. While largely known to be insecure, and thus
having fallen out of use, its simple construction leads to fast implementations
and easy encoding as a SAT problem. An attack against a chosen-state MD4 was
published by Dobbertin in 1998 \cite{Dobbertin1998}. In 2004, Wang et al.
demonstrated a collision attack against the full MD4
\cite{cryptoeprint:2004:199}, before publishing the details in 2005
\cite{Wang2005}. With reduced time complexity compared to Dobbertin, this
class of collision went on to inspire other attacks such as Sasaki et al.,
which pushed the complexity of collisions to less than two MD4 evaluations
\cite{Sasaki2007}.

Outside of work published in 2006 by Mironov and Zhang \cite{Mironov2006},
the authors have found little evidence of the application of SAT solvers
to hash function collisions and MD4 specifically. The Mironov and Zhang work
was focused on the feasibility of such an encoding and the time for finding
new examples of collisions in the Wang et al. class.

We present the following four improvements in collisions.
\begin{enumerate}
    \itemsep0em
    \item Wang et al.'s attack is differential independent.
    \item Wang et al.'s attack is state independent.
    \item Wang et al.'s attack is highly malleable.
    \item The neighborhood of Wang et al.'s attack contains other collision classes.
\end{enumerate}

\section{Terminology}
We first begin by giving a brief set of terminology:

A \textbf{differential} is the signed XOR difference between two input blocks
to a hash function.

The \textbf{differential path} is the set of state transition constraints
which produce a full or partial collision.

The \textbf{constraints} of a collision are the necessary and/or sufficient
boolean formulas for generating a collision, typically as a function of the
state blocks.

A \textbf{state transition constraint} is a constraint on the intermediate
state variables, i.e., round variables. A constraint is \textbf{weak} if
it is a any difference, and a constraint is \textbf{strong} if it is a
signed difference. A equality constraint can either be weak or strong. This
is roughly equivalent to a differential path.

\textbf{Input block constraints}, \textbf{input state constraints}, and
\textbf{output state constraints} are all analogously defined as state
transition constraints.

A \textbf{property} of a collision is any constraint placed upon the
input block which exists independent from the constraints of a collision.

An attack is \textbf{constraint-based} if there exists a SAT encoding of the
attack which can generate multiple collisions.

A collision is \textbf{differential-independent} if there are multiple
input block differentials which satisfy the state transition constraints.

A collision is \textbf{malleable} if collisions can be found with different
properties such as ASCII or JSON input blocks, or with fixed internal text.

A collision is \textbf{state-independent} if there exist alternative
starting states which produce collisions. A state-independent collision
is termed \textbf{strong} if one block can collide under two or more starting
states.

A \textbf{collision class} is defined by the differential path and optionally
differential. The \textbf{neighborhood of a collision class} is a set of
alternative differential paths with low distance between differences.
For example, consider the differential path with $\Delta(I_{0 ... 48}) = 0$;
then the differential path with $\Delta(I_0) = 1$ has low distance to
$\Delta(I_0) = 0$, and hence is said to be in the neighborhood of
$\Delta(I_{0 ... 48})$.

\section{Notation}

As we are not concerned with the theoretical underpinnings of the SAT models,
nor will we use the model to provide general proofs, we use the following
notations. This notation is heavily influenced by our use of the $bc2cnf$
utility by Junttila \cite{circuits}.

A constraint, $c = (n, e)$ is a mapping between a name, $n$, and an equation,
$e$, of boolean clauses. These equations can be simple ($a \land b$) or complex
($a \lxor (c \land d) \lxor (a \lor b)$). A model, $S = (C, A)$, is an ordered
pair of a set, $C$, of constraints and a set, $A$, a set of names which must
evaluate to true. Note that input variables are implicit and thus left
unspecified.

We construct the base model for MD4 as follows. Let $s_{0} .. s_{127}$ be the
input state variables. Let $b_{0} .. b_{511}$ be the input block. We define
$i_0 .. i_{1535}$ as the bits of the intermediate state variables generated
via MD4, per RFC 1320 \cite{rfc1320}, and $o_{0} .. o_{128}$ as the resulting
output state. With the $s$'s and $b$'s left unspecified, the model for MD4
is thus $C_{MD4} = \{(i_0, e(i_0)), ..., (o_{128}, e(o_{128})) \}$. As we
are interested in collisions, we use $C_{MD4, h_1}$ to denote the construction
of MD4 with variables $i$, $o$, $s$, and $b$ prefixed with $h_1$.

We then define $I_{md4}$ to be the set of initial state values, with
$I_{md4, h_1}$ to be those state values with names prefixed by $h_1$. We also
define $C_{collision, h_1, h_2}$ to be the two-way collision between
$h_1$ and $h_2$; that is, all output bits are equal.

\section{Differential Independent}

We note that Wang et al.'s attack on MD4 is differential independent. Let
$C_{wangs, h_1, h_2}$ be the set of the state transition constraints on page 15
and sufficient constraints on page 16 \cite{Wang2005}. Let
$\bar{\Delta}_{wangs}$ be the negation of the specified differential on page 7
\cite{Wang2005}. Then we claim that
\begin{align*}
S = ( \{ C_{MD4, h_1} \cup C_{MD4, h_2} \cup C_{wangs, h_1, h_2} \cup
\bar{\Delta}_{wangs} \cup C_{collision, h_1, h_2} \cup I_{md4, h_1} \cup
I_{md4, h_2} \}, \{ C_{wangs}, \bar{\Delta}_{wangs},
C_{collision h_1, h_2} \})
\end{align*}
 is a model which encodes a collision between $h_1$
and $h_2$ satisfying the conditions of Wang's attack but with a diffferent
differential.

Claim: the model $S$ above is satisfiable. We give the following examples
as proof.

By negating each found differential and re-running we can find all possible
differentials. This is a total of 238 differentials. We note the signed
block differences below.

\section{State Independent}

We note that Wang et al.'s is strongly state independent. We construct a model
$S$ with four instances of MD4, $h_1$, $h_2$, $h_3$, $h_4$, which share an input
block and have a non-zero input state delta.

We thus claim that
\begin{align*}
S = ( \{ C_{MD4, h_1}, C_{MD4, h_2}, C_{MD4, h_3}, C_{MD4, h_4},
C_{wangs, h_1, h_2}, C_{wangs, h_3, h_4}, \Delta_{wangs, h_1, h_2},
\Delta_{wangs, h_3, h_4}, C_{collision, h_1, h_2}, C_{collision, h_3, h_4},
\Delta_{any, h_1s, h_2s} \}, \{
C_{wangs, h_1, h_2}, C_{wangs, h_3, h_4}, \Delta_{wangs, h_1, h_2},
\Delta_{wangs, h_3, h_4}, C_{collision, h_1, h_2}, C_{collision, h_3, h_4},
\Delta_{any, h_1s, h_2s} \} )
\end{align*}
encodes the above properties and is satisifable. We reproduce several
examples below.

\section{Highly Malleable}
In the above style above, we note that Wang et al.'s is malleable, and highly
malleable under alternative differentials. We produce the following examples;
though we note that it is not possible with SAT prove that the attack is
fully malleable and can produce collisions with any property.

\section{Neighborhood}

{\footnotesize \bibliographystyle{acm}

\bibliography{thesis}}

% \theendnotes

\end{document}

### 2017-11-27 dsn-2018/bare_conf.tex
Initial draft of dsn-2018 paper
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
\usepackage{hyperref,supertabular,amssymb}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Measuring the Utility of a Collision:\\ The Neighborhood of MD4 Collisions}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Alexander M. Scheel}
\IEEEauthorblockA{Iowa State University\\
Ames, IA, USA \\
Email: alexander.m.scheel@gmail.com}
\and
\IEEEauthorblockN{Eric W. D. Rozier}
\IEEEauthorblockA{Iowa State University\\
Ames, IA, USA \\
Email: erozier@iastate.edu}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
We propose a series of metrics for evaluating the utility of collision classes
and develop a set of techniques for generating new collision classes in hash
functions utilizing a logical cryptanalaysis framework. We demonstrate the
effectiveness of this framework on the hash function MD4 and show that the
existing attacks are of high utility, but do not allow for second preimage
attacks under an expansive neighborhood. We partially analyze over 35,000 new
collision classes for their utility under these methods.
\end{abstract}

% no keywords
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

Cryptographic hash functions form the core of many protocols. From file
integrity checks to verify long term storage of data, to cache invalidation
techniques, and use as a building block in network protocols such as Kerberos
and TLS, this class of functions necessarily has some strong guarantees about
the properties of these functions. There are three basic properties hash
functions must have to be considered cryptographically secure:
\begin{itemize}
    \item Preimage Resistance: It should be computationally hard to find
        the inverse of a hash function.
    \item Second Preimage Resistance: Given an input block, it should be
        computationally hard to find a second block which hashes to the
        same value as the first block.
    \item Collision Resistance: It should be computationally hard to find two
        blocks which hash to the same value.
\end{itemize}
Note that a second preimage is necessarily a collision and is also a stronger
result.

From an attacker's perspective, finding a preimage or second preimage has been
difficult. To the author's knowledge, the current bound for a preimage attack
in MD5 is $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. Further,
while J. Kelsey and B. Schneier proposed a method for finding second preimages
in less than $2^n$, we note that this requires rather rather long messages
\cite{SchneierSecondPreimage}. However, collision attacks are well within reach
for adversaries, and well past only theoretical attacks. The work of
H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
M. Schlaffer \cite{Schlaffer2006} and others demonstrate the ease with which
collisions can be found.

From a cryptographic perspective, the existence of a feasible collision attack
breaks the requisite properties, thus deprecating the function's use. In some
instances however, there continues to be widespread use of deprecated hash
functions. One widespread example is the continued use by Git of SHA-1, despite
M. Stevens et al. having published the first known full SHA-1 collision in
February of 2017 \cite{cryptoeprint:2017:190}. However, this attack had a bound
of $2^{63.1}$ -- faster than the theoretical $2^{80}$ but still requiring
significant investment in compute resources.

We define a series of criteria for analyzing the utility of a collision class.
Collision classes of high utility are closer to a second preimage attack in the
amount of flexibility they provide to an attacker, whereas classes of low
utility may not affect all systems which use a particular hash function.
Furthermore, we demonstrate new techniques for working with collisions under
a logical cryptanalysis framework. Throughout this paper, we make use of the
MD4 hash function due to the ease of generating new classes of collisions.

\section{Terminology \& Notation}

\textbf{TODO}.

\section{The Utility of a Collision}

We propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit-step neighbors a collision class has.
    \item The maximum count of zeros in a the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness for having the property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

The first metric evaluates the flexibility of the differential path. A
differential path with more flexibility will have more differentials which
produce blocks with the given differential path. More differentials implies
a greater flexibility in choice of colliding block, and possibly allowing
for multiple collisions for a given colliding block. Furthermore, a collision
with more differentials is more likely to satisfy the last metric, having a
pair of blocks--both ASCII--which produce a collision.

The second metric evaluates the density of the neighborhood of a collision
class. If a collision resides in a dense neighborhood, it provides more
possible collision classes to search for a second preimage, chosen prefix, or
other structure desired in a collision. If however, a collision class has no
neighbors, then it cannot be used to find other possible classes for other
input blocks.

The maximum quantity of zeros (or ones) in the binary representation of a
block serve as a measure of the extremes to which a collision can be pushed.
This can additionally be extended to any suitable bit pattern in any base to
provide a more relevant metric as desired by the system under study.

The fourth metric evaluates the utility of a collision when the internal state
of the hash function is unknown. If a collision occurs under multiple initial
values, this could be used to attack some systems where user provided input is
appended to unknown data and then hashed. If a block collides under a suitably
large number of initial values, the attack becomes highly likely to occur
successfully. However, measuring the exact number of initial values a block
collides under is beyond the scope of this work.

The last metric is similar to the third in that it looks for specific bit
patterns in a collision. ASCII is one example of a widely used constraint
system. Further examples, such as JSON, XML, etc., may likewise be
supplemented based on the specifics of the system.

If a collision class satisfies many of these properties, then it is more
flexible and thus more likely to be used to target deployed systems. If,
however, a collision class does not satisfy these properties, it impact is
likely severely limited in scope, and may not provide useful information to
find other collision classes which have higher utility.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank the Department of Electrical and Computer
Engineering.

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}

% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{thesis}

% that's all folks
\end{document}

### 2017-11-27 dsn-2018/bare_conf.tex
Outline remaining portion of paper
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 %
-%\usepackage{amsmath}
+\usepackage{amsmath}
 % A popular package from the American Mathematical Society that provides
@@
 %
-%\usepackage{algorithmic}
+\usepackage{algorithmic}
 % algorithmic.sty was written by Peter Williams and Rogerio Brito.
@@
 
+\section{Techniques for Logical Cryptanalysis}
+
+The following techniques have been extensively tested on MD4 and partially
+tested on MD5 and believe to apply fully to MD5. They may or may not apply
+to any later hash function, such as SHA-1, SHA-2, or SHA-3, and have not been
+tested yet.
+
+In the following sections, we use the collisions of X. Wang \cite{Wang2005},
+Y. Sasaki \cite{Sasaki2007}, M. Schlaffer \cite{Schlaffer2006}, H. Dobbertin
+\cite{Dobbertin1998}, and P. Kasselman \cite{KasselmanMD4} for examples.
+
+\subsection{Distance Metric}
+
+We introduce a distance function between collision classes by the number of
+differences in intermediate rounds deltas. We find justification for this
+metric in the existing literature on MD4: H. Dobbertin's \cite{Dobbertin1998}
+and P. Kasselman's \cite{KasselmanMD4} classes have distance $1$ under this
+metric.
+
+That is:
+\begin{align}
+    \delta(C_1, C_2) = \| \{ i : I(C_1, i) \neq I(C_2, i) \} \|
+\end{align}
+
+For the example collisions, below are distances under this metric:
+\textbf{TODO}.
+
+\subsection{Neighborhood}
+
+We define the neighborhood of a collision class to be all other collisions
+which have a distance of one.
+
+That is:
+\begin{align}
+    N(C) = \{ C_j : \delta(C, C_j) = 1 \}
+\end{align}
+
+For the example collisions, below are sizes of the neighborhoods:
+\textbf{TODO}.
+
+\subsection{Reduced Rounds}
+\subsubsection{Abstract}
+
+\textbf{TODO} - Describe technique applied to tags
+
+Give some reasoning for why it works, example graphs.
+
+\subsubsection{Neighborhood}
+
+\textbf{TODO} - Describe technique applied neighborhood extensions
+
+Give some reasoning for why it works, example graphs.
+
+\section{Empirical Results}
+\subsection{Unique Differentials}
+
+\textbf{TODO}
+
+Include graphic of count of unique differentials for all 35k collision classes
+
+Separate table for known attacks as data points. Examples in appendix?
+
+\subsection{Unit-Step Neighborhood}
+
+\textbf{TODO}
+
+Include graphic of count of number of neighbors for all 35k collision classes
+
+Separate table for known attacks as data points.
+
+\subsection{Zeroes \& Ones}
+
+\textbf{TODO}
+
+Include graphic of count of max number of zeros/ones for all 35k collision
+classes.
+
+Separate table for known attacks as data points. Examples as appendix.
+
+\subsection{Multicollisions}
+
+\textbf{TODO}
+
+Include table detailing how many have multicollisions and how many don't.
+Hypothesis: all do since all existing datapoints do.
+
+Separate table for known attacks as data points. Examples as appendix.
+
+\subsection{ASCII Blocks}
+
+\textbf{TODO}
+
+Include table detailing how many have ASCII blocks and how many don't.
+
+Separate table for known attacks as data points. Examples as appendix.
+
+\section{Open Research}
+
+The framework can be seen here \textbf{TODO}. Complete data set available
+upon request.
+
 % An example of a floating figure using the graphicx package.
@@
 \section{Conclusion}
+
 The conclusion goes here.
@@
 The authors would like to thank the Department of Electrical and Computer
-Engineering.
+Engineering and Nathan Mulbrook for providing hardware.
 

### 2017-11-27 dsn-2018/bare_conf.tex
Updated abstract and title
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 % Do not put math or special symbols in the title.
-\title{Measuring the Utility of a Collision:\\ The Neighborhood of MD4 Collisions}
+\title{Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4}
 
@@
 \begin{abstract}
-We propose a series of metrics for evaluating the utility of collision classes
-and develop a set of techniques for generating new collision classes in hash
-functions utilizing a logical cryptanalaysis framework. We demonstrate the
-effectiveness of this framework on the hash function MD4 and show that the
-existing attacks are of high utility, but do not allow for second preimage
-attacks under an expansive neighborhood. We partially analyze over 35,000 new
-collision classes for their utility under these methods.
+The discovery of fast collision attacks in cryptographic hash functions has
+traditionally resulted in the immediate deprecation of that hash function.
+In this paper we propose five scalable and practical metrics for evaluating
+the utility of collision classes based on boolean constraints and show that
+the published attacks by X. Wang, Y. Sasaki, P. Kasselman, H. Dobbertin, and
+M. Schlaffer in MD4 have high utility. We expand on existing attacks by
+developing a series of techniques based on logical cryptanalysis to find
+over 35,000 collisions in MD4 based on existing collisions, through the novel
+definition of a collision neighborhood. We demonstrate new techniques for
+inductively building full collisions from reduced round variants of MD4. We
+propose these techniques as a mechanism for measuring hash trustworthiness and
+discuss potential applications to real-world systems.
 \end{abstract}
@@
 \textbf{TODO}.
+
+$\mathcal{C}$ for collision class. Ordered tuple.
+$\mathcal{F}$ for family of collision classes. Set.
+$I$ is an indexing function.
 
@@
 
-\subsection{Distance Metric}
+\subsection{Distance Metrics}
 
-We introduce a distance function between collision classes by the number of
-differences in intermediate rounds deltas. We find justification for this
-metric in the existing literature on MD4: H. Dobbertin's \cite{Dobbertin1998}
-and P. Kasselman's \cite{KasselmanMD4} classes have distance $1$ under this
-metric.
+We introduce a distance function, $\delta$ between collision classes by the
+number of differences in intermediate rounds deltas. That is, given two
+collision classes, $C_1, C_2 \in \mathcal{C}$:
+\begin{align}
+    \delta(C_1, C_2) = \big| \{ i : I(C_1, i) \neq I(C_2, i) \} \big|
+\end{align}
+We find justification for this metric in the existing literature on MD4:
+H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
+collision classes have distance $1$ under this metric.
 
+Please refer to Table \ref{table:distance} for the distances between existing
+collisions in MD4.
+
+\begin{table}
+    \caption{Distance between Existing Collision Classes}
+    \label{table:distance}
+    \begin{tabular}{c c c c c c}
+        & X. W. & Y. S. & P. K. & H. D. & M. S. \\
+        X. Wang's & 0 & 21 & 26 & 27 & 12 \\
+        Y. Sasaki's & & 0 & 25 & 26 & 2 \\
+        P. Kasselman's &  &  & 0 & 1 & 25 \\
+        H. Dobbertin's &  &  &  & 0 & 26 \\
+        M. Schlaffer's &  &  &  &  & 0 \\
+    \end{tabular}
+\end{table}
+
+We can define a similar distance function, $\Delta$, between families of
+collision classes by cardinality of the symmetric difference in the two
+collision families. That is, given two families of collision classes,
+$F_1, F_2 \in \mathcal{F}$:
+\begin{align}
+    \Delta(F_1, F_2) = \big| (F_1 \cup F_2) \setminus (F_1 \cap F_2) \big|
+\end{align}
+This is convenient for when the specifics of the differential path do not
+matter, merely that there exist at least one collision class of the specified
+form.
+
+\subsection{Neighborhoods}
+
+We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
+set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
 That is:
 \begin{align}
-    \delta(C_1, C_2) = \| \{ i : I(C_1, i) \neq I(C_2, i) \} \|
+    N(C, d) & = \{ C_j : \delta(C, C_j) = d \} \\
+    N(C) & = N(C, 1)
 \end{align}
+For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
+may optionally be ommited, in which case the unit distance is implied. Thus,
+$C_{Kasselman} \in N(C_{Dobbertin})$.
 
-For the example collisions, below are distances under this metric:
-\textbf{TODO}.
+The implications of neighborhoods are discussed further in section
+\ref{empirical:neighborhood}.
 
-\subsection{Neighborhood}
-
-We define the neighborhood of a collision class to be all other collisions
-which have a distance of one.
-
-That is:
-\begin{align}
-    N(C) = \{ C_j : \delta(C, C_j) = 1 \}
-\end{align}
-
-For the example collisions, below are sizes of the neighborhoods:
-\textbf{TODO}.
+We can similarly define the neighborhood of a family of collision classes,
+$F \in \mathcal{F}$, to be the set of all other families of collision classes
+at a fixed distance, $d \in \mathbb{N}$, from $F$.
 
@@
 
-\textbf{TODO} - Describe technique applied to tags
-
-Give some reasoning for why it works, example graphs.
+\textbf{TODO} - Describe technique applied to tags.
 
@@
 \textbf{TODO}
+Please refer to Table \ref{table:differentials} for the number of differential
+paths...
 
-Include graphic of count of unique differentials for all 35k collision classes
+\begin{table}
+    \caption{Number of Differentials for Existing Collision Classes}
+    \label{table:differentials}
+    \begin{tabular}{c c c c}
+        \textbf{Attack} & \textbf{Size} & \textbf{Attack} & \textbf{Size} \\
+        X. Wang's & 64 & Y. Sasaki's & 4 \\
+        H. Dobbertin's & 32 & P. Kasselman's & 32 \\
+        M. Schlaffer's & 64 & & \\
+    \end{tabular}
+\end{table}
 
-Separate table for known attacks as data points. Examples in appendix?
-
-\subsection{Unit-Step Neighborhood}
+\subsection{Unit-Step Neighborhood} \label{empirical:neighborhood}
 
@@
 
-Include graphic of count of number of neighbors for all 35k collision classes
+Please refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
+of existing collisions. In particular, note that Y. Sasaki's attack--which
+claimed to improve upon X. Wang's attack--has a larger neighborhood size, and
+likewise with P. Kasselman's attack which improved upon H. Dobbertin's attack.
 
-Separate table for known attacks as data points.
+\begin{table}
+    \caption{Neighborhood Sizes for Existing Collision Classes}
+    \label{table:neighborhood}
+    \begin{tabular}{c c c c}
+        \textbf{Attack} & \textbf{Size} & \textbf{Attack} & \textbf{Size} \\
+        X. Wang's & 54 & Y. Sasaki's & 157 \\
+        H. Dobbertin's & 55 & P. Kasselman's & 60 \\
+        M. Schlaffer's & 100 & & \\
+    \end{tabular}
+\end{table}
 
@@
 
-Separate table for known attacks as data points. Examples as appendix.
+See Table \ref{table:ascii} for results. Note that the validation of
+P. Kasselman's and H. Dobbertin's results do not hold for the latter attacks
+by X. Wang, Y. Sasaki, or M. Schlaffer. Thus, while the latter attacks have
+been viewed as being of better quality, under this particular metric,
+P. Kasselman's and H. Dobbertin's collision classes are better.
+
+\begin{table}
+    \caption{ASCII Blocks in Existing Collision Classes}
+    \label{table:ascii}
+    \begin{tabular}{c c c}
+        \textbf{Attack} & \textbf{Single Block} & \textbf{Both Blocks} \\
+        X. Wang's & true & false \\
+        Y. Sasaki's & true & false \\
+        P. Kasselman's & true & \textbf{true} \\
+        H. Dobbertin's & true & \textbf{true} \\
+        M. Schlaffer's & true & false \\
+    \end{tabular}
+\end{table}
 

### 2017-11-28 dsn-2018/bare_conf.tex
Flatten subsubsections, expand techniques
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 \usepackage{cite}
-\usepackage{hyperref,supertabular,amssymb}
+\usepackage[hidelinks]{hyperref}
+\usepackage{supertabular,amssymb}
 % cite.sty was written by Donald Arseneau
@@
 
-$\mathcal{C}$ for collision class. Ordered tuple.
-$\mathcal{F}$ for family of collision classes. Set.
-$I$ is an indexing function.
+$\mathcal{C}$ for collision class. Ordered tuple, indexable.
+$\mathcal{F}$ for family of collision classes. Ordered set, indexable.
+$[i]$ is an indexing function.
+$R$ is number of rounds of $F$ or $C$.
 
-\section{The Utility of a Collision}
+\section{Measuring the Utility of a Collision}
 
@@
 \begin{align}
-    \delta(C_1, C_2) = \big| \{ i : I(C_1, i) \neq I(C_2, i) \} \big|
+    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
 \end{align}
@@
 
-Please refer to Table \ref{table:distance} for the distances between existing
+Refer to Table \ref{table:distance} for the distances between existing
 collisions in MD4.
@@
 \begin{align}
-    N(C, d) & = \{ C_j : \delta(C, C_j) = d \} \\
+    N(C, d) & = \{ C_j \in \mathcal{C} : \delta(C, C_j) = d \} \\
     N(C) & = N(C, 1)
@@
 For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
-may optionally be ommited, in which case the unit distance is implied. Thus,
+may optionally be omitted, in which case the unit distance is implied. Thus,
 $C_{Kasselman} \in N(C_{Dobbertin})$.
 
-The implications of neighborhoods are discussed further in section
-\ref{empirical:neighborhood}.
+The implications of collision class neighborhoods are discussed further in
+section \ref{empirical:neighborhood}.
 
@@
 $F \in \mathcal{F}$, to be the set of all other families of collision classes
-at a fixed distance, $d \in \mathbb{N}$, from $F$.
+at a fixed distance, $d \in \mathbb{N}$, from $F$. That is:
+\begin{align}
+    \mathcal{N}(F, d) & = \{ F_j \in \mathcal{F} : \delta(F, F_j) = d \} \\
+    \mathcal{N}(F) & = N(F, 1)
+\end{align}
+The distance parameter, $d$, may optionally be omitted, in which case the unit
+distance is implied.
 
-\subsection{Reduced Rounds}
-\subsubsection{Abstract}
+Neighborhoods can be classified into three types: \textit{expansion},
+\textit{internal}, and \textit{mixed}. Let $d$ be fixed. An \textit{expansion}
+neighborhood of a collision class, $C \in \mathcal{C}$, is the neighborhood
+restricted only to those collision classes which only differ in rounds external
+to the collision family of $C$. An \textit{internal} neighborhood of $C$ is the
+neighborhood restricted only to those collision classes which differ in rounds
+internal to the collision family of $C$. An \textit{mixed} neighborhood of $C$
+is the neighborhood restricted only to collisions which differ in a round
+internal and a round external to the collision family of $C$. That is:
+\begin{align*}
+    N_{exp}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \in F(C), C_j[i] = C[i] \} \\
+    N_{int}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \not\in F(C), C_j[i] = C[i] \} \\
+    N_{mix}(C, d) = \{ & C_{j} \in N(C, d) : \exists i \in F(C), C_j[i] \neq C[i] \\
+        & \text{ and } \exists k \not\in F(C), C_j[k] \neq C[k] \}
+\end{align*}
 
-\textbf{TODO} - Describe technique applied to tags.
+\subsection{Family Similarity}
+We define a relation among families of collisions across rounds. Let $F_1$ and
+$F_2$ be two different families of collisions. Then we say that $F_1$ and $F_2$
+are \textit{similar}, and notate it $F_1 \lesssim F_2$, if $R(F_1) \leq R(F_2)$
+and $F_1 \subseteq F_2$. Note that, when $R(F_1) = R(F_2)$, this is equivalent
+to saying that $F_2$ is in some expansion neighborhood of $F_1$. Further, when
+$R(F_1) < R(F_2)$ and $F_1 = F_2$, then we say that $F_2$ is the
+\textit{trivial extension} of $F_1$.
 
-\subsubsection{Neighborhood}
+We claim that the following statements are true:
+\begin{enumerate}
+    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
+        $F' \lesssim F$ and $R(F') + 4 = R(F)$.
+    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
+        $F \lesssim F'$.
+\end{enumerate}
+
+\textbf{TODO} - Describe justification? Later sections?
+
+\subsection{Class Similarity}
+
+We claim that the aforementioned statements hold when considering individual
+collision classes instead of families of collision classes, but with less
+likelihood.
 
@@
 
+\subsection{Miscellaneous}
+
+\textbf{TODO} - Describe chosen prefix, second preimage, ASCII, and other
+tips and tricks for faster model runs.
+
 \section{Empirical Results}
@@
 \textbf{TODO}
-Please refer to Table \ref{table:differentials} for the number of differential
+Refer to Table \ref{table:differentials} for the number of differential
 paths...
@@
 
-Please refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
+Refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
 of existing collisions. In particular, note that Y. Sasaki's attack--which
@@
 \end{table}
+
+\section{New Collision}
+
+\section{Broader Impacts}
+
+\textbf{TODO} - Discuss impacts on real-world systems like MD5, SHA-1, etc.
 

### 2017-11-28 dsn-2018/bare_conf.tex
Outline changes

@@
 % affiliations
-\author{\IEEEauthorblockN{Alexander M. Scheel}
-\IEEEauthorblockA{Iowa State University\\
-Ames, IA, USA \\
-Email: alexander.m.scheel@gmail.com}
-\and
-\IEEEauthorblockN{Eric W. D. Rozier}
-\IEEEauthorblockA{Iowa State University\\
-Ames, IA, USA \\
-Email: erozier@iastate.edu}}
-
+\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
+\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
+Ames, Iowa, USA 50011\\
+Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}
 % conference papers do not typically use \thanks and this command
@@
 
+\textbf{The contributions of this paper are...}
+
+\begin{itemize}
+\item We...
+\end{itemize}
+
 We define a series of criteria for analyzing the utility of a collision class.
@@
 
+The remainder of this paper is organized as follows... In Section
+\ref{Sec:Terminology} we...
+
+\section{Related Work}
+\label{Sec:Related}
+
+Broad context... intersection... subsection introductions.
+
+\subsection{Differential Cryptanalysis}
+
+\subsection{Satisfiability Solving}
+
 \section{Terminology \& Notation}
+\label{Sec:Terminology}
 
@@
 
-\section{Broader Impacts}
+\section{Discussion}
 
 \textbf{TODO} - Discuss impacts on real-world systems like MD5, SHA-1, etc.
-
-\section{Open Research}
-
-The framework can be seen here \textbf{TODO}. Complete data set available
-upon request.
 
@@
 
+\section{Future Work}
+
+The framework can be seen here \textbf{TODO}. Complete data set available
+upon request.
+
 % conference papers do not normally have an appendix

### 2017-11-29 dsn-2018/bare_conf.tex
Expand Introduction, Related Work...
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 techniques, and use as a building block in network protocols such as Kerberos
-and TLS, this class of functions necessarily has some strong guarantees about
-the properties of these functions. There are three basic properties hash
+and TLS, this class of functions necessarily has strong guarantees about
+properties of its members. There are three basic properties hash
 functions must have to be considered cryptographically secure:
 \begin{itemize}
-    \item Preimage Resistance: It should be computationally hard to find
+    \item \textbf{Preimage Resistance}: It should be computationally hard to find
         the inverse of a hash function.
-    \item Second Preimage Resistance: Given an input block, it should be
+    \item \textbf{Second Preimage Resistance}: Given an input block, it should be
         computationally hard to find a second block which hashes to the
         same value as the first block.
-    \item Collision Resistance: It should be computationally hard to find two
+    \item \textbf{Collision Resistance}: It should be computationally hard to find two
         blocks which hash to the same value.
@@
 Note that a second preimage is necessarily a collision and is also a stronger
-result.
+result in practice.
 
@@
 functions. One widespread example is the continued use by Git of SHA-1, despite
-M. Stevens et al. having published the first known full SHA-1 collision in
-February of 2017 \cite{cryptoeprint:2017:190}. However, this attack had a bound
-of $2^{63.1}$ -- faster than the theoretical $2^{80}$ but still requiring
-significant investment in compute resources.
+M. Stevens et al. publishing a full SHA-1 collision in February of 2017
+\cite{cryptoeprint:2017:190}. However, this attack had a bound of $2^{63.1}$ --
+faster than the theoretical $2^{80}$ but still requiring significant compute
+resources.
 
-\textbf{The contributions of this paper are...}
-
+We seek to address the problem of evaluating the use cases for a collision,
+to determine--or partially determine--whether or not a collision affects
+a particular system. This paper contributes several novel techniques to
+logical cryptanalysis and uses these techniques to derive metrics of the
+trustworthiness of hash functions. We claim the following results are novel
+in the field:
 \begin{itemize}
-\item We...
+    \item We define the neighborhood of a class of collisions and show that
+        it is frequently non-empty.
+    \item We discuss techniques for finding useful collisions within a
+        class of collisions.
+    \item We propose additional techniques for building new classes of
+        collisions and show that collisions in MD4 can be found inductively.
+    \item We propose five universal metrics for evaluating the utility of
+        a class of collisions.
 \end{itemize}
 
-We define a series of criteria for analyzing the utility of a collision class.
-Collision classes of high utility are closer to a second preimage attack in the
-amount of flexibility they provide to an attacker, whereas classes of low
-utility may not affect all systems which use a particular hash function.
-Furthermore, we demonstrate new techniques for working with collisions under
-a logical cryptanalysis framework. Throughout this paper, we make use of the
-MD4 hash function due to the ease of generating new classes of collisions.
+\textbf{TODO - Transition}
 
-The remainder of this paper is organized as follows... In Section
-\ref{Sec:Terminology} we...
+We propose that collision classes of high utility are closer to a second
+preimage attack in the amount of flexibility they provide to an attacker,
+whereas classes of low utility may not affect all systems which use a
+particular hash function. Furthermore, we demonstrate new techniques for
+working with collisions under a logical cryptanalysis framework. Throughout
+this paper, we make use of the MD4 hash function due to the ease of generating
+new classes of collisions.
 
-\section{Related Work}
-\label{Sec:Related}
+\textbf{TODO - Finish Organization}
 
-Broad context... intersection... subsection introductions.
+The remainder of this paper is organized as follows. In Section
+\ref{Sec:Related} we discuss previous results and show how our research
+expands upon prior work. In Section \ref{Sec:Terminology}, we discuss the
+terminology and notations we use throughout the remainder of the paper.
+In Section \ref{Sec:Intuition}, we give the intuition between measuring
+the utility of a collision and motivate why it is useful. In Section
+\ref{Sec:Logical}, we present new techniques in logical cryptanalysis to
+analyze collisions in hash functions.
+
+\section{Related Work} \label{Sec:Related}
+
+\textbf{Broad context... intersection... subsection introductions.}
+
+This paper draws upon the public cryptanalysis of MD4, the wide availability of
+high quality SAT solvers, and previous work in logical cryptanalysis.
 
@@
 
-\subsection{Satisfiability Solving}
+\textbf{TODO}
 
-\section{Terminology \& Notation}
-\label{Sec:Terminology}
+Differential Cryptanalysis has been the major technique behind the discovery
+of collisions in cryptographic hash functions. Its importance can be seen
+everywhere from X. Wang's attacks on MD4 \cite{Wang2005} to more recent
+cryptanalysis of general purpose hash functions such as Murmur3 by
+J. Aumasson, D. Bernstein, and M. Bo{\ss}let \cite{murmur3DC}. However, while
+differential cryptanalysis plays an important role, techniques for automated
+analysis of hash functions are still an emerging area. In
+\cite{cryptoeprint:2004:199}, X. Wang claimed to have found the MD4 collision
+``with hand calculation'', though in \cite{Wang2005}, she specifies message
+modification techniques which can be implemented in code.
+
+Our work removes the need for finding message modification techniques by
+specifying the differential path as part of the SAT formula and letting the
+solver find a pair of satisfying messages which follow the differential path
+and produce a collision. Further constraints can then be placed upon this
+model, such as a chosen prefix or desired start state. We believe performing
+these attacks by hand to be difficult, and note that message modification
+techniques may affect the chosen prefix.
+
+\subsection{Logical Cryptanalysis}
+
+\textbf{TODO}
+
+Logical Cryptanalysis and its encoding as SAT likely started under the work of
+F. Massacci in early 2000 with his paper titled ``Logical Cryptanalysis as a
+SAT Problem: Encoding and Analysis of the U.S. Data Encryption Standard''.
+
+\subsection{Trustworthy Computing}
+
+\textbf{TODO - Assigned @erozier}
+
+\section{Terminology \& Notation}  \label{Sec:Terminology}
 
@@
 
-\section{Measuring the Utility of a Collision}
+\section{Intuition on the Utility of a Collision} \label{Sec:Intuition}
 
@@
 
-\section{Techniques for Logical Cryptanalysis}
+\section{Techniques for Logical Cryptanalysis} \label{Sec:Logical}
 
@@
 H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
-collision classes have distance $1$ under this metric.
+collision classes have distance $1$ under this metric. We extend this distance
+function to include an \textit{absolute} distance, measuring the number of
+intermediate rounds with non-zero differences. We denote this as $\delta(C_1)$.
 
@@
     \label{table:distance}
-    \begin{tabular}{c c c c c c}
-        & X. W. & Y. S. & P. K. & H. D. & M. S. \\
-        X. Wang's & 0 & 21 & 26 & 27 & 12 \\
-        Y. Sasaki's & & 0 & 25 & 26 & 2 \\
-        P. Kasselman's &  &  & 0 & 1 & 25 \\
-        H. Dobbertin's &  &  &  & 0 & 26 \\
-        M. Schlaffer's &  &  &  &  & 0 \\
+    \begin{tabular}{c c c c c c c}
+        & X. W. & Y. S. & P. K. & H. D. & M. S. & Absolute \\
+        X. Wang's & 0 & 21 & 26 & 27 & 12 & 18 \\
+        Y. Sasaki's & & 0 & 25 & 26 & 2 & 16 \\
+        P. Kasselman's &  &  & 0 & 1 & 25 & 11 \\
+        H. Dobbertin's &  &  &  & 0 & 26 & 12 \\
+        M. Schlaffer's &  &  &  &  & 0 & 17 \\
     \end{tabular}
@@
 
-\section{New Collision}
+\section{New Collision} \label{Sec:Collision}
 
-\section{Discussion}
+\section{Discussion of Impacts} \label{Sec:Impacts}
 
@@
 
-\section{Conclusion}
+\section{Conclusion} \label{Sec:Conclusion}
 
@@
 % use section* for acknowledgment
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 
 The authors would like to thank the Department of Electrical and Computer
-Engineering and Nathan Mulbrook for providing hardware.
+Engineering and XXXXX XXXXX for providing hardware.
 

### 2017-11-29 dsn-2018/bare_conf.tex
Expanded section on related work
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 while J. Kelsey and B. Schneier proposed a method for finding second preimages
-in less than $2^n$, we note that this requires rather rather long messages
+in less than $2^n$, we note that this requires rather long messages
 \cite{SchneierSecondPreimage}. However, collision attacks are well within reach
@@
 
-\subsection{Differential Cryptanalysis}
+\subsection{On Differential Cryptanalysis}
 
-\textbf{TODO}
+\textbf{TODO - expand?}
 
@@
 
-\subsection{Logical Cryptanalysis}
+\subsection{On Logical Cryptanalysis}
 
-\textbf{TODO}
+\textbf{TODO - expand?}
 
 Logical Cryptanalysis and its encoding as SAT likely started under the work of
-F. Massacci in early 2000 with his paper titled ``Logical Cryptanalysis as a
-SAT Problem: Encoding and Analysis of the U.S. Data Encryption Standard''.
+F. Massacci in 1999 with his paper titled ``Using walk-SAT and rel-sat for
+cryptographic key search''. \cite{Massacci:1999:UWR:1624218.1624261}. However,
+much of the early work by F. Massacci was focused on symmetric and asymmetric
+ciphers (in \cite{Massacci:1999:UWR:1624218.1624261},
+\cite{MASSACCI2000LogicalCA}, and \cite{FIORINI2003101}). It wasn't until the
+work of D. Jovanovi{\'{c}} and P. Jani{\v{c}}i{\'{c}} in \cite{LogicalAnalysis}
+that exploring collisions in the context of SAT was introduced, and until
+the work of I. Mironov and L. Zhang in \cite{Mironov2006} that this was
+studied for a X. Wang's collision class.
 
-\subsection{Trustworthy Computing}
+We make use of the \textit{bc2cnf} utility by T. Junttila (available here
+\cite{circuits}) as the basis for our models. This allows us to encode hash
+functions as generic structures in the circuit description language and
+ease the algorithmic creation of models which build on top of them. We then
+use CryptoMiniSat5 by M. Soos (\cite{CryptoMiniSat5}) to run the models and
+check for a satisfying witness, often an example collision.
+
+With the exception of the work by I. Mironov, the authors find that much of
+the previous work in the field has been focused on evaluating the performance
+of various SAT solvers. We deviate from this by instead focusing on evaluating
+the properties of the hash function--not the SAT solver. We have developed new
+techniques for using a SAT solver to inductively build full collisions in MD4
+from collisions in reduced-round MD4 and to build addition examples of
+collision classes from a single instance of a collision in MD4.
+
+\subsection{On Trustworthy Computing}
 
@@
 \section{Intuition on the Utility of a Collision} \label{Sec:Intuition}
+
+\textbf{TODO - introductory paragraph}
 

### 2017-11-29 dsn-2018/bare_conf.tex
Update conventions and notations
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 the published attacks by X. Wang, Y. Sasaki, P. Kasselman, H. Dobbertin, and
-M. Schlaffer in MD4 have high utility. We expand on existing attacks by
+M. Schl{\"a}ffer in MD4 have high utility. We expand on existing attacks by
 developing a series of techniques based on logical cryptanalysis to find
@@
 H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
-M. Schlaffer \cite{Schlaffer2006} and others demonstrate the ease with which
+M. Schl{\"a}ffer \cite{Schlaffer2006} and others demonstrate the ease with which
 collisions can be found.
@@
 
-\textbf{TODO}.
+\subsection{Conventions}
+\begin{itemize}
+    \item We assume ordered tuples and strings are indexable via square
+        brackets. Occasionally we use subscripts for referring to
+        elements of tuples with named members.
+\end{itemize}
 
-$\mathcal{C}$ for collision class. Ordered tuple, indexable.
-$\mathcal{F}$ for family of collision classes. Ordered set, indexable.
-$[i]$ is an indexing function.
-$R$ is number of rounds of $F$ or $C$.
+\textbf{TODO}
+
+\subsection{Collisions in Hash Functions}
+
+We use the following terminology when discussing the notion of collisions in
+a hash function.
+
+\begin{itemize}
+    \item A \textit{cryptographic hash function} is a function:
+        \begin{align*}
+            h: S_{i} \times B \rightarrow S_{o}
+        \end{align*}
+        which satisfies the usual properties of cryptographic hash functions
+        (preimage, second preimage, collision, deterministic, etc.). We denote
+        the input state space as $S_{i}$, the input block space as $B$, and the
+        output state space as $S_{o}$. Usually the input state space and the
+        output state space are the same, so we omit the subscripts, $S$. We
+        assume that these are sets of binary strings of a fixed length. In the
+        case of MD4, $S$ is all binary strings of length 128, and $B$ is all
+        binary strings of length 512.
+
+    \item An \textit{input} to a hash function is an ordered pair, $i = (s, b)$,
+        where $s \in S$ and $b \in B$. The \textit{input space} is simply
+        the domain of the hash function, which we denote as $\mathcal{I}$.
+        If $i \in \mathcal{I}$ is an input, we reference the state as $i_s$
+        and the input block as $i_b$.
+
+    \item We use the term \textit{intermediate state variables} to refer to
+        intermediate steps in the computation of a hash function under
+        a specific input. Typically these are the outputs of
+        one-way compression functions as part of the Merkel-Damg{\aa}rd
+        construct. If $i$ is an input to a hash function, we use $I(i)$ to
+        denote the intermediate state variables. Formally, this is represented
+        as an ordered tuple of binary strings of the size of the updated state.
+        In the case of MD4, this is an ordered tuple of cardinality 48, where
+        each index is a binary string of length 32, for a total of 1536
+        intermediate state variables. We denote the space that $I(i)$ maps
+        into as $\mathcal{V}$. We define the number of rounds as $R$.
+
+    \item We use the notation $R(I)$
+
+    \item A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
+        input space of a hash function with $\big|L\big| \geq 2$, such that
+        there exists an output $v \in S$ such that for all inputs $i \in L$,
+        $h(i) = c$.
+
+        \begin{itemize}
+            \item We define a \textit{simple collision} to be any collision
+                with cardinality exactly two.
+
+            \item We define a \textit{strong collision} to be a collision which
+                has multiple blocks which hash to the same output under the same
+                input  state. That is, if $L$ is a collision, then
+                $\forall i \in L$, $i_s = c$ for some $c \in S$ for $L$ to be
+                a strong collision.
+
+            \item We define a \textit{multicollision} to be a collision which
+                has multiple input states which hash to the same output under
+                a single input block. That is, if $L$ is a collision, then
+                $\forall i \in L$, $i_b = c$ for some $c \ in B$ for $L$ to be
+                a multicollision.
+        \end{itemize}
+
+    \item For a simple collision, we say that the \textit{differential path}
+        is the difference between the two sets of intermediate state
+        variables. \textbf{TODO -- HELP define formal notation?}
+            \begin{itemize}
+                \item Signed - Wang's, we don't use.
+                \item Unsigned - more common, actually used by us.
+            \end{itemize}
+
+    \item A \textit{collision class}, $C$, is an arbitrary differential path.
+        If there exists at least one collision pair which has that differential
+        path, we call that collision class non-empty. We denote the set of all
+        non-empty collision classes as $\mathcal{C}$. When we wish to convert
+        between a collision, $c \in L$ and a collision class,
+        $C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$.
+
+    \item A \textit{family of collision classes}, $F$, is the indices of a
+        collision class, $C$, which have any non-zero difference. That is,
+        $F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$.
+\end{itemize}
+
+\subsection{Known Collision Classes}
+
+We define the following set of known collision classes.
+
+\begin{itemize}
+    \item $C_{Wang}$, to be the collision class introduced by X. Wang
+        \textit{et al.} in \cite{Wang2005}.
+    \item $C_{Sasaki}$, to be the collision class introduced by Y. Sasaki
+        \textit{et al.} in \cite{Sasaki2007}.
+    \item $C_{Dobbertin}$, to be the collision class introduced by
+        H. Dobbertin in \cite{Dobbertin1998}.
+    \item $C_{Kasselman}$, to be the collision class introduced by
+        P. Kasselman in \cite{KasselmanMD4}.
+    \item $C_{Schlaffer}$, to be the collision class introduced by
+        M. Schl{\"a}ffer in \cite{Schlaffer2006}.
+\end{itemize}
+
+\subsection{Distance Metrics}
+
+We introduce a distance function, $\delta$ between collision classes by the
+number of differences in intermediate rounds deltas. That is, given two
+collision classes, $C_1, C_2 \in \mathcal{C}$:
+\begin{align}
+    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
+\end{align}
+
+We can define a similar distance function, $\Delta$, between families of
+collision classes by cardinality of the symmetric difference in the two
+collision families. That is, given two families of collision classes,
+$F_1, F_2 \in \mathcal{F}$:
+\begin{align}
+    \Delta(F_1, F_2) = \big| (F_1 \cup F_2) \setminus (F_1 \cap F_2) \big|
+\end{align}
+This is convenient for when the specifics of the differential path do not
+matter, merely that there exist at least one collision class of the specified
+form.
+
+\subsection{Neighborhoods}
+
+We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
+set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
+That is:
+\begin{align}
+    N(C, d) & = \{ C_j \in \mathcal{C} : \delta(C, C_j) = d \} \\
+    N(C) & = N(C, 1)
+\end{align}
+For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
+may optionally be omitted, in which case the unit distance is implied. Thus,
+$C_{Kasselman} \in N(C_{Dobbertin})$.
+
+We can similarly define the neighborhood of a family of collision classes,
+$F \in \mathcal{F}$, to be the set of all other families of collision classes
+at a fixed distance, $d \in \mathbb{N}$, from $F$. That is:
+\begin{align}
+    \mathcal{N}(F, d) & = \{ F_j \in \mathcal{F} : \delta(F, F_j) = d \} \\
+    \mathcal{N}(F) & = N(F, 1)
+\end{align}
+The distance parameter, $d$, may optionally be omitted, in which case the unit
+distance is implied.
+
+Neighborhoods can be classified into three types: \textit{expansion},
+\textit{internal}, and \textit{mixed}. Let $d$ be fixed. An \textit{expansion}
+neighborhood of a collision class, $C \in \mathcal{C}$, is the neighborhood
+restricted only to those collision classes which only differ in rounds external
+to the collision family of $C$. An \textit{internal} neighborhood of $C$ is the
+neighborhood restricted only to those collision classes which differ in rounds
+internal to the collision family of $C$. An \textit{mixed} neighborhood of $C$
+is the neighborhood restricted only to collisions which differ in a round
+internal and a round external to the collision family of $C$. That is:
+\begin{align*}
+    N_{exp}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \in F(C), C_j[i] = C[i] \} \\
+    N_{int}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \not\in F(C), C_j[i] = C[i] \} \\
+    N_{mix}(C, d) = \{ & C_{j} \in N(C, d) : \exists i \in F(C), C_j[i] \neq C[i] \\
+        & \text{ and } \exists k \not\in F(C), C_j[k] \neq C[k] \}
+\end{align*}
 
@@
 In the following sections, we use the collisions of X. Wang \cite{Wang2005},
-Y. Sasaki \cite{Sasaki2007}, M. Schlaffer \cite{Schlaffer2006}, H. Dobbertin
+Y. Sasaki \cite{Sasaki2007}, M. Schl{\"a}ffer \cite{Schlaffer2006}, H. Dobbertin
 \cite{Dobbertin1998}, and P. Kasselman \cite{KasselmanMD4} for examples.
@@
 
-We introduce a distance function, $\delta$ between collision classes by the
-number of differences in intermediate rounds deltas. That is, given two
-collision classes, $C_1, C_2 \in \mathcal{C}$:
-\begin{align}
-    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
-\end{align}
 We find justification for this metric in the existing literature on MD4:
@@
         H. Dobbertin's &  &  &  & 0 & 26 & 12 \\
-        M. Schlaffer's &  &  &  &  & 0 & 17 \\
+        M. Schl{\"a}ffer's &  &  &  &  & 0 & 17 \\
     \end{tabular}
 \end{table}
-
-We can define a similar distance function, $\Delta$, between families of
-collision classes by cardinality of the symmetric difference in the two
-collision families. That is, given two families of collision classes,
-$F_1, F_2 \in \mathcal{F}$:
-\begin{align}
-    \Delta(F_1, F_2) = \big| (F_1 \cup F_2) \setminus (F_1 \cap F_2) \big|
-\end{align}
-This is convenient for when the specifics of the differential path do not
-matter, merely that there exist at least one collision class of the specified
-form.
-
-\subsection{Neighborhoods}
-
-We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
-set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
-That is:
-\begin{align}
-    N(C, d) & = \{ C_j \in \mathcal{C} : \delta(C, C_j) = d \} \\
-    N(C) & = N(C, 1)
-\end{align}
-For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
-may optionally be omitted, in which case the unit distance is implied. Thus,
-$C_{Kasselman} \in N(C_{Dobbertin})$.
-
-The implications of collision class neighborhoods are discussed further in
-section \ref{empirical:neighborhood}.
-
-We can similarly define the neighborhood of a family of collision classes,
-$F \in \mathcal{F}$, to be the set of all other families of collision classes
-at a fixed distance, $d \in \mathbb{N}$, from $F$. That is:
-\begin{align}
-    \mathcal{N}(F, d) & = \{ F_j \in \mathcal{F} : \delta(F, F_j) = d \} \\
-    \mathcal{N}(F) & = N(F, 1)
-\end{align}
-The distance parameter, $d$, may optionally be omitted, in which case the unit
-distance is implied.
-
-Neighborhoods can be classified into three types: \textit{expansion},
-\textit{internal}, and \textit{mixed}. Let $d$ be fixed. An \textit{expansion}
-neighborhood of a collision class, $C \in \mathcal{C}$, is the neighborhood
-restricted only to those collision classes which only differ in rounds external
-to the collision family of $C$. An \textit{internal} neighborhood of $C$ is the
-neighborhood restricted only to those collision classes which differ in rounds
-internal to the collision family of $C$. An \textit{mixed} neighborhood of $C$
-is the neighborhood restricted only to collisions which differ in a round
-internal and a round external to the collision family of $C$. That is:
-\begin{align*}
-    N_{exp}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \in F(C), C_j[i] = C[i] \} \\
-    N_{int}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \not\in F(C), C_j[i] = C[i] \} \\
-    N_{mix}(C, d) = \{ & C_{j} \in N(C, d) : \exists i \in F(C), C_j[i] \neq C[i] \\
-        & \text{ and } \exists k \not\in F(C), C_j[k] \neq C[k] \}
-\end{align*}
 
@@
 Refer to Table \ref{table:differentials} for the number of differential
-paths...
+paths.
 
@@
         H. Dobbertin's & 32 & P. Kasselman's & 32 \\
-        M. Schlaffer's & 64 & & \\
+        M. Schl{\"a}ffer's & 64 & & \\
     \end{tabular}
@@
         H. Dobbertin's & 55 & P. Kasselman's & 60 \\
-        M. Schlaffer's & 100 & & \\
+        M. Schl{\"a}ffer's & 100 & & \\
     \end{tabular}
@@
 \textbf{TODO}
-
-Include graphic of count of max number of zeros/ones for all 35k collision
-classes.
 
@@
 P. Kasselman's and H. Dobbertin's results do not hold for the latter attacks
-by X. Wang, Y. Sasaki, or M. Schlaffer. Thus, while the latter attacks have
+by X. Wang, Y. Sasaki, or M. Schl{\"a}ffer. Thus, while the latter attacks have
 been viewed as being of better quality, under this particular metric,
@@
         H. Dobbertin's & true & \textbf{true} \\
-        M. Schlaffer's & true & false \\
+        M. Schl{\"a}ffer's & true & false \\
     \end{tabular}
@@
 
-\section{New Collision} \label{Sec:Collision}
+\section{New Collisions} \label{Sec:Collision}
+
+Through the exploration of the neighborhood technique, we have found over
+35,000 new differential paths which originated from the five original
+collisions in MD4. This shows that the neighborhood definition is sufficient
+to produce a large quantity of new collision classes.
 

### 2017-11-29 dsn-2018/bare_conf.tex
Extend new collisions sections
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 collisions in MD4. This shows that the neighborhood definition is sufficient
-to produce a large quantity of new collision classes.
+to produce a large quantity of new collision classes. Further, we give
+evidence for propagating neighborhoods across reduced round versions of MD4,
+and show a new collision in MD4 with new differential path.
+
+\subsection{Extensions of Prior Work}
+
+We noted in table \ref{table:neighborhood} that the unit-step neighborhood of
+the existing attacks were non-empty. By repeatedly expanding neighborhoods of
+known collision classes, we can construct paths in the collision space. We
+made the following observations:
+
+\begin{enumerate}
+    \item The space of collision classes is not smooth under the unit distance
+        relation. That is, starting at $C_{Schlaffer}$ and evaluating
+        successive neighborhoods, moving towards $C_{Wang}$ whenever the
+        distance decreases does not lead to a simple path of length
+        $\delta(C_{Schlaffer}, C_{Wang}) = 17$. Our lower bound on the actual
+        distance is 29, but after 14 rounds of expansion, we only reduced the
+        distance to 15.
+    \item Even under successive unit neighborhoods, the family of input block
+        differentials remains roughly the same among. This can be partially
+        counteracted by moving to reduced-round spaces, evaluating the
+        neighborhood, and expanding back to a collision in full-round MD4.
+        That is, there is a detectable signature of sharing significant
+        input block differential structure with the original collision.
+        With 35,918 unique classes of collisions which had 176 unique
+        families of collisions, there were only three unique families of
+        input block differentials: those of the starting collision classes.
+\end{enumerate}
+
+This last point show that, while there may be a number of unique differentials
+for a given collision class, and potentially many collision classes within
+some neighborhood of the original collision class, by analyzing the structure
+of the input block differential, systems can potentially prohibit successful
+collision attacks by preventing structures of input block differences.
+
+\textbf{TODO - figures and graphics on dataset}
+
+\subsection{New Differentials}
+
+We present the following new differentials as part of our work:
+
+\textbf{TODO - figures and graphics on new differentials/differential paths}
 
@@
 
-\textbf{TODO} - Discuss impacts on real-world systems like MD5, SHA-1, etc.
+\textbf{TODO} - Recap impacts on real-world systems like MD5, SHA-1, etc.
 

### 2017-11-29 dsn-2018/bare_conf.tex
More intuition
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-\textbf{TODO - introductory paragraph}
-
-We propose the following metrics for evaluating the utility of a collision
+\textbf{TODO - audit introductory paragraphs}
+
+We seek to build new intuition about collisions in hash functions,
+specifically, collisions in MD4. By developing techniques from logical
+cryptanalysis, we seek to complement and extend the results from
+differential cryptanalysis. In particular, we wish to use logical cryptanalysis
+to discuss the impacts of collisions in real-world systems. In this paper,
+we introduce and give justification for general techniques, and present some
+results of these techniques.
+
+To begin, real world systems are most vulnerable to two attacks: the preimage
+and second preimage. For systems relying on e.g., password validation via
+comparing two hashes, preimage attacks would allow inverting the hash,
+increasing the risk associated with losing a database of hashed passwords.
+On the other hand, systems relying on message validation, such as signature
+checks or file integrity checks would be easily broken with fast second
+preimage techniques. Both of these attacks are intractable, thus protecting
+real-world systems. However, much research has been devoted to collision
+attacks, producing efficient results.
+
+Thus, we seek to evaluate the relative difference between a collision
+attack and second preimage attacks, to give some measure of distance between
+them. Several high profile attacks against hash function usage have been
+made, including \textbf{TODO -- cite}, with MD5, cloning a CA certificate
+and creating malicious executables, and with SHA-1, a pair of PDFs with the
+same hash. However, many of these techniques exploit not the hash, but the
+flexibility of the file format, relying instead on embedding arbitrary,
+collidable data and later checking for the presence of one of the matching
+blocks. This is especially obvious in the MD5 executable and SHA-1 PDF attacks.
+In some cases, such as trying to produce a rogue CA certificate that matches
+an existing root, this is not possible and thus relies on searching a wide
+enough set of objects for a possible match to the collision.
+
+Seeing the need to evaluate how much flexibility exists in a collision,
+we propose the following metrics for evaluating the utility of a collision
 class:
@@
 \section{Conclusion} \label{Sec:Conclusion}
-
 The conclusion goes here.

### 2017-11-30 dsn-2018/bare_conf.tex
Changes suggested by KMayo
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-Cryptographic hash functions form the core of many protocols. From file
-integrity checks to verify long term storage of data, to cache invalidation
-techniques, and use as a building block in network protocols such as Kerberos
-and TLS, this class of functions necessarily has strong guarantees about
-properties of its members. There are three basic properties hash
+Cryptographic hash functions form the core of many protocols. With widespread
+use as file integrity checks to verify long term storage of data, as cache
+invalidation techniques, and as a building block in network protocols such as
+Kerberos and TLS, this class of functions necessarily has strong guarantees
+about properties of its members. There are three basic properties hash
 functions must have to be considered cryptographically secure:
@@
 faster than the theoretical $2^{80}$ but still requiring significant compute
-resources.
+resources to find.
 
 We seek to address the problem of evaluating the use cases for a collision,
-to determine--or partially determine--whether or not a collision affects
-a particular system. This paper contributes several novel techniques to
-logical cryptanalysis and uses these techniques to derive metrics of the
-trustworthiness of hash functions. We claim the following results are novel
-in the field:
+to determine whether or not a collision affects a particular system. This
+paper contributes several novel techniques to logical cryptanalysis and uses
+these techniques to derive metrics of the trustworthiness of hash functions.
+We claim the following results are novel in the field:
 \begin{itemize}

### 2017-12-02 dsn-2018/bare_conf.tex
Added collisions to paper
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>
Updates to introduction, related work
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 Cryptographic hash functions form the core of many protocols. With widespread
-use as file integrity checks to verify long term storage of data, as cache
-invalidation techniques, and as a building block in network protocols such as
-Kerberos and TLS, this class of functions necessarily has strong guarantees
-about properties of its members. There are three basic properties hash
-functions must have to be considered cryptographically secure:
+use of hash functions as file integrity checks to verify long term storage of
+data, as cache invalidation techniques, and as a building block in network
+protocols such as Kerberos and TLS, this class of functions necessarily has
+strong guarantees about properties of its members. There are three properties
+hash functions must have to be considered cryptographically secure:
 \begin{itemize}
@@
 \end{itemize}
-Note that a second preimage is necessarily a collision and is also a stronger
-result in practice.
+Note that while a second preimage is necessarily a collision, it is also a
+stronger result with higher use for an attacker.
 
 From an attacker's perspective, finding a preimage or second preimage has been
-difficult. To the author's knowledge, the current bound for a preimage attack
-in MD5 is $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. Further,
-while J. Kelsey and B. Schneier proposed a method for finding second preimages
+difficult: the current published bound for a preimage attack in MD5 is a cost
+of $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. Further, while
+J. Kelsey and B. Schneier proposed a method for finding second preimages
 in less than $2^n$, we note that this requires rather long messages
 \cite{SchneierSecondPreimage}. However, collision attacks are well within reach
-for adversaries, and well past only theoretical attacks. The work of
+of adversaries, and well past theoretical attacks. The work of
 H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
@@
 to determine whether or not a collision affects a particular system. This
-paper contributes several novel techniques to logical cryptanalysis and uses
-these techniques to derive metrics of the trustworthiness of hash functions.
-We claim the following results are novel in the field:
+paper contributes several techniques to logical cryptanalysis and uses these
+techniques to derive metrics of the trustworthiness of hash functions. We claim
+the following results are novel in the field:
 \begin{itemize}
@@
 \end{itemize}
-
-\textbf{TODO - Transition}
-
 We propose that collision classes of high utility are closer to a second
@@
 whereas classes of low utility may not affect all systems which use a
-particular hash function. Furthermore, we demonstrate new techniques for
-working with collisions under a logical cryptanalysis framework. Throughout
-this paper, we make use of the MD4 hash function due to the ease of generating
-new classes of collisions.
+particular hash function.
 
 \textbf{TODO - Finish Organization}
-
-The remainder of this paper is organized as follows. In Section
+The remainder of this paper is organized as fol
+lows. In Section
 \ref{Sec:Related} we discuss previous results and show how our research
@@
 
-\textbf{Broad context... intersection... subsection introductions.}
-
-This paper draws upon the public cryptanalysis of MD4, the wide availability of
-high quality SAT solvers, and previous work in logical cryptanalysis.
+This paper draws heavily upon the public cryptanalysis of MD4, the wide
+availability of high quality SAT solvers, and previous work in logical
+cryptanalysis. In particular, we look at ways that SAT solving can aid
+cryptanalysis instead of using cryptanalysis to benchmark SAT solvers. We
+then apply them to aid the understanding of attacks against real world
+systems, and propose general metrics that target different types of systems.
+Towards this, we look at prior work in two main areas: differential and logical
+cryptanalysis.
 
 \subsection{On Differential Cryptanalysis}
-
-\textbf{TODO - expand?}
 
@@
 differential cryptanalysis plays an important role, techniques for automated
-analysis of hash functions are still an emerging area. In
-\cite{cryptoeprint:2004:199}, X. Wang claimed to have found the MD4 collision
-``with hand calculation'', though in \cite{Wang2005}, she specifies message
-modification techniques which can be implemented in code.
+analysis of hash functions are still an emerging area. Further, techniques
+are being developed to make hash functions impervious to differential
+analysis.
 
@@
 model, such as a chosen prefix or desired start state. We believe performing
-these attacks by hand to be difficult, and note that message modification
-techniques may affect the chosen prefix.
+these attacks by hand to be difficult, and note that developed message
+modification techniques may affect chosen prefixes. Thus, we seek to develop
+techniques to replace differential cryptanalysis with logical cryptanalysis.
 
 \subsection{On Logical Cryptanalysis}
-
-\textbf{TODO - expand?}
 
@@
 much of the early work by F. Massacci was focused on symmetric and asymmetric
-ciphers (in \cite{Massacci:1999:UWR:1624218.1624261},
-\cite{MASSACCI2000LogicalCA}, and \cite{FIORINI2003101}). It wasn't until the
-work of D. Jovanovi{\'{c}} and P. Jani{\v{c}}i{\'{c}} in \cite{LogicalAnalysis}
-that exploring collisions in the context of SAT was introduced, and until
-the work of I. Mironov and L. Zhang in \cite{Mironov2006} that this was
-studied for a X. Wang's collision class.
-
-We make use of the \textit{bc2cnf} utility by T. Junttila (available here
-\cite{circuits}) as the basis for our models. This allows us to encode hash
-functions as generic structures in the circuit description language and
-ease the algorithmic creation of models which build on top of them. We then
-use CryptoMiniSat5 by M. Soos (\cite{CryptoMiniSat5}) to run the models and
-check for a satisfying witness, often an example collision.
-
-With the exception of the work by I. Mironov, the authors find that much of
-the previous work in the field has been focused on evaluating the performance
-of various SAT solvers. We deviate from this by instead focusing on evaluating
-the properties of the hash function--not the SAT solver. We have developed new
+ciphers
+\cite{Massacci:1999:UWR:1624218.1624261, MASSACCI2000LogicalCA, FIORINI2003101}.
+It wasn't until the work of D. Jovanovi{\'{c}} and P. Jani{\v{c}}i{\'{c}}
+that exploring collisions in the context of SAT was introduced
+\cite{LogicalAnalysis}, and until the work of I. Mironov and L. Zhang that
+this was studied for an existing collision class \cite{Mironov2006}.
+
+However, much of the work in this area is focused on benchmarking SAT solvers,
+and not on evaluating techniques for using SAT solvers to aid reasoning and
+understanding. E. Homsirikamol et al.\ show that SAT solvers can be used to
+brute force finding collisions in SHA-3, but with limited utility and limited
+to a small number of rounds \cite{Homsirikamol2012}. On the other hand,
+advances by V. Nossum in his master's thesis shows that additional work on
+encoding is also important: a new 32-bit modular addition circuit produced
+shorter run times \cite{nossum_2012}.
+
+We make use of the \textit{bc2cnf} utility by T. Junttila as the basis for
+our models \cite{circuits}. This allows us to encode hash functions as generic
+structures in the circuit description language and ease the algorithmic
+creation of models which build on top of them. We then use CryptoMiniSat5 by
+M. Soos to run the models and check for a satisfying witness, often an
+example collision \cite{CryptoMiniSat5}. We deviate from evaluating the
+performance of SAT solvers or encodings of models by instead focusing on
+evaluating the properties of the hash function. We have developed new
 techniques for using a SAT solver to inductively build full collisions in MD4
@@
     \item $C_{Wang}$, to be the collision class introduced by X. Wang
-        \textit{et al.} in \cite{Wang2005}.
+        et al. in \cite{Wang2005}.
     \item $C_{Sasaki}$, to be the collision class introduced by Y. Sasaki
-        \textit{et al.} in \cite{Sasaki2007}.
+        et al. in \cite{Sasaki2007}.
     \item $C_{Dobbertin}$, to be the collision class introduced by
@@
 
+\section{ASCII Collisions}
+
+\subsection{Dobbertin}
+
+Default IV
+
+Block 1:
+
+3f474422
+624f2b46
+683b6c7e
+48227d5c
+4f2a4f61
+206c573e
+61622234
+227a2633
+53482622
+706a6727
+503e2b3e
+7a636c76
+7\textbf{b}752620
+3f2e5728
+236a6376
+45654a75
+
+Block 2:
+
+3f474422
+624f2b46
+683b6c7e
+48227d5c
+4f2a4f61
+206c573e
+61622234
+227a2633
+53482622
+706a6727
+503e2b3e
+7a636c76
+7\textbf{c}752620
+3f2e5728
+236a6376
+45654a75
+
+\subsection{Kasselman}
+
+Default IV
+
+Block 1:
+
+7e273957
+597e7f4a
+7c773d7c
+4034376f
+69643257
+515f5269
+6a4d665a
+29416b3a
+7f377a64
+7e232455
+6d436c56
+26717321
+3\textbf{6}65774f
+6f2f4421
+77787656
+26236479
+
+7e273957
+597e7f4a
+7c773d7c
+4034376f
+69643257
+515f5269
+6a4d665a
+29416b3a
+7f377a64
+7e232455
+6d436c56
+26717321
+3\textbf{7}65774f
+6f2f4421
+77787656
+26236479
+
 \section{Discussion of Impacts} \label{Sec:Impacts}

### 2017-12-03 dsn-2018/bare_conf.tex
Minor word changes
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-We introduce a distance function, $\delta$ between collision classes by the
+We introduce a distance function, $\delta$, between collision classes by the
 number of differences in intermediate rounds deltas. That is, given two
@@
 We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
-set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
-That is:
+set of all other non-empty collision classes at a fixed distance,
+$d \in \mathbb{N}$, from $C$. That is:
 \begin{align}
@@
 
-We seek to build new intuition about collisions in hash functions,
-specifically, collisions in MD4. By developing techniques from logical
+We seek to build new intuition about collisions in hash functions
+and particularly collisions in MD4. By developing techniques using logical
 cryptanalysis, we seek to complement and extend the results from
@@
 The following techniques have been extensively tested on MD4 and partially
-tested on MD5 and believe to apply fully to MD5. They may or may not apply
-to any later hash function, such as SHA-1, SHA-2, or SHA-3, and have not been
-tested yet.
-
-In the following sections, we use the collisions of X. Wang \cite{Wang2005},
-Y. Sasaki \cite{Sasaki2007}, M. Schl{\"a}ffer \cite{Schlaffer2006}, H. Dobbertin
-\cite{Dobbertin1998}, and P. Kasselman \cite{KasselmanMD4} for examples.
+tested on MD5 and believe to apply fully to MD5. They likely apply to any
+hash function with similar structure, but as of yet, are untested.
 

### 2017-12-03 dsn-2018/bare_conf.tex
Extend Techniques section
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-\subsection{Distance Metrics}
+\subsection{Distance Metrics} \label{terminology:distance}
 
@@
 \end{align}
+We extend this distance function to include an \textit{absolute} distance,
+measuring the number of intermediate rounds with non-zero differences.
+We notate this as $\delta(C_1)$.
 
@@
 possible collision classes to search for a second preimage, chosen prefix, or
-other structure desired in a collision. If however, a collision class has no
+other structure desired in a collision. If, however, a collision class has no
 neighbors, then it cannot be used to find other possible classes for other
@@
 flexible and thus more likely to be used to target deployed systems. If,
-however, a collision class does not satisfy these properties, it impact is
+however, a collision class does not satisfy these properties, its impact is
 likely severely limited in scope, and may not provide useful information to
@@
 
-We find justification for this metric in the existing literature on MD4:
+We find justification for the distance metric defined in Section
+\ref{terminology:distance} in the existing literature on MD4:
 H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
-collision classes have distance $1$ under this metric. We extend this distance
-function to include an \textit{absolute} distance, measuring the number of
-intermediate rounds with non-zero differences. We denote this as $\delta(C_1)$.
+collision classes have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
+under this metric. The distance function allows us to find collision classes
+based off of existing classes. By limiting ourselves to unit distances, we
+maximally reuse an existing collision and limit the amount of work the SAT
+solver needs to do. Further, rather than searching for any unit-distance
+new collision class, we can easily split our search to look in each of the
+48 rounds of MD4, allowing us to parallelize the search for neighbors.
 
@@
 \subsection{Family Similarity}
-We define a relation among families of collisions across rounds. Let $F_1$ and
-$F_2$ be two different families of collisions. Then we say that $F_1$ and $F_2$
-are \textit{similar}, and notate it $F_1 \lesssim F_2$, if $R(F_1) \leq R(F_2)$
-and $F_1 \subseteq F_2$. Note that, when $R(F_1) = R(F_2)$, this is equivalent
-to saying that $F_2$ is in some expansion neighborhood of $F_1$. Further, when
-$R(F_1) < R(F_2)$ and $F_1 = F_2$, then we say that $F_2$ is the
-\textit{trivial extension} of $F_1$.
-
-We claim that the following statements are true:
+
+We define a relation among families of collisions across rounds as follows.
+Let $F_1$ and $F_2$ be two different families of collisions. Then we say that
+$F_1$ and $F_2$ are \textit{similar}, and notate it $F_1 \lesssim F_2$, if
+$R(F_1) \leq R(F_2)$ and $F_1 \subseteq F_2$. Note that, when
+$R(F_1) = R(F_2)$, this is equivalent to saying that $F_2$ is in some
+expansion neighborhood of $F_1$. Further, when $R(F_1) < R(F_2)$ and
+$F_1 = F_2$, then we say that $F_2$ is the \textit{trivial extension} of
+$F_1$.
+
+We claim that the following statements are true for MD4:
 \begin{enumerate}
@@
 
-\textbf{TODO} - Describe justification? Later sections?
-
-\subsection{Class Similarity}
+From an intuitive standpoint, what this is says is that collisions have
+witnesses for the existence in reduced round versions of the function, and
+further, that it is sufficient to search reduced round versions to find
+candidates for extension into additional rounds. This likely derives from
+the iterative construction of MD4 and the particular choice of block schedule.
+
+\textbf{TODO -- Empirical justification / graphical representation?}
+
+\subsubsection{Collision Class Similarity}
 
@@
 collision classes instead of families of collision classes, but with less
-likelihood.
-
-\textbf{TODO} - Describe technique applied neighborhood extensions
-
-Give some reasoning for why it works, example graphs.
-
-\subsection{Miscellaneous}
-
-\textbf{TODO} - Describe chosen prefix, second preimage, ASCII, and other
-tips and tricks for faster model runs.
-
-\section{Empirical Results}
+likelihood. That is, we can transplant a $n$ round collision into a $n+4$
+round and consider a neighborhood around it to find a new collision.
+
+\subsubsection{Second Preimage Similarity}
+
+If the models are constrained for a second preimage, we note that attack again
+works, but requires searching wide and wider differential paths in small round
+sizes. In particular, despite strong evidence for this pattern holding in
+$n \leq 24$ rounds, outside of the collision family $\{0, 16\}$ in both 28 and
+32 round MD4--which does not extend any further--we have been unable to find
+any other differential path in those rounds by exhaustive brute force search
+up to a differential path of length 8. Further searches have been done of
+36, 40, 44, and 48 rounds up to a differential path of length 5 with a similar
+lack of results. Thus we believe the general technique to hold, but the size
+of the search space for a second preimage becomes prohibitively large, and
+further, that differential paths become sufficiently long to prohibit easy
+finding of second preimages.
+
+\subsubsection{Chosen Prefix Similarity}
+
+Unlike second preimage attacks, chosen prefix attacks have a higher likelihood
+of working, and give better results. However, we have found that most chosen
+prefixes work in the neighborhoods of existing collisions, and thus are not
+convinced that inductively building a chosen prefix collision is a useful
+technique by itself.
+
+\subsection{Unjustified Performance Observations}
+
+We share the general tips for working with models of MD4 from a SAT solver
+performance perspective, without justification for results. Note that these
+results may be limited in usefulness and may not be uniformly applicable.
+\begin{itemize}
+    \item Unless searching for a second preimage or chosen prefix with
+        sufficient number of bits ($\geq 128$), we find models perform better
+        leaving the initial state unconstrained.
+    \item Adding ASCII block constraints significantly and negatively impact
+        the performance of finding a collision.
+    \item Keeping small search radius on neighborhoods involving families
+        of collision classes is important for good performance. However,
+        once sufficient witnesses have been found across rounds, missing
+        families can be found by various filling-in techniques by applying
+        the $\lesssim$ relation.
+\end{itemize}
+
+\section{Empirical Results:\\The Utility of Existing Collisions}
+
+\textbf{TODO -- Introduction?}
+
 \subsection{Unique Differentials}
 
-\textbf{TODO}
-Refer to Table \ref{table:differentials} for the number of differential
-paths.
+Table \ref{table:differentials} states the
 

### 2017-12-03 dsn-2018/bare_conf.tex
Expanded Empirical Results section
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-Table \ref{table:differentials} states the
+Table \ref{table:differentials} states the number of unique differentials,
+including the specified one, which produce a differential path conforming to
+each of the collision classes. We note in particular that while Y. Sasaki's
+attack is generally an improvement upon X. Wang's, Y. Sasaki's attack has fewer
+input block differentials and thus less flexibility in this aspect. Further,
+M. Schl{\"a}ffer's attack -- also building off the work of X. Wang -- has the
+same number of differentials as thus is also more flexible than Y. Sasaki's.
 
@@
 
-\textbf{TODO}
-
 Refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
-of existing collisions. In particular, note that Y. Sasaki's attack--which
-claimed to improve upon X. Wang's attack--has a larger neighborhood size, and
-likewise with P. Kasselman's attack which improved upon H. Dobbertin's attack.
+of existing collisions. In particular, note that both Y. Sasaki's attack and
+M. Schl{\"a}ffer's attack -- which both claimed to improve upon X. Wang's
+attack--have a larger neighborhood sizes, and likewise with P. Kasselman's
+attack which improved upon H. Dobbertin's attack. Thus, the newer collisions
+occur in more dense subsets of the collision class space.
 
@@
 
-\subsection{Zeroes \& Ones}
-
-\textbf{TODO}
-
-Separate table for known attacks as data points. Examples as appendix.
+\subsection{Zeroes}
+
+See table \ref{table:zeroes} for the maximum count of zeroes in a block
+for existing collisions. Note that X. Wang's attack is the leader, with 509
+maximum zero bits in the collision pair. Further, while Y. Sasaki improved
+the bound in other areas, this was not one of them. Of note too,
+H. Dobbertin's and P. Kasselman's collisions are essentially identical in
+this metric.
+
+\begin{table}
+    \caption{Number of Zeroes in a Colliding Block}
+    \label{table:zeroes}
+    \begin{tabular}{c c c c}
+        \textbf{Attack} & \textbf{Count} & \textbf{Attack} & \textbf{Count} \\
+        X. Wang's & 509 & Y. Sasaki's & 494 \\
+        H. Dobbertin's & 504 & P. Kasselman's & 504 \\
+        M. Schl{\"a}ffer's & 506 & & \\
+    \end{tabular}
+\end{table}
 
@@
 
-\textbf{TODO}
-
-Include table detailing how many have multicollisions and how many don't.
-Hypothesis: all do since all existing datapoints do.
-
-Separate table for known attacks as data points. Examples as appendix.
+We note that while these collisions are state independent and include
+multicollisions, none of these attacks are always multicollisions. Many of the
+blocks we've found only collide under the initial state value. Because of this,
+these collisions are only useful if you know the current state where a
+colliding pair is to be injected. Thus, this is one area in which collisions
+could potentially improve.
 
@@
 
-\textbf{TODO}
-
-Include table detailing how many have ASCII blocks and how many don't.
-
-See Table \ref{table:ascii} for results. Note that the validation of
-P. Kasselman's and H. Dobbertin's results do not hold for the latter attacks
-by X. Wang, Y. Sasaki, or M. Schl{\"a}ffer. Thus, while the latter attacks have
-been viewed as being of better quality, under this particular metric,
-P. Kasselman's and H. Dobbertin's collision classes are better.
+See Table \ref{table:ascii} for results of finding ASCII block pairs. Note that
+this table validates P. Kasselman's and H. Dobbertin's results of finding
+two ASCII block messages. However, these results do not hold for later
+collisions by X. Wang, Y. Sasaki, or M. Schl{\"a}ffer. Thus, while the latter
+attacks have been viewed as being of better quality, under this particular
+metric, P. Kasselman's and H. Dobbertin's collision classes are better.
 
@@
 \end{table}
+
+\subsection{Recapitulation}
+
+Overall, we feel that while the given collision classes have value and utility,
+they do not constitute full second preimage attacks. Further, simple
+neighborhood search does not lead to collisions in every case. Therefore,
+additional techniques are necessary for expanding the collision neighborhood.
 

### 2017-12-03 dsn-2018/bare_conf.tex
Extended citations
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 difficult: the current published bound for a preimage attack in MD5 is a cost
-of $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. Further, while
+of $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. In MD4, the
+best second preimage attack by H. Yu et al. gives a bound of $2^{-56}$
+likelihood, but uses destructive message modification techniques to achieve
+collisions for all other messages \cite{cryptoeprint:2007:206}. Further, while
 J. Kelsey and B. Schneier proposed a method for finding second preimages
@@
 
-\textbf{TODO - Finish Organization}
-The remainder of this paper is organized as fol
-lows. In Section
+The remainder of this paper is organized as follows. In Section
 \ref{Sec:Related} we discuss previous results and show how our research
@@
 \ref{Sec:Logical}, we present new techniques in logical cryptanalysis to
-analyze collisions in hash functions.
+analyze collisions in hash functions. In Section \ref{Sec:Empirical}, we
+discuss the utility of existing collisions under the metrics described in
+\ref{Sec:Intuition}. In Section \ref{Sec:Collision}, we discuss new
+differential paths and the effectiveness of various techniques presented
+in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Impacts}, we
+discuss broader impacts of this research.
 
@@
 M. Soos to run the models and check for a satisfying witness, often an
-example collision \cite{CryptoMiniSat5}. We deviate from evaluating the
-performance of SAT solvers or encodings of models by instead focusing on
-evaluating the properties of the hash function. We have developed new
-techniques for using a SAT solver to inductively build full collisions in MD4
-from collisions in reduced-round MD4 and to build addition examples of
-collision classes from a single instance of a collision in MD4.
+example collision \cite{Soos:2009:ESS:1575471.1575502, CryptoMiniSat5}. We
+deviate from evaluating the performance of SAT solvers or encodings of models
+by instead focusing on evaluating the properties of the hash function. We have
+developed new techniques for using a SAT solver to inductively build full
+collisions in MD4 from collisions in reduced-round MD4 and to build addition
+examples of collision classes from a single instance of a collision in MD4.
 
@@
         elements of tuples with named members.
+    \item We use MD4 as specified in RFC 1320 \cite{rfc1320}.
 \end{itemize}
@@
 them. Several high profile attacks against hash function usage have been
-made, including \textbf{TODO -- cite}, with MD5, cloning a CA certificate
-and creating malicious executables, and with SHA-1, a pair of PDFs with the
-same hash. However, many of these techniques exploit not the hash, but the
-flexibility of the file format, relying instead on embedding arbitrary,
-collidable data and later checking for the presence of one of the matching
-blocks. This is especially obvious in the MD5 executable and SHA-1 PDF attacks.
-In some cases, such as trying to produce a rogue CA certificate that matches
-an existing root, this is not possible and thus relies on searching a wide
-enough set of objects for a possible match to the collision.
+made. For MD5, these include cloning a CA certificate \cite{Stevens2009}
+and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
+PDFs with the same hash \cite{cryptoeprint:2017:190}. However, many of these
+techniques exploit not the hash, but the flexibility of the file format,
+relying instead on embedding arbitrary, collidable data and later checking for
+the presence of one of the matching blocks. This is especially obvious in the
+MD5 executable and SHA-1 PDF attacks. The work by M. Stevens to perform the
+CA Collision, on the other hand, was a chosen prefix attack extending
+previous work from 2007 \cite{Stevens2007}.
 
@@
 \section{Empirical Results:\\The Utility of Existing Collisions}
+\label{Sec:Empirical}
 
@@
 
-\subsection{New Differentials}
-
-We present the following new differentials as part of our work:
-
-\textbf{TODO - figures and graphics on new differentials/differential paths}
-
-\section{ASCII Collisions}
-
-\subsection{Dobbertin}
-
-Default IV
-
-Block 1:
-
-3f474422
-624f2b46
-683b6c7e
-48227d5c
-4f2a4f61
-206c573e
-61622234
-227a2633
-53482622
-706a6727
-503e2b3e
-7a636c76
-7\textbf{b}752620
-3f2e5728
-236a6376
-45654a75
-
-Block 2:
-
-3f474422
-624f2b46
-683b6c7e
-48227d5c
-4f2a4f61
-206c573e
-61622234
-227a2633
-53482622
-706a6727
-503e2b3e
-7a636c76
-7\textbf{c}752620
-3f2e5728
-236a6376
-45654a75
-
-\subsection{Kasselman}
-
-Default IV
-
-Block 1:
-
-7e273957
-597e7f4a
-7c773d7c
-4034376f
-69643257
-515f5269
-6a4d665a
-29416b3a
-7f377a64
-7e232455
-6d436c56
-26717321
-3\textbf{6}65774f
-6f2f4421
-77787656
-26236479
-
-7e273957
-597e7f4a
-7c773d7c
-4034376f
-69643257
-515f5269
-6a4d665a
-29416b3a
-7f377a64
-7e232455
-6d436c56
-26717321
-3\textbf{7}65774f
-6f2f4421
-77787656
-26236479
-
 \section{Discussion of Impacts} \label{Sec:Impacts}

### 2017-12-03 dsn-2018/bare_conf.tex
Changes suggested by KMayo
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 terminology and notations we use throughout the remainder of the paper.
-In Section \ref{Sec:Intuition}, we give the intuition between measuring
+In Section \ref{Sec:Intuition}, we give the intuition behind measuring
 the utility of a collision and motivate why it is useful. In Section
@@
 
-Differential Cryptanalysis has been the major technique behind the discovery
+Differential cryptanalysis has been the major technique behind the discovery
 of collisions in cryptographic hash functions. Its importance can be seen
@@
 
-Logical Cryptanalysis and its encoding as SAT likely started under the work of
+Logical cryptanalysis and its encoding as SAT likely started under the work of
 F. Massacci in 1999 with his paper titled ``Using walk-SAT and rel-sat for
@@
 developed new techniques for using a SAT solver to inductively build full
-collisions in MD4 from collisions in reduced-round MD4 and to build addition
+collisions in MD4 from collisions in reduced-round MD4 and to build additional
 examples of collision classes from a single instance of a collision in MD4.
@@
 This is convenient for when the specifics of the differential path do not
-matter, merely that there exist at least one collision class of the specified
+matter, merely that there exists at least one collision class of the specified
 form.
@@
 neighborhood restricted only to those collision classes which differ in rounds
-internal to the collision family of $C$. An \textit{mixed} neighborhood of $C$
+internal to the collision family of $C$. A \textit{mixed} neighborhood of $C$
 is the neighborhood restricted only to collisions which differ in a round
@@
 To begin, real world systems are most vulnerable to two attacks: the preimage
-and second preimage. For systems relying on e.g., password validation via
+and second preimage. For systems relying on password validation via
 comparing two hashes, preimage attacks would allow inverting the hash,
@@
 flexibility of a collision class, whereas the latter two are merely looking
-for a single witness for having the property. Depending on the scenario,
+for a single witness having such a property. Depending on the scenario,
 specific properties of a collision may be of more interest than others.
@@
 produce blocks with the given differential path. More differentials implies
-a greater flexibility in choice of colliding block, and possibly allowing
+a greater flexibility in choice of colliding block, and possibly allows
 for multiple collisions for a given colliding block. Furthermore, a collision
@@
 The following techniques have been extensively tested on MD4 and partially
-tested on MD5 and believe to apply fully to MD5. They likely apply to any
+tested on MD5 and are believed to apply fully to MD5. They likely apply to any
 hash function with similar structure, but as of yet, are untested.

### 2017-12-04 dsn-2018/bare_conf.tex
Changes suggested by KMayo
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
     \item We discuss techniques for finding useful collisions within a
-        class of collisions.
+        class of collisions using logical cryptanalysis.
     \item We propose additional techniques for building new classes of
@@
 solver needs to do. Further, rather than searching for any unit-distance
-new collision class, we can easily split our search to look in each of the
+new collision classes, we can easily split our search to look in each of the
 48 rounds of MD4, allowing us to parallelize the search for neighbors.
@@
 
-From an intuitive standpoint, what this is says is that collisions have
-witnesses for the existence in reduced round versions of the function, and
+From an intuitive standpoint, what this says is that collisions have
+witnesses for their existence in reduced round versions of the function, and
 further, that it is sufficient to search reduced round versions to find
@@
 collision classes instead of families of collision classes, but with less
-likelihood. That is, we can transplant a $n$ round collision into a $n+4$
+likelihood. That is, we can transplant an $n$ round collision into an $n+4$
 round and consider a neighborhood around it to find a new collision.
@@
 If the models are constrained for a second preimage, we note that attack again
-works, but requires searching wide and wider differential paths in small round
+works, but requires searching wider and wider differential paths in small round
 sizes. In particular, despite strong evidence for this pattern holding in
@@
         the performance of finding a collision.
-    \item Keeping small search radius on neighborhoods involving families
+    \item Keeping a small search radius on neighborhoods involving families
         of collision classes is important for good performance. However,
         once sufficient witnesses have been found across rounds, missing
-        families can be found by various filling-in techniques by applying
-        the $\lesssim$ relation.
+        families within the round can be found by various filling-in
+        techniques through applying the $\lesssim$ relation.
 \end{itemize}
@@
 Table \ref{table:differentials} states the number of unique differentials,
-including the specified one, which produce a differential path conforming to
+including the specified one, which produces a differential path conforming to
 each of the collision classes. We note in particular that while Y. Sasaki's
@@
 M. Schl{\"a}ffer's attack -- also building off the work of X. Wang -- has the
-same number of differentials as thus is also more flexible than Y. Sasaki's.
+same number of differentials and is thus also more flexible than Y. Sasaki's.
 
@@
 M. Schl{\"a}ffer's attack -- which both claimed to improve upon X. Wang's
-attack--have a larger neighborhood sizes, and likewise with P. Kasselman's
+attack--have larger neighborhood sizes, and likewise with P. Kasselman's
 attack which improved upon H. Dobbertin's attack. Thus, the newer collisions
@@
 they do not constitute full second preimage attacks. Further, simple
-neighborhood search does not lead to collisions in every case. Therefore,
+neighborhood searches do not lead to collisions in every case. Therefore,
 additional techniques are necessary for expanding the collision neighborhood.
@@
 to produce a large quantity of new collision classes. Further, we give
-evidence for propagating neighborhoods across reduced round versions of MD4,
-and show a new collision in MD4 with new differential path.
+evidence for propagating neighborhoods across reduced round versions of MD4.
 
@@
     \item Even under successive unit neighborhoods, the family of input block
-        differentials remains roughly the same among. This can be partially
-        counteracted by moving to reduced-round spaces, evaluating the
-        neighborhood, and expanding back to a collision in full-round MD4.
-        That is, there is a detectable signature of sharing significant
-        input block differential structure with the original collision.
-        With 35,918 unique classes of collisions which had 176 unique
-        families of collisions, there were only three unique families of
-        input block differentials: those of the starting collision classes.
+        differentials remains roughly the same among successive neighborhoods.
+        This can be partially counteracted by moving to reduced-round spaces,
+        evaluating the neighborhood, and expanding back to a collision in
+        full-round MD4. That is, there is a detectable signature of sharing
+        significant input block differential structure with the original
+        collision. With 35,918 unique classes of collisions which had 176
+        unique families of collisions, there were only three unique families
+        of input block differentials: those of the starting collision classes.
 \end{enumerate}
 
-This last point show that, while there may be a number of unique differentials
+This last point shows that, while there may be a number of unique differentials
 for a given collision class, and potentially many collision classes within

### 2017-12-05 dsn-2018/bare_conf.tex
DSN Figures
DSN progress

@@
 % http://www.tug.org/applications/pdftex
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{wrapfig}
 
@@
 
-Cryptographic hash functions form the core of many protocols. With widespread
-use of hash functions as file integrity checks to verify long term storage of
+Cryptographic hash functions form the core of many protocols which
+trust that cryptographic hash functions preserve certain basic
+properties. Cryptograpic hash functions see widespread
+use for file integrity checks to verify long term storage of
 data, as cache invalidation techniques, and as a building block in network
-protocols such as Kerberos and TLS, this class of functions necessarily has
-strong guarantees about properties of its members. There are three properties
-hash functions must have to be considered cryptographically secure:
+protocols such as Kerberos and TLS.  The classes of functions suitable
+for these use cases necessarily have
+strong guarantees about properties of its members. 
+An ideal cryptographic hash has five main features: determinism - the
+same message always results in the same hash; performability -
+computing a hash for a message should not be algorithmically
+difficult; one-way - it should be infeasible to generate a message
+from its hash except by trying all possible messages; sensitive to
+initial conditions - small changes to a message should change the hash
+value to a degree that the new hash value appears uncorrelated with
+the old hash value; unique digests - it should be infeasible to find
+two different messages with the same hash value.
+
+The most important aspect of these can be formally defined as three
+properties which hash functions must have to be considered cryptographically secure:
 \begin{itemize}
@@
 of $2^{123.4}$ by Y. Sasaki and K. Aoki \cite{SasakiPreimage}. In MD4, the
-best second preimage attack by H. Yu et al. gives a bound of $2^{-56}$
+best second preimage attack by H. Yu et al. gives a bound of $\frac{1}{2^{56}}$
 likelihood, but uses destructive message modification techniques to achieve
@@
 in less than $2^n$, we note that this requires rather long messages
-\cite{SchneierSecondPreimage}. However, collision attacks are well within reach
-of adversaries, and well past theoretical attacks. The work of
+\cite{SchneierSecondPreimage} which limits their utility to
+attackers. Collision attacks, however, are well within reach
+of adversaries, and have progressed past theoretical attacks into feasibility. The work of
 H. Dobbertin \cite{Dobbertin1998}, X. Wang \cite{cryptoeprint:2004:199},
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Figs/graph-intro-crop.pdf}
+\caption{A visualization of the heuristic unit-distance neighborhood
+  of valid collisions from the Dobbertin and Kasselman collisions.}
+\label{Fig:Graph-Intro}
+\end{center}
+\end{figure}
+
 From a cryptographic perspective, the existence of a feasible collision attack
-breaks the requisite properties, thus deprecating the function's use. In some
-instances however, there continues to be widespread use of deprecated hash
-functions. One widespread example is the continued use by Git of SHA-1, despite
+breaks the requisite properties, thus deprecating the function's use
+for trustworthy applications. In many practical instances, however,
+communities continue to use deprecated hash functions in a widespread fashion.
+One such example of widespread use post-deprecation is the continued use by Git of SHA-1, despite
 M. Stevens et al. publishing a full SHA-1 collision in February of 2017
-\cite{cryptoeprint:2017:190}. However, this attack had a bound of $2^{63.1}$ --
-faster than the theoretical $2^{80}$ but still requiring significant compute
+\cite{cryptoeprint:2017:190}. This attack had a bound of $2^{63.1}$
+\emdash faster than the theoretical $2^{80}$ but still requiring significant compute
 resources to find.
@@
 
+\begin{figure*}
+\begin{center}
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{Figs/differential-1289.png}
+    \caption{Collision 1289}
+    \label{Fig:1289}
+  \end{subfigure}\\
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{Figs/differential-1250.png}
+    \caption{Collision 1250}
+    \label{Fig:1250}
+  \end{subfigure}\\
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{Figs/transitions.png}
+    \caption{Differential Path Transitions}
+    \label{Fig:Transitions}
+  \end{subfigure}\\
+\end{center}
+\end{figure*}
+
 Differential cryptanalysis has been the major technique behind the discovery
@@
 \subsection{Unit-Step Neighborhood} \label{empirical:neighborhood}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Figs/graph-clique-crop.pdf}
+\caption{}
+\label{Fig:Graph-Clique}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Figs/graph-neighborhood-crop.pdf}
+\caption{}
+\label{Fig:Graph-Neighborhood}
+\end{center}
+\end{figure}
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
 computing a hash for a message should not be algorithmically
-difficult; one-way - it should be infeasible to generate a message
-from its hash except by trying all possible messages; sensitive to
+difficult; non-reversability - it should be infeasible to generate a message
+from its hash except by trying all possible messages; sensitivity to
 initial conditions - small changes to a message should change the hash
@@
 
-The most important aspect of these can be formally defined as three
-properties which hash functions must have to be considered cryptographically secure:
+The most important aspect of these can be formally defined under three
+properties which hash function implementations must have to be
+considered cryptographically secure in practice:
 \begin{itemize}
@@
 \cite{cryptoeprint:2017:190}. This attack had a bound of $2^{63.1}$
-\emdash faster than the theoretical $2^{80}$ but still requiring significant compute
+--- faster than the theoretical $2^{80}$ but still requiring significant compute
 resources to find.
@@
 We seek to address the problem of evaluating the use cases for a collision,
-to determine whether or not a collision affects a particular system. This
-paper contributes several techniques to logical cryptanalysis and uses these
-techniques to derive metrics of the trustworthiness of hash functions. We claim
+to determine whether or not a collision affects a particular system.
+The central thesis of this research is that logical cryptanalysis
+allows us to formalize the structure inherent to the underlying hash
+function such that new collisions can be found within reasonable time
+on commodity hardware.  We assert that when such structure exists, it
+is \textbf{imperitive to discontinue use of such hash functions as
+  they are fundamentally untrustworthy for critical applications.}  An
+example of such structure showing the unit distance neighborhood
+(defined in Section \ref{Sec:Terminology}) of the known collisions
+from Dobbertin and Kasselman is presented in Figure \ref{Fig:Graph-Intro}.
+This paper contributes several techniques to logical cryptanalysis and uses these
+techniques to derive metrics of the trustworthiness for hash functions. We claim
 the following results are novel in the field:
@@
 \subsection{On Differential Cryptanalysis}
+
+Differential cryptanalysis has been the major technique behind the discovery
+of collisions in cryptographic hash functions. Its importance can be seen
+everywhere from X. Wang's attacks on MD4 \cite{Wang2005} to more recent
+cryptanalysis of general purpose hash functions such as Murmur3 by
+J. Aumasson, D. Bernstein, and M. Bo{\ss}let \cite{murmur3DC}. However, while
+differential cryptanalysis plays an important role, techniques for automated
+analysis of hash functions are still an emerging area. Further, techniques
+are being developed to make hash functions impervious to differential
+analysis.
+
+Our work removes the need for finding message modification techniques by
+specifying the differential path as part of the SAT formula and letting the
+solver find a pair of satisfying messages which follow the differential path
+and produce a collision. Further constraints can then be placed upon this
+model, such as a chosen prefix or desired start state. We believe performing
+these attacks by hand to be difficult, and note that developed message
+modification techniques may affect chosen prefixes. Thus, we seek to develop
+techniques to replace differential cryptanalysis with logical cryptanalysis.
+
+\subsection{On Logical Cryptanalysis}
+
+Logical cryptanalysis and its encoding as SAT likely started under the work of
+F. Massacci in 1999 with his paper titled ``Using walk-SAT and rel-sat for
+cryptographic key search''. \cite{Massacci:1999:UWR:1624218.1624261}. However,
+much of the early work by F. Massacci was focused on symmetric and asymmetric
+ciphers
+\cite{Massacci:1999:UWR:1624218.1624261, MASSACCI2000LogicalCA, FIORINI2003101}.
+It wasn't until the work of D. Jovanovi{\'{c}} and P. Jani{\v{c}}i{\'{c}}
+that exploring collisions in the context of SAT was introduced
+\cite{LogicalAnalysis}, and until the work of I. Mironov and L. Zhang that
+this was studied for an existing collision class \cite{Mironov2006}.
+
+However, much of the work in this area is focused on benchmarking SAT solvers,
+and not on evaluating techniques for using SAT solvers to aid reasoning and
+understanding. E. Homsirikamol et al.\ show that SAT solvers can be used to
+brute force finding collisions in SHA-3, but with limited utility and limited
+to a small number of rounds \cite{Homsirikamol2012}. On the other hand,
+advances by V. Nossum in his master's thesis shows that additional work on
+encoding is also important: a new 32-bit modular addition circuit produced
+shorter run times \cite{nossum_2012}.
+
+We make use of the \textit{bc2cnf} utility by T. Junttila as the basis for
+our models \cite{circuits}. This allows us to encode hash functions as generic
+structures in the circuit description language and ease the algorithmic
+creation of models which build on top of them. We then use CryptoMiniSat5 by
+M. Soos to run the models and check for a satisfying witness, often an
+example collision \cite{Soos:2009:ESS:1575471.1575502, CryptoMiniSat5}. We
+deviate from evaluating the performance of SAT solvers or encodings of models
+by instead focusing on evaluating the properties of the hash function. We have
+developed new techniques for using a SAT solver to inductively build full
+collisions in MD4 from collisions in reduced-round MD4 and to build additional
+examples of collision classes from a single instance of a collision in MD4.
+
+\subsection{On Trustworthy Computing}
+
+\textbf{TODO - Assigned @erozier}
+
+\section{Terminology \& Notation}  \label{Sec:Terminology}
+
+\subsection{Conventions}
+\begin{itemize}
+    \item We assume ordered tuples and strings are indexable via square
+        brackets. Occasionally we use subscripts for referring to
+        elements of tuples with named members.
+    \item We use MD4 as specified in RFC 1320 \cite{rfc1320}.
+\end{itemize}
+
+\textbf{TODO}
+
+\subsection{Collisions in Hash Functions}
+
+We use the following terminology when discussing the notion of collisions in
+a hash function.
+
+\begin{itemize}
+    \item A \textit{cryptographic hash function} is a function:
+        \begin{align*}
+            h: S_{i} \times B \rightarrow S_{o}
+        \end{align*}
+        which satisfies the usual properties of cryptographic hash functions
+        (preimage, second preimage, collision, deterministic, etc.). We denote
+        the input state space as $S_{i}$, the input block space as $B$, and the
+        output state space as $S_{o}$. Usually the input state space and the
+        output state space are the same, so we omit the subscripts, $S$. We
+        assume that these are sets of binary strings of a fixed length. In the
+        case of MD4, $S$ is all binary strings of length 128, and $B$ is all
+        binary strings of length 512.
+
+    \item An \textit{input} to a hash function is an ordered pair, $i = (s, b)$,
+        where $s \in S$ and $b \in B$. The \textit{input space} is simply
+        the domain of the hash function, which we denote as $\mathcal{I}$.
+        If $i \in \mathcal{I}$ is an input, we reference the state as $i_s$
+        and the input block as $i_b$.
+
+    \item We use the term \textit{intermediate state variables} to refer to
+        intermediate steps in the computation of a hash function under
+        a specific input. Typically these are the outputs of
+        one-way compression functions as part of the Merkel-Damg{\aa}rd
+        construct. If $i$ is an input to a hash function, we use $I(i)$ to
+        denote the intermediate state variables. Formally, this is represented
+        as an ordered tuple of binary strings of the size of the updated state.
+        In the case of MD4, this is an ordered tuple of cardinality 48, where
+        each index is a binary string of length 32, for a total of 1536
+        intermediate state variables. We denote the space that $I(i)$ maps
+        into as $\mathcal{V}$. We define the number of rounds as $R$.
+
+    \item We use the notation $R(I)$
+
+    \item A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
+        input space of a hash function with $\big|L\big| \geq 2$, such that
+        there exists an output $v \in S$ such that for all inputs $i \in L$,
+        $h(i) = c$.
+
+        \begin{itemize}
+            \item We define a \textit{simple collision} to be any collision
+                with cardinality exactly two.
+
+            \item We define a \textit{strong collision} to be a collision which
+                has multiple blocks which hash to the same output under the same
+                input  state. That is, if $L$ is a collision, then
+                $\forall i \in L$, $i_s = c$ for some $c \in S$ for $L$ to be
+                a strong collision.
+
+            \item We define a \textit{multicollision} to be a collision which
+                has multiple input states which hash to the same output under
+                a single input block. That is, if $L$ is a collision, then
+                $\forall i \in L$, $i_b = c$ for some $c \ in B$ for $L$ to be
+                a multicollision.
+        \end{itemize}
+
+    \item For a simple collision, we say that the \textit{differential path}
+        is the difference between the two sets of intermediate state
+        variables. \textbf{TODO -- HELP define formal notation?}
+            \begin{itemize}
+                \item Signed - Wang's, we don't use.
+                \item Unsigned - more common, actually used by us.
+            \end{itemize}
+
+    \item A \textit{collision class}, $C$, is an arbitrary differential path.
+        If there exists at least one collision pair which has that differential
+        path, we call that collision class non-empty. We denote the set of all
+        non-empty collision classes as $\mathcal{C}$. When we wish to convert
+        between a collision, $c \in L$ and a collision class,
+        $C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$.
+
+    \item A \textit{family of collision classes}, $F$, is the indices of a
+        collision class, $C$, which have any non-zero difference. That is,
+        $F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$.
+\end{itemize}
+
+\subsection{Known Collision Classes}
+
+We define the following set of known collision classes.
+
+\begin{itemize}
+    \item $C_{Wang}$, to be the collision class introduced by X. Wang
+        et al. in \cite{Wang2005}.
+    \item $C_{Sasaki}$, to be the collision class introduced by Y. Sasaki
+        et al. in \cite{Sasaki2007}.
+    \item $C_{Dobbertin}$, to be the collision class introduced by
+        H. Dobbertin in \cite{Dobbertin1998}.
+    \item $C_{Kasselman}$, to be the collision class introduced by
+        P. Kasselman in \cite{KasselmanMD4}.
+    \item $C_{Schlaffer}$, to be the collision class introduced by
+        M. Schl{\"a}ffer in \cite{Schlaffer2006}.
+\end{itemize}
+
+\subsection{Distance Metrics} \label{terminology:distance}
+
+We introduce a distance function, $\delta$, between collision classes by the
+number of differences in intermediate rounds deltas. That is, given two
+collision classes, $C_1, C_2 \in \mathcal{C}$:
+\begin{align}
+    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
+\end{align}
+We extend this distance function to include an \textit{absolute} distance,
+measuring the number of intermediate rounds with non-zero differences.
+We notate this as $\delta(C_1)$.
+
+We can define a similar distance function, $\Delta$, between families of
+collision classes by cardinality of the symmetric difference in the two
+collision families. That is, given two families of collision classes,
+$F_1, F_2 \in \mathcal{F}$:
+\begin{align}
+    \Delta(F_1, F_2) = \big| (F_1 \cup F_2) \setminus (F_1 \cap F_2) \big|
+\end{align}
+This is convenient for when the specifics of the differential path do not
+matter, merely that there exists at least one collision class of the specified
+form.
+
+We introduce this distance function as a heuristic measure of the
+sensitivity of the hash function to small changes in the input.
+
+\subsection{Neighborhoods}
+
+We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
+set of all other non-empty collision classes at a fixed distance,
+$d \in \mathbb{N}$, from $C$. That is:
+\begin{align}
+    N(C, d) & = \{ C_j \in \mathcal{C} : \delta(C, C_j) = d \} \\
+    N(C) & = N(C, 1)
+\end{align}
+For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
+may optionally be omitted, in which case the unit distance is implied. Thus,
+$C_{Kasselman} \in N(C_{Dobbertin})$.
+
+We can similarly define the neighborhood of a family of collision classes,
+$F \in \mathcal{F}$, to be the set of all other families of collision classes
+at a fixed distance, $d \in \mathbb{N}$, from $F$. That is:
+\begin{align}
+    \mathcal{N}(F, d) & = \{ F_j \in \mathcal{F} : \delta(F, F_j) = d \} \\
+    \mathcal{N}(F) & = N(F, 1)
+\end{align}
+The distance parameter, $d$, may optionally be omitted, in which case the unit
+distance is implied.
+
+Neighborhoods can be classified into three types: \textit{expansion},
+\textit{internal}, and \textit{mixed}. Let $d$ be fixed. An \textit{expansion}
+neighborhood of a collision class, $C \in \mathcal{C}$, is the neighborhood
+restricted only to those collision classes which only differ in rounds external
+to the collision family of $C$. An \textit{internal} neighborhood of $C$ is the
+neighborhood restricted only to those collision classes which differ in rounds
+internal to the collision family of $C$. A \textit{mixed} neighborhood of $C$
+is the neighborhood restricted only to collisions which differ in a round
+internal and a round external to the collision family of $C$. That is:
+\begin{align*}
+    N_{exp}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \in F(C), C_j[i] = C[i] \} \\
+    N_{int}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \not\in F(C), C_j[i] = C[i] \} \\
+    N_{mix}(C, d) = \{ & C_{j} \in N(C, d) : \exists i \in F(C), C_j[i] \neq C[i] \\
+        & \text{ and } \exists k \not\in F(C), C_j[k] \neq C[k] \}
+\end{align*}
+
+When the neighborhood defined by $\delta = 1$ is frequently non-empty
+we assert that sufficient structure exists in the underlying hash
+function such that logical cryptanalysis via 3-CNF-SAT becomes
+feasible.  This condition is thus sufficient to render the underlying
+hash function \textbf{untrustworthy}.  Conversely it is a necessary
+condition (but possibly insufficient) that a trustworthy hash function
+has an empty neighborhood defined by $\delta = 1$.
+
+Furthermore this condition extends to $\delta = d$ for some small
+value of $d$ as determined by computational resources such that $d$
+becomes a trustworthiness metric under constraints of current
+3-CNF-SAT solving capabilities within reasonable time bounds.
+
+\section{Intuition on the Utility of a Collision} \label{Sec:Intuition}
+
+\textbf{TODO - audit introductory paragraphs}
+
+We seek to build new intuition about collisions in hash functions
+and particularly collisions in MD4. By developing techniques using logical
+cryptanalysis, we seek to complement and extend the results from
+differential cryptanalysis. In particular, we wish to use logical cryptanalysis
+to discuss the impacts of collisions in real-world systems. In this paper,
+we introduce and give justification for general techniques, and present some
+results of these techniques.
+
+To begin, real world systems are most vulnerable to two attacks: the preimage
+and second preimage. For systems relying on password validation via
+comparing two hashes, preimage attacks would allow inverting the hash,
+increasing the risk associated with losing a database of hashed passwords.
+On the other hand, systems relying on message validation, such as signature
+checks or file integrity checks would be easily broken with fast second
+preimage techniques. Both of these attacks are intractable, thus protecting
+real-world systems. However, much research has been devoted to collision
+attacks, producing efficient results.
+
+Thus, we seek to evaluate the relative difference between a collision
+attack and second preimage attacks, to give some measure of distance between
+them. Several high profile attacks against hash function usage have been
+made. For MD5, these include cloning a CA certificate \cite{Stevens2009}
+and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
+PDFs with the same hash \cite{cryptoeprint:2017:190}. However, many of these
+techniques exploit not the hash, but the flexibility of the file format,
+relying instead on embedding arbitrary, collidable data and later checking for
+the presence of one of the matching blocks. This is especially obvious in the
+MD5 executable and SHA-1 PDF attacks. The work by M. Stevens to perform the
+CA Collision, on the other hand, was a chosen prefix attack extending
+previous work from 2007 \cite{Stevens2007}.
+
+Seeing the need to evaluate how much flexibility exists in a collision,
+we propose the following metrics for evaluating the utility of a collision
+class:
+\begin{enumerate}
+    \item The number of unique differentials a collision class has.
+    \item The number of unit-step neighbors a collision class has.
+    \item The maximum count of zeros in a the binary representation of a
+        colliding block (and likewise with ones).
+    \item Whether there exists a block which collides under multiple initial
+        values.
+    \item Whether or not zero, one, or both of the blocks in a collision may
+        be of ASCII values under any input block difference.
+\end{enumerate}
+Note that the first three are quantitative measures providing some measure of
+flexibility of a collision class, whereas the latter two are merely looking
+for a single witness having such a property. Depending on the scenario,
+specific properties of a collision may be of more interest than others.
+
+The first metric evaluates the flexibility of the differential path. A
+differential path with more flexibility will have more differentials which
+produce blocks with the given differential path. More differentials implies
+a greater flexibility in choice of colliding block, and possibly allows
+for multiple collisions for a given colliding block. Furthermore, a collision
+with more differentials is more likely to satisfy the last metric, having a
+pair of blocks--both ASCII--which produce a collision.
+
+The second metric evaluates the density of the neighborhood of a collision
+class. If a collision resides in a dense neighborhood, it provides more
+possible collision classes to search for a second preimage, chosen prefix, or
+other structure desired in a collision. If, however, a collision class has no
+neighbors, then it cannot be used to find other possible classes for other
+input blocks.
+
+The maximum quantity of zeros (or ones) in the binary representation of a
+block serve as a measure of the extremes to which a collision can be pushed.
+This can additionally be extended to any suitable bit pattern in any base to
+provide a more relevant metric as desired by the system under study.
+
+The fourth metric evaluates the utility of a collision when the internal state
+of the hash function is unknown. If a collision occurs under multiple initial
+values, this could be used to attack some systems where user provided input is
+appended to unknown data and then hashed. If a block collides under a suitably
+large number of initial values, the attack becomes highly likely to occur
+successfully. However, measuring the exact number of initial values a block
+collides under is beyond the scope of this work.
+
+The last metric is similar to the third in that it looks for specific bit
+patterns in a collision. ASCII is one example of a widely used constraint
+system. Further examples, such as JSON, XML, etc., may likewise be
+supplemented based on the specifics of the system.
+
+If a collision class satisfies many of these properties, then it is more
+flexible and thus more likely to be used to target deployed systems. If,
+however, a collision class does not satisfy these properties, its impact is
+likely severely limited in scope, and may not provide useful information to
+find other collision classes which have higher utility.
+
+\section{Techniques for Logical Cryptanalysis} \label{Sec:Logical}
 
@@
 
-Differential cryptanalysis has been the major technique behind the discovery
-of collisions in cryptographic hash functions. Its importance can be seen
-everywhere from X. Wang's attacks on MD4 \cite{Wang2005} to more recent
-cryptanalysis of general purpose hash functions such as Murmur3 by
-J. Aumasson, D. Bernstein, and M. Bo{\ss}let \cite{murmur3DC}. However, while
-differential cryptanalysis plays an important role, techniques for automated
-analysis of hash functions are still an emerging area. Further, techniques
-are being developed to make hash functions impervious to differential
-analysis.
-
-Our work removes the need for finding message modification techniques by
-specifying the differential path as part of the SAT formula and letting the
-solver find a pair of satisfying messages which follow the differential path
-and produce a collision. Further constraints can then be placed upon this
-model, such as a chosen prefix or desired start state. We believe performing
-these attacks by hand to be difficult, and note that developed message
-modification techniques may affect chosen prefixes. Thus, we seek to develop
-techniques to replace differential cryptanalysis with logical cryptanalysis.
-
-\subsection{On Logical Cryptanalysis}
-
-Logical cryptanalysis and its encoding as SAT likely started under the work of
-F. Massacci in 1999 with his paper titled ``Using walk-SAT and rel-sat for
-cryptographic key search''. \cite{Massacci:1999:UWR:1624218.1624261}. However,
-much of the early work by F. Massacci was focused on symmetric and asymmetric
-ciphers
-\cite{Massacci:1999:UWR:1624218.1624261, MASSACCI2000LogicalCA, FIORINI2003101}.
-It wasn't until the work of D. Jovanovi{\'{c}} and P. Jani{\v{c}}i{\'{c}}
-that exploring collisions in the context of SAT was introduced
-\cite{LogicalAnalysis}, and until the work of I. Mironov and L. Zhang that
-this was studied for an existing collision class \cite{Mironov2006}.
-
-However, much of the work in this area is focused on benchmarking SAT solvers,
-and not on evaluating techniques for using SAT solvers to aid reasoning and
-understanding. E. Homsirikamol et al.\ show that SAT solvers can be used to
-brute force finding collisions in SHA-3, but with limited utility and limited
-to a small number of rounds \cite{Homsirikamol2012}. On the other hand,
-advances by V. Nossum in his master's thesis shows that additional work on
-encoding is also important: a new 32-bit modular addition circuit produced
-shorter run times \cite{nossum_2012}.
-
-We make use of the \textit{bc2cnf} utility by T. Junttila as the basis for
-our models \cite{circuits}. This allows us to encode hash functions as generic
-structures in the circuit description language and ease the algorithmic
-creation of models which build on top of them. We then use CryptoMiniSat5 by
-M. Soos to run the models and check for a satisfying witness, often an
-example collision \cite{Soos:2009:ESS:1575471.1575502, CryptoMiniSat5}. We
-deviate from evaluating the performance of SAT solvers or encodings of models
-by instead focusing on evaluating the properties of the hash function. We have
-developed new techniques for using a SAT solver to inductively build full
-collisions in MD4 from collisions in reduced-round MD4 and to build additional
-examples of collision classes from a single instance of a collision in MD4.
-
-\subsection{On Trustworthy Computing}
-
-\textbf{TODO - Assigned @erozier}
-
-\section{Terminology \& Notation}  \label{Sec:Terminology}
-
-\subsection{Conventions}
-\begin{itemize}
-    \item We assume ordered tuples and strings are indexable via square
-        brackets. Occasionally we use subscripts for referring to
-        elements of tuples with named members.
-    \item We use MD4 as specified in RFC 1320 \cite{rfc1320}.
-\end{itemize}
-
-\textbf{TODO}
-
-\subsection{Collisions in Hash Functions}
-
-We use the following terminology when discussing the notion of collisions in
-a hash function.
-
-\begin{itemize}
-    \item A \textit{cryptographic hash function} is a function:
-        \begin{align*}
-            h: S_{i} \times B \rightarrow S_{o}
-        \end{align*}
-        which satisfies the usual properties of cryptographic hash functions
-        (preimage, second preimage, collision, deterministic, etc.). We denote
-        the input state space as $S_{i}$, the input block space as $B$, and the
-        output state space as $S_{o}$. Usually the input state space and the
-        output state space are the same, so we omit the subscripts, $S$. We
-        assume that these are sets of binary strings of a fixed length. In the
-        case of MD4, $S$ is all binary strings of length 128, and $B$ is all
-        binary strings of length 512.
-
-    \item An \textit{input} to a hash function is an ordered pair, $i = (s, b)$,
-        where $s \in S$ and $b \in B$. The \textit{input space} is simply
-        the domain of the hash function, which we denote as $\mathcal{I}$.
-        If $i \in \mathcal{I}$ is an input, we reference the state as $i_s$
-        and the input block as $i_b$.
-
-    \item We use the term \textit{intermediate state variables} to refer to
-        intermediate steps in the computation of a hash function under
-        a specific input. Typically these are the outputs of
-        one-way compression functions as part of the Merkel-Damg{\aa}rd
-        construct. If $i$ is an input to a hash function, we use $I(i)$ to
-        denote the intermediate state variables. Formally, this is represented
-        as an ordered tuple of binary strings of the size of the updated state.
-        In the case of MD4, this is an ordered tuple of cardinality 48, where
-        each index is a binary string of length 32, for a total of 1536
-        intermediate state variables. We denote the space that $I(i)$ maps
-        into as $\mathcal{V}$. We define the number of rounds as $R$.
-
-    \item We use the notation $R(I)$
-
-    \item A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
-        input space of a hash function with $\big|L\big| \geq 2$, such that
-        there exists an output $v \in S$ such that for all inputs $i \in L$,
-        $h(i) = c$.
-
-        \begin{itemize}
-            \item We define a \textit{simple collision} to be any collision
-                with cardinality exactly two.
-
-            \item We define a \textit{strong collision} to be a collision which
-                has multiple blocks which hash to the same output under the same
-                input  state. That is, if $L$ is a collision, then
-                $\forall i \in L$, $i_s = c$ for some $c \in S$ for $L$ to be
-                a strong collision.
-
-            \item We define a \textit{multicollision} to be a collision which
-                has multiple input states which hash to the same output under
-                a single input block. That is, if $L$ is a collision, then
-                $\forall i \in L$, $i_b = c$ for some $c \ in B$ for $L$ to be
-                a multicollision.
-        \end{itemize}
-
-    \item For a simple collision, we say that the \textit{differential path}
-        is the difference between the two sets of intermediate state
-        variables. \textbf{TODO -- HELP define formal notation?}
-            \begin{itemize}
-                \item Signed - Wang's, we don't use.
-                \item Unsigned - more common, actually used by us.
-            \end{itemize}
-
-    \item A \textit{collision class}, $C$, is an arbitrary differential path.
-        If there exists at least one collision pair which has that differential
-        path, we call that collision class non-empty. We denote the set of all
-        non-empty collision classes as $\mathcal{C}$. When we wish to convert
-        between a collision, $c \in L$ and a collision class,
-        $C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$.
-
-    \item A \textit{family of collision classes}, $F$, is the indices of a
-        collision class, $C$, which have any non-zero difference. That is,
-        $F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$.
-\end{itemize}
-
-\subsection{Known Collision Classes}
-
-We define the following set of known collision classes.
-
-\begin{itemize}
-    \item $C_{Wang}$, to be the collision class introduced by X. Wang
-        et al. in \cite{Wang2005}.
-    \item $C_{Sasaki}$, to be the collision class introduced by Y. Sasaki
-        et al. in \cite{Sasaki2007}.
-    \item $C_{Dobbertin}$, to be the collision class introduced by
-        H. Dobbertin in \cite{Dobbertin1998}.
-    \item $C_{Kasselman}$, to be the collision class introduced by
-        P. Kasselman in \cite{KasselmanMD4}.
-    \item $C_{Schlaffer}$, to be the collision class introduced by
-        M. Schl{\"a}ffer in \cite{Schlaffer2006}.
-\end{itemize}
-
-\subsection{Distance Metrics} \label{terminology:distance}
-
-We introduce a distance function, $\delta$, between collision classes by the
-number of differences in intermediate rounds deltas. That is, given two
-collision classes, $C_1, C_2 \in \mathcal{C}$:
-\begin{align}
-    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
-\end{align}
-We extend this distance function to include an \textit{absolute} distance,
-measuring the number of intermediate rounds with non-zero differences.
-We notate this as $\delta(C_1)$.
-
-We can define a similar distance function, $\Delta$, between families of
-collision classes by cardinality of the symmetric difference in the two
-collision families. That is, given two families of collision classes,
-$F_1, F_2 \in \mathcal{F}$:
-\begin{align}
-    \Delta(F_1, F_2) = \big| (F_1 \cup F_2) \setminus (F_1 \cap F_2) \big|
-\end{align}
-This is convenient for when the specifics of the differential path do not
-matter, merely that there exists at least one collision class of the specified
-form.
-
-\subsection{Neighborhoods}
-
-We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
-set of all other non-empty collision classes at a fixed distance,
-$d \in \mathbb{N}$, from $C$. That is:
-\begin{align}
-    N(C, d) & = \{ C_j \in \mathcal{C} : \delta(C, C_j) = d \} \\
-    N(C) & = N(C, 1)
-\end{align}
-For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
-may optionally be omitted, in which case the unit distance is implied. Thus,
-$C_{Kasselman} \in N(C_{Dobbertin})$.
-
-We can similarly define the neighborhood of a family of collision classes,
-$F \in \mathcal{F}$, to be the set of all other families of collision classes
-at a fixed distance, $d \in \mathbb{N}$, from $F$. That is:
-\begin{align}
-    \mathcal{N}(F, d) & = \{ F_j \in \mathcal{F} : \delta(F, F_j) = d \} \\
-    \mathcal{N}(F) & = N(F, 1)
-\end{align}
-The distance parameter, $d$, may optionally be omitted, in which case the unit
-distance is implied.
-
-Neighborhoods can be classified into three types: \textit{expansion},
-\textit{internal}, and \textit{mixed}. Let $d$ be fixed. An \textit{expansion}
-neighborhood of a collision class, $C \in \mathcal{C}$, is the neighborhood
-restricted only to those collision classes which only differ in rounds external
-to the collision family of $C$. An \textit{internal} neighborhood of $C$ is the
-neighborhood restricted only to those collision classes which differ in rounds
-internal to the collision family of $C$. A \textit{mixed} neighborhood of $C$
-is the neighborhood restricted only to collisions which differ in a round
-internal and a round external to the collision family of $C$. That is:
-\begin{align*}
-    N_{exp}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \in F(C), C_j[i] = C[i] \} \\
-    N_{int}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \not\in F(C), C_j[i] = C[i] \} \\
-    N_{mix}(C, d) = \{ & C_{j} \in N(C, d) : \exists i \in F(C), C_j[i] \neq C[i] \\
-        & \text{ and } \exists k \not\in F(C), C_j[k] \neq C[k] \}
-\end{align*}
-
-\section{Intuition on the Utility of a Collision} \label{Sec:Intuition}
-
-\textbf{TODO - audit introductory paragraphs}
-
-We seek to build new intuition about collisions in hash functions
-and particularly collisions in MD4. By developing techniques using logical
-cryptanalysis, we seek to complement and extend the results from
-differential cryptanalysis. In particular, we wish to use logical cryptanalysis
-to discuss the impacts of collisions in real-world systems. In this paper,
-we introduce and give justification for general techniques, and present some
-results of these techniques.
-
-To begin, real world systems are most vulnerable to two attacks: the preimage
-and second preimage. For systems relying on password validation via
-comparing two hashes, preimage attacks would allow inverting the hash,
-increasing the risk associated with losing a database of hashed passwords.
-On the other hand, systems relying on message validation, such as signature
-checks or file integrity checks would be easily broken with fast second
-preimage techniques. Both of these attacks are intractable, thus protecting
-real-world systems. However, much research has been devoted to collision
-attacks, producing efficient results.
-
-Thus, we seek to evaluate the relative difference between a collision
-attack and second preimage attacks, to give some measure of distance between
-them. Several high profile attacks against hash function usage have been
-made. For MD5, these include cloning a CA certificate \cite{Stevens2009}
-and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
-PDFs with the same hash \cite{cryptoeprint:2017:190}. However, many of these
-techniques exploit not the hash, but the flexibility of the file format,
-relying instead on embedding arbitrary, collidable data and later checking for
-the presence of one of the matching blocks. This is especially obvious in the
-MD5 executable and SHA-1 PDF attacks. The work by M. Stevens to perform the
-CA Collision, on the other hand, was a chosen prefix attack extending
-previous work from 2007 \cite{Stevens2007}.
-
-Seeing the need to evaluate how much flexibility exists in a collision,
-we propose the following metrics for evaluating the utility of a collision
-class:
-\begin{enumerate}
-    \item The number of unique differentials a collision class has.
-    \item The number of unit-step neighbors a collision class has.
-    \item The maximum count of zeros in a the binary representation of a
-        colliding block (and likewise with ones).
-    \item Whether there exists a block which collides under multiple initial
-        values.
-    \item Whether or not zero, one, or both of the blocks in a collision may
-        be of ASCII values under any input block difference.
-\end{enumerate}
-Note that the first three are quantitative measures providing some measure of
-flexibility of a collision class, whereas the latter two are merely looking
-for a single witness having such a property. Depending on the scenario,
-specific properties of a collision may be of more interest than others.
-
-The first metric evaluates the flexibility of the differential path. A
-differential path with more flexibility will have more differentials which
-produce blocks with the given differential path. More differentials implies
-a greater flexibility in choice of colliding block, and possibly allows
-for multiple collisions for a given colliding block. Furthermore, a collision
-with more differentials is more likely to satisfy the last metric, having a
-pair of blocks--both ASCII--which produce a collision.
-
-The second metric evaluates the density of the neighborhood of a collision
-class. If a collision resides in a dense neighborhood, it provides more
-possible collision classes to search for a second preimage, chosen prefix, or
-other structure desired in a collision. If, however, a collision class has no
-neighbors, then it cannot be used to find other possible classes for other
-input blocks.
-
-The maximum quantity of zeros (or ones) in the binary representation of a
-block serve as a measure of the extremes to which a collision can be pushed.
-This can additionally be extended to any suitable bit pattern in any base to
-provide a more relevant metric as desired by the system under study.
-
-The fourth metric evaluates the utility of a collision when the internal state
-of the hash function is unknown. If a collision occurs under multiple initial
-values, this could be used to attack some systems where user provided input is
-appended to unknown data and then hashed. If a block collides under a suitably
-large number of initial values, the attack becomes highly likely to occur
-successfully. However, measuring the exact number of initial values a block
-collides under is beyond the scope of this work.
-
-The last metric is similar to the third in that it looks for specific bit
-patterns in a collision. ASCII is one example of a widely used constraint
-system. Further examples, such as JSON, XML, etc., may likewise be
-supplemented based on the specifics of the system.
-
-If a collision class satisfies many of these properties, then it is more
-flexible and thus more likely to be used to target deployed systems. If,
-however, a collision class does not satisfy these properties, its impact is
-likely severely limited in scope, and may not provide useful information to
-find other collision classes which have higher utility.
-
-\section{Techniques for Logical Cryptanalysis} \label{Sec:Logical}
-
 The following techniques have been extensively tested on MD4 and partially
@@
 
+We find justification for the distance metric $\Delta$ in the existing
+literature due to collisions by  X. Wang \cite{cryptoeprint:2004:199} and
+M. Schl{\"a}ffer \cite{Schlaffer2006}. Their respective families are:
+
+\begin{align*}
+F(C_{Wang}) &=& (1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, \textbf{13}, 15, 16, 19, 20, 35, 36)\\
+F(C_{Schlaffer}) &=& (1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, 15, 16, 19, 20, 35, 36)
+\end{align*}
+
 \subsection{Family Similarity}
@@
 Let $F_1$ and $F_2$ be two different families of collisions. Then we say that
-$F_1$ and $F_2$ are \textit{similar}, and notate it $F_1 \lesssim F_2$, if
+$F_1$ and $F_2$ are \textit{similar}, and notate it $F_1 \precsim F_2$, if
 $R(F_1) \leq R(F_2)$ and $F_1 \subseteq F_2$. Note that, when
@@
     \item For every $F \in \mathcal{F}$, there exists $F'$ such that
-        $F' \lesssim F$ and $R(F') + 4 = R(F)$.
+        $F' \precsim F$ and $R(F') + 4 = R(F)$.
     \item For every $F \in \mathcal{F}$, there exists $F'$ such that
-        $F \lesssim F'$.
+        $F \precsim F'$.
 \end{enumerate}
@@
         families within the round can be found by various filling-in
-        techniques through applying the $\lesssim$ relation.
+        techniques through applying the $\precsim$ relation.
 \end{itemize}

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
     \item We discuss techniques for finding useful collisions within a
-        class of collisions using logical cryptanalysis.
+        class using logical cryptanalysis.
     \item We propose additional techniques for building new classes of
-        collisions and show that collisions in MD4 can be found inductively.
+        collisions and show that classes of collisions in MD4 can be found inductively.
     \item We propose five universal metrics for evaluating the utility of
@@
 in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Impacts}, we
-discuss broader impacts of this research.
+discuss the broader impacts and conclusions of this research, and
+conclude with Section \ref{Sec:Conclusion} discussing future work and
+extensions in Section \ref{Sec:Future}.
 
@@
 cryptanalysis. In particular, we look at ways that SAT solving can aid
-cryptanalysis instead of using cryptanalysis to benchmark SAT solvers. We
+cryptanalysis instead of using cryptanalysis to benchmark SAT
+solvers.  A critical assumption of our use of SAT solvers as a means
+to discover structures in hash functions which lead to untrustworthy
+behavior is the insight that while SAT is, in general, a hard problem
+for many specific and practical instances current SAT-solving
+technology is quite efficient \cite{vardi2010p}. We
 then apply them to aid the understanding of attacks against real world
@@
 
-Our work removes the need for finding message modification techniques by
+Our work advances the current state of the art by removing the need for finding message modification techniques by
 specifying the differential path as part of the SAT formula and letting the
@@
 
-However, much of the work in this area is focused on benchmarking SAT solvers,
-and not on evaluating techniques for using SAT solvers to aid reasoning and
+Much of the work in this area is focused on the use of these problems
+to benchmark the performance of SAT solvers,
+and not to evaluate techniques for using SAT solvers to aid reasoning and
 understanding. E. Homsirikamol et al.\ show that SAT solvers can be used to
@@
 
-\subsection{On Trustworthy Computing}
-
-\textbf{TODO - Assigned @erozier}
+\subsection{Trustworthy Computing}
+
+Hash functions see wide usage in modern trustworthy computing
+applications, where they are used to establish the authenticity and
+provenance of data, the credibility of image data
+\cite{friedman1993trustworthy}, sensor identity and trust
+\cite{wong2006dynamic}, meta-data validity \cite{bayer1993improving},
+hardware authorization \cite{sirer2011logical}, and other measures of
+trust in system and mission critical systems.  
+The use of deprecated hashes for commit verification in
+git\cite{loeliger2012version}, and for patch identity and trust
+verification \cite{kuhn2005secure}, or segment verification in
+peer-to-peer decentralized networks \cite{lua2005survey} is quite
+common.  The use of deprecated hashes poses a serious problem, and can
+compromise an otherwise trustworthy system or platform.  A key
+component missing from the existing literature is a good metric which
+can be used to assess the trustworthiness of a hash, and its
+suitability for a given use case.
+
+Such a metric needs to be one which allows for a variable which
+accounts for difficulty, so that trustworthiness can be assessed in a
+meaningful way in an environment with changing hardware capabilities.
 
@@
 
-\section{Discussion of Impacts} \label{Sec:Impacts}
+\section{Discussion} \label{Sec:Impacts}
 
@@
 
-\section{Future Work}
+\section{Future Work} \label{Sec:Future}
 

### 2017-12-05 dsn-2018/bare_conf.tex
Updates to future work, logical cryptanalysis techniques, and improved transition figure
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-\subsection{Distance Metrics} \label{terminology:distance}
+\subsection{Distance Heuristics} \label{terminology:distance}
 
@@
   \begin{subfigure}{\textwidth}
-    \includegraphics[width=\textwidth]{Figs/transitions.png}
+    \includegraphics[width=\textwidth]{Figs/transitions.pdf}
     \caption{Differential Path Transitions}
@@
 
-\subsection{Distance Metrics}
-
-We find justification for the distance metric defined in Section
+\subsection{Distance Heuristics}
+
+We find justification for the distance heuristic, $\delta$, defined in Section
 \ref{terminology:distance} in the existing literature on MD4:
@@
 collision classes have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
-under this metric. The distance function allows us to find collision classes
-based off of existing classes. By limiting ourselves to unit distances, we
-maximally reuse an existing collision and limit the amount of work the SAT
-solver needs to do. Further, rather than searching for any unit-distance
-new collision classes, we can easily split our search to look in each of the
-48 rounds of MD4, allowing us to parallelize the search for neighbors.
+under this metric. Since others have shown that MD4 is untrustworthy, we can
+limit our search space to small distances, exploiting the structure of MD4. This
+reduction in search space decreases the time to solve a single 3-CNF-SAT
+instance at the expense of restricting ourselves to highly connected, highly
+similar differential paths. Further, we can achieve parallel solving
+capabilities by noting that this problem is embarrassingly parallel: each
+candidate collision class can be independently checked on separate machines.
+
+The highly connected, highly similar differential paths are highlighted in
+Figure \ref{Fig:Graph-Clique}: when the neighbors of H. Dobbertin's and
+P. Kasselman's collision classes from Figure \ref{Fig:Graph-Intro} are viewed
+independently, the large cliques become apparent. Depending on the search
+strategies utilized, cliques may be expanded before new neighbors, or a
+significant portion of a vertices neighbors might belong to existing cliques,
+resulting in slow progress towards any external goal.
 
@@
 
-We find justification for the distance metric $\Delta$ in the existing
+We find justification for the distance heuristic $\Delta$ in the existing
 literature due to collisions by  X. Wang \cite{cryptoeprint:2004:199} and
 M. Schl{\"a}ffer \cite{Schlaffer2006}. Their respective families are:
-
-\begin{align*}
-F(C_{Wang}) &=& (1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, \textbf{13}, 15, 16, 19, 20, 35, 36)\\
-F(C_{Schlaffer}) &=& (1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, 15, 16, 19, 20, 35, 36)
-\end{align*}
+$
+\begin{aligned}
+F(C_{Wang}) = (& 1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, \\
+        & \textbf{13}, 15, 16, 19, 20, 35, 36) \\
+\end{aligned} \\
+\begin{aligned}
+F(C_{Schlaffer}) = (& 1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, \\
+         & 15, 16, 19, 20, 35, 36)
+\end{aligned}
+$
 
@@
 
-\textbf{TODO - figures and graphics on dataset}
+\textbf{TODO - figures and graphics on dataset, cross round, etc.}
 
@@
 \section{Conclusion} \label{Sec:Conclusion}
-The conclusion goes here.
+
+\textbf{TODO: COMPLETE}
 
@@
 
-The framework can be seen here \textbf{TODO}. Complete data set available
-upon request.
+Moving forward, we seek to solidify the utility metrics by improving upon
+second preimage and chosen preimage bounds to validate assumptions about the
+utility of collisions. Further, we seek to apply these results to other hash
+functions in the same family, such as MD5, SHA-1, and SHA-2. Lastly, it
+remains to be seen how this work impacts dissimilar constructions such as
+SHA-3 and key derivation functions such as bcrypt and scrypt, and whether
+logical cryptanalysis techniques can be applied to these categories of
+hash functions as well.
+
+This research and all associated code are open access and open source.
+The code is hosted on GitHub at \url{https://github.com/XXXXXXXXXXXXXXXXXX}.
+For additional details or to request data associated with the project, please
+contact the authors.
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
 
-\subsection{Conventions}
-\begin{itemize}
-    \item We assume ordered tuples and strings are indexable via square
-        brackets. Occasionally we use subscripts for referring to
-        elements of tuples with named members.
-    \item We use MD4 as specified in RFC 1320 \cite{rfc1320}.
-\end{itemize}
-
-\textbf{TODO}
-
-\subsection{Collisions in Hash Functions}
-
-We use the following terminology when discussing the notion of collisions in
-a hash function.
-
-\begin{itemize}
-    \item A \textit{cryptographic hash function} is a function:
+In order to define our define our metrics using logical cryptanalysis
+we first define a set of formal terminology and notation whn
+discussing our work, and the notation of collisions within a hash. 
+We make the general assumption that ordered tuples and strings are
+indexable via square brackets. We utilize subscripts when referring to
+elements of tuples with named members, as opposed to just indexable
+members.  When referring to the MD4 hash function, we use MD4 as
+specified in RFC 1320 \cite{rfc1320}.
+
+A \textit{cryptographic hash function} is a function:
 \begin{align*}
@@
 
-    \item An \textit{input} to a hash function is an ordered pair, $i = (s, b)$,
+An \textit{input} to a hash function is an ordered pair, $i = (s, b)$,
 where $s \in S$ and $b \in B$. The \textit{input space} is simply
@@
 
-    \item We use the term \textit{intermediate state variables} to refer to
+We use the term \textit{intermediate state variables} to refer to
 intermediate steps in the computation of a hash function under
@@
 intermediate state variables. We denote the space that $I(i)$ maps
-        into as $\mathcal{V}$. We define the number of rounds as $R$.
-
-    \item We use the notation $R(I)$
+into as $\mathcal{V}$. We define a set of functions $R(x)$ where $x$
+is an element derived from the sets $\mathcal{C}$ or $\mathcal{F}$
+which returns an element.
 

### 2017-12-05 dsn-2018/bare_conf.tex
Terminology
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
         case of MD4, $S$ is all binary strings of length 128, and $B$ is all
-        binary strings of length 512.
+        binary strings of length 512. We assume that the hash function can be
+        reduced to an arbitrary number of rounds, $r$. We note the hash
+        function reduced to $r$ rounds as $h_{r}$.
 
@@
         as an ordered tuple of binary strings of the size of the updated state.
-        In the case of MD4, this is an ordered tuple of cardinality 48, where
-        each index is a binary string of length 32, for a total of 1536
+        In the case of MD4, this is an ordered tuple of cardinality $r$, where
+        each index is a binary string of length 32, for a maximum of $1536$
         intermediate state variables. We denote the space that $I(i)$ maps
-        into as $\mathcal{V}$. We define the number of rounds as $R$.
-
-    \item We use the notation $R(I)$
+        into as $\mathcal{V}$.
 
@@
         there exists an output $v \in S$ such that for all inputs $i \in L$,
-        $h(i) = c$.
+        $h(i) = c$. We extend this to arbitrary round hash functions using
+        the notation $(r, L)$, to explicitly denote that $L$ is a collision
+        under $h_{r}$.
 
         \begin{itemize}
-            \item We define a \textit{simple collision} to be any collision
-                with cardinality exactly two.
-
-            \item We define a \textit{strong collision} to be a collision which
-                has multiple blocks which hash to the same output under the same
-                input  state. That is, if $L$ is a collision, then
-                $\forall i \in L$, $i_s = c$ for some $c \in S$ for $L$ to be
-                a strong collision.
-
             \item We define a \textit{multicollision} to be a collision which
@@
         is the difference between the two sets of intermediate state
-        variables. \textbf{TODO -- HELP define formal notation?}
-            \begin{itemize}
-                \item Signed - Wang's, we don't use.
-                \item Unsigned - more common, actually used by us.
-            \end{itemize}
+        variables. This is formally represented by an ordered tuple of
+        $r$ elements. We use the simple xor difference throughout this paper.
 
@@
         between a collision, $c \in L$ and a collision class,
-        $C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$.
+        $C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$. When
+        $c$ is a tuple, $(r, L)$, we say that the collision class, $C$
+        must exist over same round collisions, and thus $C$ becomes a
+        two-tuple, $(r, d)$, where $d$ is the ordered tuple denoting
+        differential path.
 
@@
         collision class, $C$, which have any non-zero difference. That is,
-        $F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$.
+        $F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$. When $C$ again is over
+        a single round, $r$, we extend $F$ likewise, to become
+        $(r, F)$. We denote the set of all families of collision classes as
+        $\mathcal{F}$.
 \end{itemize}

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-\section{Intuition on the Utility of a Collision} \label{Sec:Intuition}
-
-\textbf{TODO - audit introductory paragraphs}
-
-We seek to build new intuition about collisions in hash functions
+\section{Flexibility and the Utility of a Collision Class} \label{Sec:Intuition}
+
+We seek to build new intuition about collision classes in hash functions
 and particularly collisions in MD4. By developing techniques using logical
-cryptanalysis, we seek to complement and extend the results from
-differential cryptanalysis. In particular, we wish to use logical cryptanalysis
-to discuss the impacts of collisions in real-world systems. In this paper,
-we introduce and give justification for general techniques, and present some
+cryptanalysis, we complement and extend the results from
+differential cryptanalysis. In particular, we use logical cryptanalysis
+to discuss the impacts of collisions in real-world systems, and to
+give justification for general techniques, and the
 results of these techniques.
 
-To begin, real world systems are most vulnerable to two attacks: the preimage
+Real world systems are most vulnerable to two attacks: preimage
 and second preimage. For systems relying on password validation via
@@
 checks or file integrity checks would be easily broken with fast second
-preimage techniques. Both of these attacks are intractable, thus protecting
-real-world systems. However, much research has been devoted to collision
-attacks, producing efficient results.
-
-Thus, we seek to evaluate the relative difference between a collision
-attack and second preimage attacks, to give some measure of distance between
-them. Several high profile attacks against hash function usage have been
-made. For MD5, these include cloning a CA certificate \cite{Stevens2009}
+preimage techniques. Both of these attacks are theoretically intractable, thus protecting
+real-world systems. In practice, however, research efforts have been
+successfully devoted to the production of collision attacks, producing
+efficient results.  These attacks have real consequences for
+trustworthy systems, and evaluating their impact and risk is an
+important and underserved area of the existing literature.
+We seek to evaluate trustworthiness by building a measure of the relative difference between a collision
+attack and second preimage attacks. The literature describes a number
+of successful high profile attacks against popular, and industrially
+relevant hash functions. 
+For MD5, these include cloning a CA certificate \cite{Stevens2009}
 and creating malicious executables \cite{MD5Executables}. For SHA-1, a pair of
-PDFs with the same hash \cite{cryptoeprint:2017:190}. However, many of these
-techniques exploit not the hash, but the flexibility of the file format,
+PDFs with the same hash was successfully produced \cite{cryptoeprint:2017:190}. Many of these
+techniques, however, exploit not the hash but the flexibility of the file format
 relying instead on embedding arbitrary, collidable data and later checking for
@@
 
-Seeing the need to evaluate how much flexibility exists in a collision,
+Evaluating flexibility becomes important as it serves as a
+contrapositive to trustworthiness.  A highly flexible attack implies a
+larger set of families of collision classes, and decreasing the trust
+that should be placed in a hash function. To evaluate how much flexibility exists in a collision,
 we propose the following metrics for evaluating the utility of a collision
@@
 The first metric evaluates the flexibility of the differential path. A
-differential path with more flexibility will have more differentials which
+differential path with more flexibility will have more input-block differentials which
 produce blocks with the given differential path. More differentials implies
@@
 with more differentials is more likely to satisfy the last metric, having a
-pair of blocks--both ASCII--which produce a collision.
+pair of blocks --- both ASCII --- which produce a collision.
 
@@
 of the hash function is unknown. If a collision occurs under multiple initial
-values, this could be used to attack some systems where user provided input is
+values, this could be used to attack some systems where the input
+provided by a user is
 appended to unknown data and then hashed. If a block collides under a suitably
 large number of initial values, the attack becomes highly likely to occur
-successfully. However, measuring the exact number of initial values a block
-collides under is beyond the scope of this work.
-
-The last metric is similar to the third in that it looks for specific bit
+successfully. Measuring the exact number of initial values a block
+collides under, however, is beyond the scope of this work.
+
+The final metric is similar to the third in that it looks for specific bit
 patterns in a collision. ASCII is one example of a widely used constraint
 system. Further examples, such as JSON, XML, etc., may likewise be
-supplemented based on the specifics of the system.
+supplemented based on the specifics of the system, and the domain of
+its application and implementation.
 

### 2017-12-05 dsn-2018/bare_conf.tex

@@
 
+We use the following terminology when discussing the notion of collisions in
+a hash function.
+
 A \textit{cryptographic hash function} is a function:
@@
 case of MD4, $S$ is all binary strings of length 128, and $B$ is all
-binary strings of length 512.
+binary strings of length 512. We assume that the hash function can be
+reduced to an arbitrary number of rounds, $r$. We note the hash
+function reduced to $r$ rounds as $h_{r}$.
 
@@
 as an ordered tuple of binary strings of the size of the updated state.
-In the case of MD4, this is an ordered tuple of cardinality 48, where
-each index is a binary string of length 32, for a total of 1536
+In the case of MD4, this is an ordered tuple of cardinality $r$, where
+each index is a binary string of length 32, for a maximum of $1536$
 intermediate state variables. We denote the space that $I(i)$ maps
-into as $\mathcal{V}$. We define a set of functions $R(x)$ where $x$
-is an element derived from the sets $\mathcal{C}$ or $\mathcal{F}$
-which returns an element.
-
-    \item A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
+into as $\mathcal{V}$.
+
+A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
 input space of a hash function with $\big|L\big| \geq 2$, such that
 there exists an output $v \in S$ such that for all inputs $i \in L$,
-        $h(i) = c$.
+$h(i) = c$. We extend this to arbitrary round hash functions using
+the notation $(r, L)$, to explicitly denote that $L$ is a collision
+under $h_{r}$.
 
 \begin{itemize}
-            \item We define a \textit{simple collision} to be any collision
-                with cardinality exactly two.
-
-            \item We define a \textit{strong collision} to be a collision which
-                has multiple blocks which hash to the same output under the same
-                input  state. That is, if $L$ is a collision, then
-                $\forall i \in L$, $i_s = c$ for some $c \in S$ for $L$ to be
-                a strong collision.
-
 \item We define a \textit{multicollision} to be a collision which
@@
 
-    \item For a simple collision, we say that the \textit{differential path}
+For a simple collision, we say that the \textit{differential path}
 is the difference between the two sets of intermediate state
-        variables. \textbf{TODO -- HELP define formal notation?}
-            \begin{itemize}
-                \item Signed - Wang's, we don't use.
-                \item Unsigned - more common, actually used by us.
-            \end{itemize}
-
-    \item A \textit{collision class}, $C$, is an arbitrary differential path.
+variables. This is formally represented by an ordered tuple of
+$r$ elements. We use the simple xor difference throughout this paper.
+
+A \textit{collision class}, $C$, is an arbitrary differential path.
 If there exists at least one collision pair which has that differential
@@
 between a collision, $c \in L$ and a collision class,
-        $C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$.
-
-    \item A \textit{family of collision classes}, $F$, is the indices of a
+$C \in \mathcal{C}$, we use the notation $C = \mathcal{C}(c)$. When
+$c$ is a tuple, $(r, L)$, we say that the collision class, $C$
+must exist over same round collisions, and thus $C$ becomes a
+two-tuple, $(r, d)$, where $d$ is the ordered tuple denoting
+differential path.
+
+A \textit{family of collision classes}, $F$, is the indices of a
 collision class, $C$, which have any non-zero difference. That is,
-        $F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$.
-\end{itemize}
+$F = \{ 0 \leq i \leq R : C[i] \neq 0 \}$. When $C$ again is over
+a single round, $r$, we extend $F$ likewise, to become
+$(r, F)$. We denote the set of all families of collision classes as
+$\mathcal{F}$.
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
 important and underserved area of the existing literature.
-We seek to evaluate trustworthiness by building a measure of the relative difference between a collision
+
+We evaluate trustworthiness by building a measure of the relative difference between a collision
 attack and second preimage attacks. The literature describes a number
@@
 
-The following techniques have been extensively tested on MD4 and partially
-tested on MD5 and are believed to apply fully to MD5. They likely apply to any
-hash function with similar structure, but as of yet, are untested.
+We approach the problem of logical cryptanalysis by applying
+techniques based in 3-CNF-SAT solving to explore the underlying
+structure of various hash functions.  In this paper we present the
+first case of extensive testing of these methods on the MD4 hash
+function and the partial testing of these methods on the MD5 hash
+function.  Our results lead us to believe that these techniques likely
+fully apply to MD5, as well as any other hash function with similar structure.
 
@@
 
-We find justification for the distance heuristic, $\delta$, defined in Section
-\ref{terminology:distance} in the existing literature on MD4:
+We justify our the distance heuristic, $\delta$, defined in Section
+\ref{terminology:distance} in the existing literature on MD4.  Using
 H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
-collision classes have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
-under this metric. Since others have shown that MD4 is untrustworthy, we can
+collision classes, we find they have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
+under our metric. Since others have shown that MD4 is untrustworthy,
+we hypothesized (correctly) that we could
 limit our search space to small distances, exploiting the structure of MD4. This
@@
 instance at the expense of restricting ourselves to highly connected, highly
-similar differential paths. Further, we can achieve parallel solving
-capabilities by noting that this problem is embarrassingly parallel: each
-candidate collision class can be independently checked on separate machines.
+similar differential paths which none-the-less yield rich and
+interesting sets of collisions. We further improved our performability
+by exploiting the parallalizability of this problem
+noting that this problem is embarrassingly parallel: allowing us to
+check each
+candidate collision class independently on separate machines.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Figs/graph-clique-crop.pdf}
+\caption{Clique and pseudo-clique neighborhoods associated with
+  unit distance neighborhoods of Dobbertin and Kasselman collision
+  classes from the literature.}
+\label{Fig:Graph-Clique}
+\end{center}
+\end{figure}
 
@@
 \subsection{Unit-Step Neighborhood} \label{empirical:neighborhood}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.5\textwidth]{Figs/graph-clique-crop.pdf}
-\caption{}
-\label{Fig:Graph-Clique}
-\end{center}
-\end{figure}
 

### 2017-12-05 dsn-2018/bare_conf.tex
Family similarity
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 A \textit{collision}, $L \subseteq \mathcal{I}$ is a subset of the
-input space of a hash function with $\big|L\big| \geq 2$, such that
+input space of a hash function with $\big|L\big| = 2$, such that
 there exists an output $v \in S$ such that for all inputs $i \in L$,
@@
   a single input block. That is, if $L$ is a collision, then
-  $\forall i \in L$, $i_b = c$ for some $c \ in B$ for $L$ to be
+  $\forall i \in L$, $i_b = c$ for some $c \in B$ for $L$ to be
   a multicollision.
@@
 
-For a simple collision, we say that the \textit{differential path}
+For a collision, we say that the \textit{differential path}
 is the difference between the two sets of intermediate state
@@
 the iterative construction of MD4 and the particular choice of block schedule.
-
-\textbf{TODO -- Empirical justification / graphical representation?}
+We present Figure \ref{Fig:PrecSim} as an example of the lattice structure
+induced under $\precsim$ for example collision families.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{Figs/lattice.pdf}
+\caption{The lattice structure of $\precsim$.}
+\label{Fig:PrecSim}
+\end{center}
+\end{figure*}
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 significant portion of a vertices neighbors might belong to existing cliques,
-resulting in slow progress towards any external goal.
-
-Refer to Table \ref{table:distance} for the distances between existing
-collisions in MD4.
+resulting in slow progress towards any external goal, such as finding
+a second-preimage attack when distance must be greater than some bound.
+
+We present results in Table \ref{table:distance} which illustrate the
+distances found between published collision classes in MD4.
 
@@
 
-We find justification for the distance heuristic $\Delta$ in the existing
+We justify the distance heuristic $\Delta$ in the existing
 literature due to collisions by  X. Wang \cite{cryptoeprint:2004:199} and
@@
 $
+
+For this example $\Delta(F(C_{Wang}), F(C_{Schlaffer})) = 1$.
 

### 2017-12-05 dsn-2018/bare_conf.tex
Expand New Collisions, add better figure
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-\textbf{TODO - figures and graphics on dataset, cross round, etc.}
+Further, compared to the work presented by M. Schl{\"a}ffer and E. Oswald in
+\cite{Schlaffer2006}, this is a 35 fold increase. Increasing this bound is
+possible with low additional cost. Further, by fully considering the
+similarities of collision families under $\precsim$ and better enumerating the
+resulting candidates, additional paths with unique differentials could be
+constructed.
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 We define a relation among families of collisions across rounds as follows.
-Let $F_1$ and $F_2$ be two different families of collisions. Then we say that
-$F_1$ and $F_2$ are \textit{similar}, and notate it $F_1 \precsim F_2$, if
+Let $F_1$ and $F_2$ be two different families of collisions. We say that
+$F_1$ and $F_2$ are \textit{similar} and notate this similarity as $F_1 \precsim F_2$ if
 $R(F_1) \leq R(F_2)$ and $F_1 \subseteq F_2$. Note that, when
@@
     \item For every $F \in \mathcal{F}$, there exists $F'$ such that
-        $F' \precsim F$ and $R(F') + 4 = R(F)$.
+        $F' \precsim F$ and $R(F') + 4 = R(F)$ due to the entire
+        internal state being replaced every four rounds of MD4.
     \item For every $F \in \mathcal{F}$, there exists $F'$ such that
@@
 \end{enumerate}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{Figs/lattice.pdf}
+\caption{The lattice structure of $\precsim$ for the collision
+  families under X. Wang and Y. Sasaki.}
+\label{Fig:PrecSim}
+\end{center}
+\end{figure*}
 
@@
 We present Figure \ref{Fig:PrecSim} as an example of the lattice structure
-induced under $\precsim$ for example collision families.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{Figs/lattice.pdf}
-\caption{The lattice structure of $\precsim$.}
-\label{Fig:PrecSim}
-\end{center}
-\end{figure*}
-
-\subsubsection{Collision Class Similarity}
+induced under $\precsim$ for example collision families, and
+demonstrates the claims from our statements about MD4 for collision
+families under X. Wang and Y. Sasaki.
+
+\paragraph{Collision Class Similarity}
 
@@
 collision classes instead of families of collision classes, but with less
-likelihood. That is, we can transplant an $n$ round collision into an $n+4$
-round and consider a neighborhood around it to find a new collision.
-
-\subsubsection{Second Preimage Similarity}
+likelihood which may lead to the requirement of enlarging the
+neighborhood distance. That is, we can transplant an $n$ round collision into an $n+4$
+round and consider a neighborhood to find new collisions.  Empirically
+we have shown this is not only likely, but extant for published
+collision classes.
+
+\paragraph{Second Preimage Similarity}
 
@@
 $n \leq 24$ rounds, outside of the collision family $\{0, 16\}$ in both 28 and
-32 round MD4--which does not extend any further--we have been unable to find
+32 round MD4 --- which does not extend any further --- we have been unable to find
 any other differential path in those rounds by exhaustive brute force search
@@
 36, 40, 44, and 48 rounds up to a differential path of length 5 with a similar
-lack of results. Thus we believe the general technique to hold, but the size
-of the search space for a second preimage becomes prohibitively large, and
+lack of results. Thus we believe the general technique to hold but it
+is heuristic in nature and the size
+of the search space for a second preimage may become prohibitively
+large so as to be infeasible, and
 further, that differential paths become sufficiently long to prohibit easy
@@
 
-\subsubsection{Chosen Prefix Similarity}
-
-Unlike second preimage attacks, chosen prefix attacks have a higher likelihood
-of working, and give better results. However, we have found that most chosen
+\paragraph{Chosen Prefix Similarity}
+
+The attack also has been shown empirically to work on chosen prefix
+attacks, but once again these attacks are heuristic and not
+guaranteed.  Unlike second preimage attacks, chosen prefix attacks
+have empirically shown a higher likelihood
+of working, and have given better results in practice. We have found that most chosen
 prefixes work in the neighborhoods of existing collisions, and thus are not

### 2017-12-05 dsn-2018/bare_conf.tex

@@
 significant portion of a vertices neighbors might belong to existing cliques,
-resulting in slow progress towards any external goal, such as finding
-a second-preimage attack when distance must be greater than some bound.
-
-We present results in Table \ref{table:distance} which illustrate the
-distances found between published collision classes in MD4.
+resulting in slow progress towards any external goal.
+
+Refer to Table \ref{table:distance} for the distances between existing
+collisions in MD4.
 
@@
 
-We justify the distance heuristic $\Delta$ in the existing
+We find justification for the distance heuristic $\Delta$ in the existing
 literature due to collisions by  X. Wang \cite{cryptoeprint:2004:199} and
@@
 $
-
-For this example $\Delta(F(C_{Wang}), F(C_{Schlaffer})) = 1$.
 

### 2017-12-05 dsn-2018/bare_conf.tex
Added family48 figure
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-\subsection{Unique Differentials}
+\subsection{Unique Differentials} \label{empirical:differentials}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Figs/families48.pdf}
+\caption{A visualization of the unit-distance family relationships
+of the 35,000 new collisions in MD4. }
+\label{Fig:Families48}
+\end{center}
+\end{figure}
+
 We noted in table \ref{table:neighborhood} that the unit-step neighborhood of
@@
 
-\begin{enumerate}
-    \item The space of collision classes is not smooth under the unit distance
+The space of collision classes is not smooth under the unit distance
 relation. That is, starting at $C_{Schlaffer}$ and evaluating
@@
 distance is 29, but after 14 rounds of expansion, we only reduced the
-        distance to 15.
-    \item Even under successive unit neighborhoods, the family of input block
-        differentials remains roughly the same among successive neighborhoods.
+distance to 15. This shows that $\delta$ does not satisfy the triangle
+equality and cannot be a metric, only a heuristic.
+
+Even under successive unit neighborhoods, the family of input block
+differentials remains roughly the same among successive neighborhoods,
+echoing the work of M. Schl{\"a}ffer and E. Oswald \cite{Schlaffer2006}.
 This can be partially counteracted by moving to reduced-round spaces,
@@
 significant input block differential structure with the original
-        collision. With 35,918 unique classes of collisions which had 176
-        unique families of collisions, there were only three unique families
-        of input block differentials: those of the starting collision classes.
-\end{enumerate}
+collision. Our 35,918 collision classes belonged to 176 unique
+families of collisions. However, there were only three unique families
+of input block differentials with 100 input block differences. Note
+however, that the unlike the metric evaluations in
+Section \ref{empirical:differentials}, these collision classes were
+not evaluated for number of unique differentials; thus there are likely
+more than 100 patterns, but still only three families.
 
@@
 of the input block differential, systems can potentially prohibit successful
-collision attacks by preventing structures of input block differences.
-
-Further, compared to the work presented by M. Schl{\"a}ffer and E. Oswald in
+collision attacks by preventing structures of input block differences. Note
+too how the graph presented in \ref{Fig:Families48} shows a distinct grouping
+into three disconnected subgraphs, which fall along input block differential
+family boundaries.
+
+Further, compared to the work presented by M. Schl{\"a}ffer and E. Oswald
 \cite{Schlaffer2006}, this is a 35 fold increase. Increasing this bound is
@@
 resulting candidates, additional paths with unique differentials could be
-constructed.
+constructed. This would allow for new candidate input block differentials,
+better construction of chosen prefix attacks, and potentially to a full second
+preimage attack without relying on message modification techniques.
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
     \item The number of unique differentials a collision class has.
-    \item The number of unit-step neighbors a collision class has.
+    \item The number of unit distance neighbors a collision class has.
     \item The maximum count of zeros in a the binary representation of a
@@
 
-\subsection{Unjustified Performance Observations}
-
-We share the general tips for working with models of MD4 from a SAT solver
-performance perspective, without justification for results. Note that these
-results may be limited in usefulness and may not be uniformly applicable.
+\subsection{Intuition on Performance}
+
+Our empirical results have provided some intuition on the
+performability of our presented attacks, and the general problem of
+working with models of MD4 from a SAT solver
+performance perspective.  While our results are of a heuristic nature,
+and may not be uniformly applicable, we believe they have
+generalizable value and provide insight within the specific domain of
+collision generation.
+
 \begin{itemize}
@@
         families within the round can be found by various filling-in
-        techniques through applying the $\precsim$ relation.
+        techniques through the application of the $\precsim$ relation.
 \end{itemize}
@@
 
-\textbf{TODO -- Introduction?}
+In this section we present our empirical results, evaluating the
+utility of existing published collisions by exploring their structure
+with 3-CNF-SAT, and building representative neighborhoods using our
+distance metrics.  We present these results in the form of an
+exploration of unique differentials, exploration of unit distance
+neighborhoods, maximum count of zeros in the binary representation of
+a colliding block, the existance of blocks which collide under
+multiple initial values, and ASCII block collisions as discussed
+previously in Section \label{Sec:Intuition}.
 
@@
 
-Table \ref{table:differentials} states the number of unique differentials,
-including the specified one, which produces a differential path conforming to
+Table \ref{table:differentials} illustrates the number of unique differentials,
+including the input block differential specified by the original authors, which produces a differential path conforming to
 each of the collision classes. We note in particular that while Y. Sasaki's
@@
 
-\subsection{Unit-Step Neighborhood} \label{empirical:neighborhood}
+\subsection{Unit Distance Neighborhood} \label{empirical:neighborhood}
 
@@
 \includegraphics[width=0.5\textwidth]{Figs/graph-neighborhood-crop.pdf}
-\caption{}
+\caption{All collisions in the unit distance neighborhoods of
+  Dobbertin and Kasselman collision classes
+  organized by connectedness, and showing the inter-distance
+  paths neighborhoods of unpublished collisions.}
 \label{Fig:Graph-Neighborhood}
@@
 
-Refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
+Table \ref{table:neighborhood} presents the sizes discovered empirically of neighborhoods
 of existing collisions. In particular, note that both Y. Sasaki's attack and
@@
 
+Figure \ref{Fig:Graph-Neighborhood} graphically illustrates the
+geometry of the unit distance neighborhoods for H. Dobbertin and
+P. Kasselman's attack, illustrating the geometry of inter-distance
+paths within the newly discovered collisions.  While all nodes
+represent collisions which are unit-distance from the Dobbertin and
+Kasselman collisions, inter-distance paths may be as large as two.
+Clear structure emerges in this representation of the cliques, and
+corresponds directly to the cliques in Figure \ref{Fig:Graph-Clique}.
+
 \subsection{Zeroes}
 
-See table \ref{table:zeroes} for the maximum count of zeroes in a block
+We present results in Table \ref{table:zeroes} which represent the maximum count of zeroes in a block
 for existing collisions. Note that X. Wang's attack is the leader, with 509
@@
 
-We noted in table \ref{table:neighborhood} that the unit-step neighborhood of
+We noted in table \ref{table:neighborhood} that the unit distance neighborhood of
 the existing attacks were non-empty. By repeatedly expanding neighborhoods of

### 2017-12-05 dsn-2018/bare_conf.tex
Conclusion, acknowledgements
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-\section{Discussion} \label{Sec:Impacts}
-
-\textbf{TODO} - Recap impacts on real-world systems like MD5, SHA-1, etc.
-
 % An example of a floating figure using the graphicx package.
@@
 
-\textbf{TODO: COMPLETE}
+- Theory behind Logical Cryptanalysis:
+    - Distance heuristic.
+    - $\precsim$.
+- Tie into Trustworthiness of hash functions.
+- Demonstrated metrics.
+- Demonstrated 35k new collisions.
+- Implications to real world systems.
 
@@
 
-The authors would like to thank the Department of Electrical and Computer
-Engineering and XXXXX XXXXX for providing hardware.
+The authors would like to thank XXXXX XXXXX for providing hardware for the
+project.
 

### 2017-12-05 dsn-2018/bare_conf.tex

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-We noted in table \ref{table:neighborhood} that the unit distance neighborhood of
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Figs/families48.pdf}
+\caption{A visualization of the unit-distance family relationships
+of the 35,000 new collisions in MD4. }
+\label{Fig:Families48}
+\end{center}
+\end{figure}
+
+We noted in table \ref{table:neighborhood} that the unit-step neighborhood of
 the existing attacks were non-empty. By repeatedly expanding neighborhoods of

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
 blocks we've found only collide under the initial state value. Because of this,
-these collisions are only useful if you know the current state where a
+strict multicollision attacks are only useful if you know the current state where a
 colliding pair is to be injected. Thus, this is one area in which collisions
@@
 
-\subsection{Recapitulation}
-
 Overall, we feel that while the given collision classes have value and utility,
@@
 Through the exploration of the neighborhood technique, we have found over
-35,000 new differential paths which originated from the five original
+35,000 new differential paths which originated from the five original published
 collisions in MD4. This shows that the neighborhood definition is sufficient
@@
 of input block differentials with 100 input block differences. Note
-however, that the unlike the metric evaluations in
+however, that unlike the metric evaluations in
 Section \ref{empirical:differentials}, these collision classes were
@@
 Further, compared to the work presented by M. Schl{\"a}ffer and E. Oswald
-\cite{Schlaffer2006}, this is a 35 fold increase. Increasing this bound is
+\cite{Schlaffer2006}, this is a 35 fold increase. Increasing this yield is
 possible with low additional cost. Further, by fully considering the
@@
 constructed. This would allow for new candidate input block differentials,
-better construction of chosen prefix attacks, and potentially to a full second
+better construction of chosen prefix attacks, and could potentially lead to a full second
 preimage attack without relying on message modification techniques.

### 2017-12-05 dsn-2018/bare_conf.tex
Referenced figure of transition states
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-\section{Techniques for Logical Cryptanalysis} \label{Sec:Logical}
-
 \begin{figure*}
@@
     \includegraphics[width=\textwidth]{Figs/differential-1289.png}
-    \caption{Collision 1289}
+    \caption{Collision Class S1289}
     \label{Fig:1289}
@@
     \includegraphics[width=\textwidth]{Figs/differential-1250.png}
-    \caption{Collision 1250}
+    \caption{Collision Class S1250}
     \label{Fig:1250}
@@
     \includegraphics[width=\textwidth]{Figs/transitions.pdf}
-    \caption{Differential Path Transitions}
+    \caption{Differential Path Transitions between S1289 and S1250}
     \label{Fig:Transitions}
   \end{subfigure}\\
+  \caption{An optimal path between collision classes S1289 and S1250,
+  centered around S0 (M. Schl{\"a}ffer's Collision Class), showing each
+  intermediate collision class, transition steps, and appropriate
+  round differences.}
+  \label{Fig:Transition}
 \end{center}
 \end{figure*}
+
+An example of this intuition can be seen in Figure \ref{Fig:Transition}.
+From M. Schl{\"a}ffer's collision class, S1289 and S1250 are equidistant, with
+$\delta(C_{Schlaffer}, S1289) = 3$ and $\delta(S1289, S1250) = 3$.
+Individual step transitions between $S1289$ and $S1250$ are notated in
+Figure \ref{Fig:Transitions}, showing each step and each intermediate collision
+class. Figures \ref{Fig:1289} and \ref{Fig:1250} show the complete differential
+path for S1289 and S1250 respectively.
+
+By considering paths among neighbors, a more optimal exploration strategy can
+be constructed, favoring switching rounds before exploring cliques when the
+triangle inequality holds in the local search area. This allows for a wider
+coverage in differential paths, escaping the local collision type, and
+ultimately improving chosen prefix and second preimage attacks.
+
+\section{Techniques for Logical Cryptanalysis} \label{Sec:Logical}
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 collision attacks by preventing structures of input block differences. Note
-too how the graph presented in \ref{Fig:Families48} shows a distinct grouping
+too how the graph presented in Figure \ref{Fig:Families48} shows a distinct grouping
 into three disconnected subgraphs, which fall along input block differential
@@
 
-- Theory behind Logical Cryptanalysis:
-    - Distance heuristic.
-    - $\precsim$.
-- Tie into Trustworthiness of hash functions.
-- Demonstrated metrics.
-- Demonstrated 35k new collisions.
-- Implications to real world systems.
+We have demonstrated in this paper that our theories for logical
+cryptanalysis hold, and lead to a 35 fold increase in collision
+discovery when compared to the existing literature.  By employing our
+presented distance heuristic and the lattice structure uncovered by
+the $\precsim$ relation, we can evaluate the trustworthiness of a hash
+function as the inverse of the flexibility of known collisions for
+that hash function.  This allows us to measure the structural
+weaknesses of a hash function and determine if extant collisions are sufficient
+to render a hash function untrustworthy, and likewise to understand if
+necessary conditions related to the underlying structure are met to
+ensure trustworthiness.
+
+We demonstrate the utility of five metrics of the flexibility present
+in a collision on the basis of 1) the number of unique differentials
+within a collision class; 2) the size of the unit distance
+neighborhood for a collision class; 3) the maximum count of zeros in a
+binary representation of a colliding block; 4) the existance of blocks
+which collide under multiple initial values; and 5) the presence of
+ASCII valued inplut blocks within a collision.  These metrics proved
+capable when evaluating the flexibility of collisions in the published
+literature, and showed that flexibility is a multi-dimensional
+quantity that must be compared over multiple distance metrics, as not all
+collisions dominate universally.
+
+Our empirical results furthermore demonstrated that concerns over
+industrial use of deprecated hash functions, specifically MD4, are
+real, and that structural weaknesses exist which are highly
+exploitable and limit the trustworthiness of any system which employs
+such hash functions.  We hope the presented results lead to a
+reduction in the use of such deprecated hash functions, and that new
+hash functions are periodically evaluated when collisions are
+discovered to understand if these collisions also indicate the
+presence of structural weaknesses.
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
 % have to be run twice to clear warning/error messages.
+
+\linespread{1.05}
 
@@
 
-Moving forward, we seek to solidify the utility metrics by improving upon
+While the literature has established a set of known collisions and
+some initial algorithms for finding new collisions, the 35 fold
+increase in collisions discovered by the methods we have presented in
+this paper open a broad range of possible avenues for future work.
+Moving forward we plan to solidify the utility metrics by improving upon
 second preimage and chosen preimage bounds to validate assumptions about the
-utility of collisions. Further, we seek to apply these results to other hash
-functions in the same family, such as MD5, SHA-1, and SHA-2. Lastly, it
-remains to be seen how this work impacts dissimilar constructions such as
+utility of collisions. Further, we plan to apply these results to other hash
+functions in the same family, such as MD5, SHA-1, and SHA-2 to better
+understand if published collisions represent structural weaknesses
+which impact the trustworthiness of these broadly used algorithms, or not. Lastly, it
+remains to be seen how this work impacts dissimilar hash function constructions such as
 SHA-3 and key derivation functions such as bcrypt and scrypt, and whether
@@
 
-This research and all associated code are open access and open source.
-The code is hosted on GitHub at \url{https://github.com/XXXXXXXXXXXXXXXXXX}.
+In order to help the community build upon or work, and for the
+purposes of reproducability and ascertainment of correctness the
+research, data, and all associated code are provided on an open access
+and open source basis under the GNU General Public License Version 3
+\footnote{\url{https://www.gnu.org/licenses/gpl-3.0.en.html}}.
+The code itself is hosted on GitHub at \url{https://github.com/XXXXXXXXXXXXXXXXXX}.
 For additional details or to request data associated with the project, please
@@
 
-The authors would like to thank XXXXX XXXXX for providing hardware for the
+The authors would like to thank XXXXX XXXXX for providing access to
+the hardware used for the
 project.

### 2017-12-05 dsn-2018/bare_conf.tex
Minor changes
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 Cryptographic hash functions form the core of many protocols which
-trust that cryptographic hash functions preserve certain basic
+trust that these functions preserve certain basic
 properties. Cryptograpic hash functions see widespread
-use for file integrity checks to verify long term storage of
+use as file integrity checks to verify long term storage of
 data, as cache invalidation techniques, and as a building block in network
@@
 from Dobbertin and Kasselman is presented in Figure \ref{Fig:Graph-Intro}.
+
 This paper contributes several techniques to logical cryptanalysis and uses these
@@
 differential paths and the effectiveness of various techniques presented
-in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Impacts}, we
-discuss the broader impacts and conclusions of this research, and
-conclude with Section \ref{Sec:Conclusion} discussing future work and
+in Section \ref{Sec:Logical}. Lastly, we discuss the broader impacts and
+conclusions of this research in Section \ref{Sec:Conclusion}, discussing future work and
 extensions in Section \ref{Sec:Future}.
@@
 input block differentials and thus less flexibility in this aspect. Further,
-M. Schl{\"a}ffer's attack -- also building off the work of X. Wang -- has the
+M. Schl{\"a}ffer's attack --- also building off the work of X. Wang --- has the
 same number of differentials and is thus also more flexible than Y. Sasaki's.
@@
 of existing collisions. In particular, note that both Y. Sasaki's attack and
-M. Schl{\"a}ffer's attack -- which both claimed to improve upon X. Wang's
-attack--have larger neighborhood sizes, and likewise with P. Kasselman's
+M. Schl{\"a}ffer's attack --- which both claimed to improve upon X. Wang's
+attack---have larger neighborhood sizes, and likewise with P. Kasselman's
 attack which improved upon H. Dobbertin's attack. Thus, the newer collisions
@@
 collision attacks by preventing structures of input block differences. Note
-too how the graph presented in Figure \ref{Fig:Families48} shows a distinct grouping
+too how the graph presented in \ref{Fig:Families48} shows a distinct grouping
 into three disconnected subgraphs, which fall along input block differential
@@
 
-We have demonstrated in this paper that our theories for logical
-cryptanalysis hold, and lead to a 35 fold increase in collision
-discovery when compared to the existing literature.  By employing our
-presented distance heuristic and the lattice structure uncovered by
-the $\precsim$ relation, we can evaluate the trustworthiness of a hash
-function as the inverse of the flexibility of known collisions for
-that hash function.  This allows us to measure the structural
-weaknesses of a hash function and determine if extant collisions are sufficient
-to render a hash function untrustworthy, and likewise to understand if
-necessary conditions related to the underlying structure are met to
-ensure trustworthiness.
-
-We demonstrate the utility of five metrics of the flexibility present
-in a collision on the basis of 1) the number of unique differentials
-within a collision class; 2) the size of the unit distance
-neighborhood for a collision class; 3) the maximum count of zeros in a
-binary representation of a colliding block; 4) the existance of blocks
-which collide under multiple initial values; and 5) the presence of
-ASCII valued inplut blocks within a collision.  These metrics proved
-capable when evaluating the flexibility of collisions in the published
-literature, and showed that flexibility is a multi-dimensional
-quantity that must be compared over multiple distance metrics, as not all
-collisions dominate universally.
-
-Our empirical results furthermore demonstrated that concerns over
-industrial use of deprecated hash functions, specifically MD4, are
-real, and that structural weaknesses exist which are highly
-exploitable and limit the trustworthiness of any system which employs
-such hash functions.  We hope the presented results lead to a
-reduction in the use of such deprecated hash functions, and that new
-hash functions are periodically evaluated when collisions are
-discovered to understand if these collisions also indicate the
-presence of structural weaknesses.
+- Theory behind Logical Cryptanalysis:
+    - Distance heuristic.
+    - $\precsim$.
+- Tie into Trustworthiness of hash functions.
+- Demonstrated metrics.
+- Demonstrated 35k new collisions.
+- Implications to real world systems.
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-\linespread{1.05}
+\linespread{1.04}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{Figs/graph-intro-crop.pdf}
+\caption{A visualization of the heuristic unit-distance neighborhood
+  of valid collisions from the Dobbertin and Kasselman collisions.}
+\label{Fig:Graph-Intro}
+\end{center}
+\end{figure}
+
 The most important aspect of these can be formally defined under three
@@
 collisions can be found.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.5\textwidth]{Figs/graph-intro-crop.pdf}
-\caption{A visualization of the heuristic unit-distance neighborhood
-  of valid collisions from the Dobbertin and Kasselman collisions.}
-\label{Fig:Graph-Intro}
-\end{center}
-\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-\includegraphics[width=0.5\textwidth]{Figs/graph-clique-crop.pdf}
+\includegraphics[width=0.7\textwidth]{Figs/graph-clique-crop.pdf}
 \caption{Clique and pseudo-clique neighborhoods associated with
@@
 \end{center}
-\end{figure}
+\end{figure*}
 
@@
 multiple initial values, and ASCII block collisions as discussed
-previously in Section \label{Sec:Intuition}.
+previously in Section \ref{Sec:Intuition}.
 
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-\includegraphics[width=0.5\textwidth]{Figs/graph-neighborhood-crop.pdf}
+\includegraphics[width=0.7\textwidth]{Figs/graph-neighborhood-crop.pdf}
 \caption{All collisions in the unit distance neighborhoods of
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-\includegraphics[width=0.5\textwidth]{Figs/families48.pdf}
+\includegraphics[width=0.7\textwidth]{Figs/families48.pdf}
 \caption{A visualization of the unit-distance family relationships
-of the 35,000 new collisions in MD4. }
+of the 35,000 new collisions in MD4. Note these nodes represent
+collision \textbf{families} with solid vertices as opposed to the
+classes in previous figures represented by open verticies.}
 \label{Fig:Families48}
 \end{center}
-\end{figure}
+\end{figure*}
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
 % affiliations
-\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
-\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
-Ames, Iowa, USA 50011\\
-Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}
+
+%\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
+%\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
+%Ames, Iowa, USA 50011\\
+%Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}
+
+\vspace{3in}
+
 % conference papers do not typically use \thanks and this command
@@
 Cryptographic hash functions form the core of many protocols which
-trust that cryptographic hash functions preserve certain basic
-properties. Cryptograpic hash functions see widespread
+trust that these functions preserve certain basic
+properties. Cryptographic hash functions see widespread
 use for file integrity checks to verify long term storage of
@@
 computing a hash for a message should not be algorithmically
-difficult; non-reversability - it should be infeasible to generate a message
+difficult; non-reversibility - it should be infeasible to generate a message
 from its hash except by trying all possible messages; sensitivity to
@@
 cryptanalysis. In particular, we look at ways that SAT solving can aid
-cryptanalysis instead of using cryptanalysis to benchmark SAT
+cryptanalysis directly, where as in the literature cryptanalysis is
+primarily used to benchmark SAT
 solvers.  A critical assumption of our use of SAT solvers as a means

### 2017-12-05 dsn-2018/bare_conf.tex
Change definition
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 from its hash except by trying all possible messages; sensitivity to
-initial conditions - small changes to a message should change the hash
-value to a degree that the new hash value appears uncorrelated with
-the old hash value; unique digests - it should be infeasible to find
-two different messages with the same hash value.
+initial conditions - small changes to a message should change every bit of
+the hash value with probability $\frac{1}{2}$;
+unique digests - it should be infeasible to find two different messages with
+the same hash value.
 

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

@@
 properties. Cryptographic hash functions see widespread
-use for file integrity checks to verify long term storage of
+use as file integrity checks to verify long term storage of
 data, as cache invalidation techniques, and as a building block in network
@@
 for these use cases necessarily have
-strong guarantees about properties of its members.
+strong guarantees about the properties of its members.
 An ideal cryptographic hash has five main features: determinism - the
@@
 We have demonstrated in this paper that our theories for logical
-cryptanalysis hold, and lead to a 35 fold increase in collision
+cryptanalysis hold: namely the existance of useful distance
+heuristics; the existance of a similarity operator, $\precsim$,
+between families of collision classes; and the existance of a useful
+heuristic structure for second preimage and chosen prefix attacks.
+These theories lead to techniques which result in a 35 fold increase in collision
 discovery when compared to the existing literature.  By employing our

### 2017-12-05 dsn-2018/bare_conf.tex
Added comma
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 In order to help the community build upon or work, and for the
-purposes of reproducability and ascertainment of correctness the
+purposes of reproducability and ascertainment of correctness, the
 research, data, and all associated code are provided on an open access

### 2018-02-02 honors-thesis/usenix.tex

@@
 \section{Introduction}
-\subsection{History of SAT in Cryptography}
+\subsection{Brief History of SAT in Cryptography}
+
+Much of the original research for developing the use of SAT with cryptography
+was done by Fabio Massacci currently at the University of Trento \cite{CryptoSAT}.
+The focus of this work was largely two fold: cryptoanalysis against the DES
+and faking signatures in the RSA with $e=3$. More information can be found
+here \cite{MASSACCI2000LogicalCA} and here \cite{FIORINI2003101}, but note that
+this work predates all of the collisions discussed in this article.
 
@@
 other examples under the same differential and weak state transition constraints.
-Note that an extensible collision is inherently constraint-based.
+Note that if a collision is extensible, it is also constraint-based.
 
 A collision is \textbf{differential-independent} if there are multiple
-differentials which satisfy the intermediate state constr
+differentials which satisfy the state transition constraints.
+
 A collision is \textbf{malleable} if collisions can be found with different
@@
 starting states which produce collisions. A state-independent collision
-is termed strong if one collision can collide under two or more starting
+is termed \textbf{strong} if one block can collide under two or more starting
 states.
@@
 We now introduce a series of notations to ease construction of cryptographic
-SAT models. $b := AND_{i = 0}^{32}(a)$. Needs to be
+SAT models.
+
+We define
+\begin{align*}
+    c_{l}^{l + k - j} := EQ_{i = j}^{k}(a_i, b_i)
+\end{align*}
+for $j \geq 0, k \geq j, l \geq 0$ to be equivalent to:
+\begin{align*}
+    \{ c_l & := a_j == b_j, \\
+    c_{l+1} & := a_{j+1} == b_{j+1}, \\
+    ..., \\
+    c_{l+k-j} & := a_{k} == b_{k} \} \\
+\end{align*}
+In particular, $c$ is a tuple of boolean equations with cardinality $|c| = k-j$,
+defining an equality relation between input sets $a$ and $b$, at index starting
+at $j$ and ending at $k$ (inclusive).
 
@@
 \subsubsection{State-Independent}
+h1: f7f3fe7437e33bee3590392936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d31498806e520608c6738da70170b:40a8b995bd89b97662b4516e1df317f2
+h2: f7f3fe7437e33b6e3590399936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d30498806e520608c6738da70170b:40a8b995bd89b97662b4516e1df317f2
+h3: f7f3fe7437e33bee3590392936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d31498806e520608c6738da70170b:aabcc2e21ea0e529178ca7dcb21240a2
+h4: f7f3fe7437e33b6e3590399936aeab37e0b848329173bf0f08a252a0281e3284f4e4fd90626c4cbb3350ac3e36694d34703d30498806e520608c6738da70170b:aabcc2e21ea0e529178ca7dcb21240a2
+c Total time               : 127.12
+
 \subsubsection{Malleable}
@@
 
+\subsubsection{State Independent}
+h1: 13d340f35ff2c0ace19bcd73f5c153aa816105701b5050a2c7e0ab260adf14026e0a20b1f39a8f4be37e60aad37d5b1100140082c228676c4d1c8098c095d023:746094a6c1cf11f96614c893f6b54d67
+h2: 13d340035ff2c0ace19bcdf3f5c153aa816105f01b5050a2c7e0ab260adf14026e0a2031f39a8f4be37e60aad37d5b1100140002c228676c4d1c8098c095d023:746094a6c1cf11f96614c893f6b54d67
+h3: 13d340f35ff2c0ace19bcd73f5c153aa816105701b5050a2c7e0ab260adf14026e0a20b1f39a8f4be37e60aad37d5b1100140082c228676c4d1c8098c095d023:ef64866bf9eedb5a723d7153bcce781f
+h4: 13d340035ff2c0ace19bcdf3f5c153aa816105f01b5050a2c7e0ab260adf14026e0a2031f39a8f4be37e60aad37d5b1100140002c228676c4d1c8098c095d023:ef64866bf9eedb5a723d7153bcce781f
+c Total time               : 165.32
+
 \section{MD5}
@@
 \subsection{Xie's MD5 Collisions}
+\subsection{Sasaki's MD5 Collisions}
 

### 2018-02-15 dsn-2018/rebuttal-response.txt
Initial rebuttal
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

Thank you to all of the reviewers. We have fixed the grammatical and
typographical errors in the paper and will work on better organizing the
paper to highlight our reviewers' comments.

This paper has two major parts: 1) suggesting metrics for measuring the
trustworthiness of a hash function and 2) developing new logical
cryptanalysis (LC) techniques for analyzing the structure of hash functions.
We selected MD4 not to show that it was untrustworthy, but because it was
widely agreed to be so. In the interests of reproducibility, we apply our
metrics to a known-untrustworthy function, which has multiple pre-existing
attacks against it. Further, our new LC techniques are more broadly applicable
to the entire class of Merkel-Damgård; while we have preliminary results on
MD5 and SHA-1, these are immature and not in the scope of this paper. However,
while the utility metrics apply to SHA-3 (and broadly to most standardized
hash functions), the new LC techniques will not immediately transfer. We will
update the submission to make these distinctions more clear.

To Reviewer #285E, we argue that these are five baseline metrics and not
necessarily meant to be comprehensive for all use cases. Traditionally,
the discovery of collisions in a hash function has resulted in its assumed
untrustworthiness for all use cases of that function. Metrics 1 and 2 are
independent of the target system and try to categorize how much additional
information is gained from finding a single collision (e.g., if a collision
class has many neighbors, this implies that learning about that one class
gives you multiple other candidate classes, and thus that the system is highly
structured and unsuitable for use as a cryptographically secure hash function).
Metric 3 is suitable for systems where the attacker has limited control and
most closely needs a second-preimage attack (e.g., attacking the PKI via
colliding into an existing root certificate). Metric 4 is suitable for systems
where the attacker has full control over the contents of the block, but limited
control over where that block occurs in context to other blocks. Metric 5
is most suitable to systems like source code repositories, JSON API endpoints,
or shell scripts, where only a limited character set is normal. For example,
to use a collision in SHA-1 to attack Git, metrics 4 and 5 are important, and
3 is only important if the target code needs to make sense; metrics 1 and 2
are not necessary, but useful if the proposed collision fails to achieve the
desired effect.

We are leaving aside two main features: weightings of these metrics and any
applicability of these metrics to preimage attacks, as they tend to be rare
and thus hard to study from an empirical perspective. However, we leave aside
weightings and an overarching measure of "how far is a collision from a second
preimage" for the reason that it isn't strictly necessary for every system,
and no one weighting can completely encompass all use cases for a hash
function. Thus, part of our desire to publish is to begin a dialogue with the
broader community regarding what other metrics are important, and what their
effect on the general trustworthiness of a hash is.

To Reviewer #285D, we argue that, while DSN doesn't explicitly call for
cryptographic research, our work falls into the area of
"models...for...assessing dependable systems". Because cryptographic hash
functions are well specificied and determanistic, we can model the impact of a
collision attack on a system in a SAT framework. In addition to the five
general metrics discussed above, explicitly evaluating the trustworthiness
for a specific use case is a novel advancement and well within the scope of
DSN. We present the new LC techniques, in part, to better demonstrate how we
arrived at a ~10-fold increase in collision classes and also as a basis for
a larger empirical study of the effectiveness of these metrics. Further,
part of our desire to publish in DSN is to begin a dialogue with the
broader community around dependable systems regarding what other metrics
are important to their specific use cases.

Thank you for your time.

### 2018-02-15 dsn-2018/rebuttal-response.txt
Spelling
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 "models...for...assessing dependable systems". Because cryptographic hash
-functions are well specificied and determanistic, we can model the impact of a
+functions are well specified and deterministic, we can model the impact of a
 collision attack on a system in a SAT framework. In addition to the five

### 2018-02-15 dsn-2018/rebuttal-response.txt
Updates to rebuttal
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-This paper has two major parts: 1) suggesting metrics for measuring the
-trustworthiness of a hash function and 2) developing new logical
-cryptanalysis (LC) techniques for analyzing the structure of hash functions.
-We selected MD4 not to show that it was untrustworthy, but because it was
-widely agreed to be so. In the interests of reproducibility, we apply our
-metrics to a known-untrustworthy function, which has multiple pre-existing
-attacks against it. Further, our new LC techniques are more broadly applicable
-to the entire class of Merkel-Damgård; while we have preliminary results on
-MD5 and SHA-1, these are immature and not in the scope of this paper. However,
-while the utility metrics apply to SHA-3 (and broadly to most standardized
-hash functions), the new LC techniques will not immediately transfer. We will
+This paper has two major parts:
+    1) suggesting metrics for measuring the trustworthiness of a hash function
+    2) developing new logical cryptanalysis techniques (LCT) for analyzing
+        the structure of hash functions.
+We recognize that MD4 is untrustworthy and use it for the interests of
+reproducibility. We sought to test our new metrics against a
+known-untrustworthy hash which has multiple pre-existing attacks against it.
+Further, our new LC techniques are more broadly applicable to the entire class
+of Merkel-Damgård; while we have preliminary results on MD5 and SHA-1, these
+results are immature and not in the scope of this paper. However, while the
+utility metrics apply to SHA-3 (and broadly to most standardized hash
+functions), the new LC techniques will not immediately transfer. We will
 update the submission to make these distinctions more clear.
@@
 the discovery of collisions in a hash function has resulted in its assumed
-untrustworthiness for all use cases of that function. Metrics 1 and 2 are
-independent of the target system and try to categorize how much additional
-information is gained from finding a single collision (e.g., if a collision
-class has many neighbors, this implies that learning about that one class
-gives you multiple other candidate classes, and thus that the system is highly
-structured and unsuitable for use as a cryptographically secure hash function).
-Metric 3 is suitable for systems where the attacker has limited control and
-most closely needs a second-preimage attack (e.g., attacking the PKI via
-colliding into an existing root certificate). Metric 4 is suitable for systems
-where the attacker has full control over the contents of the block, but limited
-control over where that block occurs in context to other blocks. Metric 5
-is most suitable to systems like source code repositories, JSON API endpoints,
-or shell scripts, where only a limited character set is normal. For example,
-to use a collision in SHA-1 to attack Git, metrics 4 and 5 are important, and
-3 is only important if the target code needs to make sense; metrics 1 and 2
-are not necessary, but useful if the proposed collision fails to achieve the
-desired effect.
+untrustworthiness for all use cases of that function.
+
+    - Metrics 1 and 2 are independent of the target system and try to
+        quantify the information is gained from finding a single collision.
+        (E.g., if a collision has a large number of neighbors, it implies
+        structure in the hash function, making it untrustworthy)
+    - Metric 3 is suitable for systems where the attacker has limited control
+        and needs a second-preimage attack. (E.g., attacking the PKI via
+        colliding into an existing root certificate).
+    - Metric 4 is suitable for systems where the attacker has full control
+        over the contents of the block, but limited control over where that
+        block occurs in context to other blocks.
+    - Metric 5 is most suitable to systems like source code repositories or
+        shell scripts, where only a limited character set is normal.
+For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are
+important, and 3 is only important if the target code needs to make sense;
+metrics 1 and 2 are not necessary, but useful if the given collision fails.
 

### 2018-02-15 dsn-2018/rebuttal-response.txt
Updates to rebuttal
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 Thank you to all of the reviewers. We have fixed the grammatical and
-typographical errors in the paper and will work on better organizing the
-paper to highlight our reviewers' comments.
+typographical errors in the paper and will work on better organizing
+the paper to highlight our reviewers' comments.
 
@@
         the structure of hash functions.
-We recognize that MD4 is untrustworthy and use it for the interests of
+We recognize that MD4 is untrustworthy and use it in the interest of
 reproducibility. We sought to test our new metrics against a
 known-untrustworthy hash which has multiple pre-existing attacks against it.
-Further, our new LC techniques are more broadly applicable to the entire class
-of Merkel-Damgård; while we have preliminary results on MD5 and SHA-1, these
-results are immature and not in the scope of this paper. However, while the
-utility metrics apply to SHA-3 (and broadly to most standardized hash
-functions), the new LC techniques will not immediately transfer. We will
-update the submission to make these distinctions more clear.
+Furthermore, our new LCTs are applicable to the entire class of Merkel-Damgård,
+but while we have preliminary results on MD5 and SHA-1, they are immature and
+not in the scope of this paper. The utility metrics apply to SHA-3 but the new
+LCTs will not immediately transfer. We will update the submission to make
+these distinctions more clear.
 
@@
 the discovery of collisions in a hash function has resulted in its assumed
-untrustworthiness for all use cases of that function.
-
+untrustworthiness for all use cases of that function; our proposed metrics
+seek to address this shortcoming in the literature.
     - Metrics 1 and 2 are independent of the target system and try to
@@
 For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are
-important, and 3 is only important if the target code needs to make sense;
-metrics 1 and 2 are not necessary, but useful if the given collision fails.
+important, and 3 is relevant when the target code needs to make sense; 1 and
+2 are not necessary, but useful if the given collision fails. Thus, the only
+aspect we leave aside is how to weight these individual metrics to measure
+the distance from a collision attack to a second preimage attack, but we
+feel that this weighting is specific to individual use cases.
 
-We are leaving aside two main features: weightings of these metrics and any
-applicability of these metrics to preimage attacks, as they tend to be rare
-and thus hard to study from an empirical perspective. However, we leave aside
-weightings and an overarching measure of "how far is a collision from a second
-preimage" for the reason that it isn't strictly necessary for every system,
-and no one weighting can completely encompass all use cases for a hash
-function. Thus, part of our desire to publish is to begin a dialogue with the
-broader community regarding what other metrics are important, and what their
-effect on the general trustworthiness of a hash is.
-
-To Reviewer #285D, we argue that, while DSN doesn't explicitly call for
-cryptographic research, our work falls into the area of
-"models...for...assessing dependable systems". Because cryptographic hash
-functions are well specified and deterministic, we can model the impact of a
-collision attack on a system in a SAT framework. In addition to the five
-general metrics discussed above, explicitly evaluating the trustworthiness
-for a specific use case is a novel advancement and well within the scope of
-DSN. We present the new LC techniques, in part, to better demonstrate how we
-arrived at a ~10-fold increase in collision classes and also as a basis for
-a larger empirical study of the effectiveness of these metrics. Further,
-part of our desire to publish in DSN is to begin a dialogue with the
-broader community around dependable systems regarding what other metrics
-are important to their specific use cases.
+To Reviewer #285D, we argue that our work falls into DSN’s call for papers
+which describe "models...for...assessing dependable systems". Because
+cryptographic hash functions are well specified and deterministic, we can model
+the impact of a collision attack on a system in a SAT framework. In addition to
+the five general metrics discussed above, explicitly evaluating the
+trustworthiness for a specific use case is a novel advancement and well within
+the scope of DSN. We present the new LCTs as basis for an empirical study of
+the effectiveness of these metrics and techniques. Further, part of our desire
+to publish in DSN is to begin a dialogue with the broader community around
+dependable systems regarding what other metrics are important to their specific
+use cases.
 

### 2018-02-16 dsn-2018/rebuttal-response.txt
Updates from @ewdavi
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
-Thank you to all of the reviewers. We have fixed the grammatical and
-typographical errors in the paper and will work on better organizing
-the paper to highlight our reviewers' comments.
+Thank you to all of the reviewers. Grammatical/typographical errors have been fixed. We will improve the organization as suggested.
 
-This paper has two major parts:
-    1) suggesting metrics for measuring the trustworthiness of a hash function
-    2) developing new logical cryptanalysis techniques (LCT) for analyzing
-        the structure of hash functions.
-We recognize that MD4 is untrustworthy and use it in the interest of
-reproducibility. We sought to test our new metrics against a
-known-untrustworthy hash which has multiple pre-existing attacks against it.
-Furthermore, our new LCTs are applicable to the entire class of Merkel-Damgård,
-but while we have preliminary results on MD5 and SHA-1, they are immature and
-not in the scope of this paper. The utility metrics apply to SHA-3 but the new
-LCTs will not immediately transfer. We will update the submission to make
-these distinctions more clear.
+We make two main contributions:
 
-To Reviewer #285E, we argue that these are five baseline metrics and not
-necessarily meant to be comprehensive for all use cases. Traditionally,
-the discovery of collisions in a hash function has resulted in its assumed
-untrustworthiness for all use cases of that function; our proposed metrics
-seek to address this shortcoming in the literature.
-    - Metrics 1 and 2 are independent of the target system and try to
-        quantify the information is gained from finding a single collision.
-        (E.g., if a collision has a large number of neighbors, it implies
-        structure in the hash function, making it untrustworthy)
-    - Metric 3 is suitable for systems where the attacker has limited control
-        and needs a second-preimage attack. (E.g., attacking the PKI via
-        colliding into an existing root certificate).
-    - Metric 4 is suitable for systems where the attacker has full control
-        over the contents of the block, but limited control over where that
-        block occurs in context to other blocks.
-    - Metric 5 is most suitable to systems like source code repositories or
-        shell scripts, where only a limited character set is normal.
-For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are
-important, and 3 is relevant when the target code needs to make sense; 1 and
-2 are not necessary, but useful if the given collision fails. Thus, the only
-aspect we leave aside is how to weight these individual metrics to measure
-the distance from a collision attack to a second preimage attack, but we
-feel that this weighting is specific to individual use cases.
+1) metrics for measuring the trustworthiness of a hash.
+2) new logical cryptanalysis techniques (LCT) for analyzing the structure of hash functions.
 
-To Reviewer #285D, we argue that our work falls into DSN’s call for papers
-which describe "models...for...assessing dependable systems". Because
-cryptographic hash functions are well specified and deterministic, we can model
-the impact of a collision attack on a system in a SAT framework. In addition to
-the five general metrics discussed above, explicitly evaluating the
-trustworthiness for a specific use case is a novel advancement and well within
-the scope of DSN. We present the new LCTs as basis for an empirical study of
-the effectiveness of these metrics and techniques. Further, part of our desire
-to publish in DSN is to begin a dialogue with the broader community around
-dependable systems regarding what other metrics are important to their specific
-use cases.
+We recognize MD4 as untrustworthy, but we use it for reproducibility.
+We tested our novel metrics against a known-untrustworthy hash with multiple pre-existing attacks to measure impact and improvement of our methods.
+Our new LCTs are applicable to the entire class of Merkel-Damgard hash functions.
+We have preliminary results on MD5 and SHA-1, full treatment exceeds the scope and page limit of this paper.
+Utility metrics apply to SHA-3, but LCTs do not trivially transfer.
+We will update the submission to make these distinctions clear.
 
-Thank you for your time.
+#285E, we present five baseline metrics, but do not say they are comprehensive.
+Traditionally, the discovery of collisions in a hash function has resulted in its assumed untrustworthiness for all use cases of that function; our proposed metrics seek to address this shortcoming.
+
+- Metrics 1 and 2 are independent of the target system and try to quantify the information gained from finding a single collision.
+The neighborhood density implies structure in the hash function, providing a formal measure of untrustworthiness.
+
+- Metric 3 is suitable for systems where the attacker has limited control and needs a second-preimage attack. (E.g., attacking the PKI via colliding into a certificate).
+
+- Metric 4 is suitable for systems where the attacker has full control over the contents of the block, but limited control over where that block occurs in context to other blocks.
+
+- Metric 5 is most suitable to systems like source code repositories or shell scripts, where only a limited character set is normal.
+
+For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are important, and 3 is relevant when the target code needs to make sense; 1 and 2 are not necessary, but useful if the given collision fails.
+Thus, the only aspect we leave aside is how to weight these individual metrics to measure the distance from a collision attack to a second preimage attack, but we feel that this weighting is specific to individual use cases.
+
+#285D, we argue that our work falls into DSN’s call for papers which describe "models...for...assessing dependable systems".
+Cryptographic hash functions are well specified and deterministic and thus we can model the impact of a collision attack on a system in a SAT framework.
+In addition to the five general metrics discussed above, explicitly evaluating the untrustworthiness for a specific use case is a novel advancement and well within the scope of DSN.
+We present the new LCTs as basis for an empirical study of the effectiveness of these metrics and techniques.
+
+We thank all reviewers for their comments, and have incorporated them into our submission.

### 2018-03-05 dsn-2018/bare_conf.tex
DSN 2018 Submission - Rejected (#5)
Submission rejected; draft + updates.

* Initial draft of dsn-2018 paper (#3)

Authored by E. Rozier (@erozier2), A. Scheel (@cipherboy) with feedback
from K. Mayo (@katmayo7), N. De Graaf (@NathanDeGraaf),
N. Mulbrook and others.

* Updated for typos

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Store versioned paper

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Initial rebuttal

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Spelling

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2017-12-05 dsn-2018/bare_conf.tex
@@
 
-%\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
-%\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
-%Ames, Iowa, USA 50011\\
-%Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}
+\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
+\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
+Ames, Iowa, USA 50011\\
+Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}
 
@@
 differential paths and the effectiveness of various techniques presented
-in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Impacts}, we
+in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Conclusion}, we
 discuss the broader impacts and conclusions of this research, and
-conclude with Section \ref{Sec:Conclusion} discussing future work and
-extensions in Section \ref{Sec:Future}.
+conclude by discussing future work and extensions in Section \ref{Sec:Future}.
 
@@
 behavior is the insight that while SAT is, in general, a hard problem
-for many specific and practical instances current SAT-solving
+for many specific and practical instances, current SAT-solving
 technology is quite efficient \cite{vardi2010p}. We
@@
 
-In order to define our define our metrics using logical cryptanalysis
+In order to define our metrics using logical cryptanalysis
 we first define a set of formal terminology and notation whn
@@
 measuring the number of intermediate rounds with non-zero differences.
-We notate this as $\delta(C_1)$.
+We denote this as $\delta(C_1)$.
 
@@
 
-Furthermore this condition extends to $\delta = d$ for some small
+Furthermore, this condition extends to $\delta = d$ for some small
 value of $d$ as determined by computational resources such that $d$
@@
     \item The number of unit distance neighbors a collision class has.
-    \item The maximum count of zeros in a the binary representation of a
+    \item The maximum count of zeros in the binary representation of a
         colliding block (and likewise with ones).
@@
 
-We justify our the distance heuristic, $\delta$, defined in Section
+We justify our distance heuristic, $\delta$, defined in Section
 \ref{terminology:distance} in the existing literature on MD4.  Using
@@
 strategies utilized, cliques may be expanded before new neighbors, or a
-significant portion of a vertices neighbors might belong to existing cliques,
+significant portion of a vertex's neighbors might belong to existing cliques,
 resulting in slow progress towards any external goal.
@@
 neighborhoods, maximum count of zeros in the binary representation of
-a colliding block, the existance of blocks which collide under
+a colliding block, the existence of blocks which collide under
 multiple initial values, and ASCII block collisions as discussed
@@
 
-We present results in Table \ref{table:zeroes} which represent the maximum count of zeroes in a block
+We present results in Table \ref{table:zeros} which represent the maximum count of zeros in a block
 for existing collisions. Note that X. Wang's attack is the leader, with 509
@@
     \caption{Number of Zeroes in a Colliding Block}
-    \label{table:zeroes}
+    \label{table:zeros}
     \begin{tabular}{c c c c}
@@
 Through the exploration of the neighborhood technique, we have found over
-35,000 new differential paths which originated from the five original published
+35,000 new differential paths which originated from the five originally published
 collisions in MD4. This shows that the neighborhood definition is sufficient
@@
 
-We noted in table \ref{table:neighborhood} that the unit distance neighborhood of
+We noted in Table \ref{table:neighborhood} that the unit distance neighborhood of
 the existing attacks were non-empty. By repeatedly expanding neighborhoods of
@@
 neighborhood for a collision class; 3) the maximum count of zeros in a
-binary representation of a colliding block; 4) the existance of blocks
+binary representation of a colliding block; 4) the existence of blocks
 which collide under multiple initial values; and 5) the presence of
-ASCII valued inplut blocks within a collision.  These metrics proved
+ASCII valued input blocks within a collision.  These metrics proved
 capable when evaluating the flexibility of collisions in the published
@@
 
-In order to help the community build upon or work, and for the
-purposes of reproducability and ascertainment of correctness the
+In order to help the community build upon our work, and for the
+purposes of reproducability and ascertainment of correctness, the
 research, data, and all associated code are provided on an open access
@@
 The authors would like to thank XXXXX XXXXX for providing access to
-the hardware used for the
-project.
+the hardware used for the project.
 

### 2018-03-26 s2018-math490/thesis/thesis.tex
Start on thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

\documentclass[10pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasmathematics} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Keccak Thesis}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author{Alexander Scheel}

\affil{Iowa State University}

% Please give the surname of the lead author for the running footer
\leadauthor{Scheel}

% Please add here a significance statement to explain the relevance of your work

% Please include corresponding author, author contribution and author declaration information
\correspondingauthor{E-mail: alexander.m.scheel\@gmail.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:

\begin{abstract}
    TODO
\end{abstract}

\keywords{MATH 490 $|$ Advisor: Dr. Clifford Bergman $|$ SHA-3 $|$ hash\_framework }

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
% \thispagestyle{firststyle}
% \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% Bibliography
% \bibliography{pnas-sample}

\section{A Series of Introductions} \label{sec:intro}
    This paper presents the work of the author towards the completion of the
requirements for the MATH 490 Independent Study course under Dr. Clifford
Bergman at Iowa State University of Science and Technology. This work was an
extension of the author's Honors Project under Dr. Eric Bergman and utilized
the resulting
\href{https://github.com/cipherboy/hash_framework}{hash\_framework} project.
Additional artifacts related to this project can be seen in the
\href{https://github.com/cipherboy/keccak-attacks}{keccak-attacks} repository.

    Within this section are a series of introductions which provide necessary
background on the topics of cryptography, hash functions, the development of
Keccak/SHA-3, and its structure. While certain sections can be skipped if the
reader has the prerequisite knowledge, hopefully all readers find the material
engaging and useful. Following these introductions, this paper presents the
analysis of the Keccak/SHA-3 hash function before concluding with a final
evaluation and further work.

\subsection{Introduction on the Topics of Cryptography} \label{sec:i:crypto}

    While there have been many applications of mathematics, few are as
demanding and shrouded in secrecy as that of cryptography. Cryptography exists
because of the fundamental need of civilizations, governments, and individuals
to keep secrets secure from devoted adversaries.

\subsection{Introduction on the Topics of Hash Functions} \label{sec:i:hash}

    Within cryptography's collection of algorithms, few are are as useful as
hash functions have proven to be to cryptographers and non-cryptographers
alike. A hash function maps arbitrary length binary strings to binary strings
of a fixed length.

\subsection{Introduction on the Development of Keccak/SHA-3} \label{sec:i:keccak}

    FIPS 202 \cite{NIST202}.

\subsection{Introduction on the Structure of SHA-3} \label{sec:i:structure}

\section{Mathematical Properties of the Five Round Functions} \label{sec:properties}

\subsubsection{Bijectivity} \label{sec:p:bijectivity}

\subsubsection{Order of the Permutation} \label{sec:p:order}

\subsubsection{Inverse of the Permutation} \label{sec:p:inverse}

\subsubsection{XOR ($\oplus$) Distributivity} \label{sec:p:xor}

\subsubsection{Generalizations of $\theta$ and $\chi$} \label{sec:p:generalizations}

\subsubsection{Choice of Parameters for $\theta$ and $\chi$} \label{sec:p:generalizations}

\section{Marginal and Differential Properties of the Five Round Functions} \label{sec:differentials}

\subsection{Margin Properties} \label{sec:d:margin}

\subsection{Differential Properties} \label{sec:d:diff}

\subsection{Input Margin Impact on Differential Properties} \label{sec:d:input}

\subsection{Output Margin Impact on Differential Properties} \label{sec:d:output}

\section{Exhaustive Collision Searches} \label{sec:search}

\section{Fixed Point Attacks} \label{sec:fixed}

\subsection{Full Fixed Points} \label{sec:f:full}

\subsection{Partial Fixed Points} \label{sec:f:partial}

\section{Conclusions and Further Work} \label{sec:conclusion}

\section{Bibliography} \label{sec:bibliography}

\bibliography{thesis}

\end{document}

### 2018-03-27 s2018-math490/thesis/thesis.tex
Add outline
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 \begin{abstract}
+    % TODO
     TODO
@@
 
+% TODO
+% More information about structure of analysis section.
+
 \subsection{Introduction on the Topics of Cryptography} \label{sec:i:crypto}
@@
 to keep secrets secure from devoted adversaries.
+
+% TODO Outline
+%    - Historic
+%        - Ciphers vs Codes
+%
+%    - Modern
+%        - Symmetric
+%        - Asymmetric
+%        - Hash Functions
+%        - Building Protocols
 
@@
 
+% TODO Outline
+%   - Properties
+%   - History of Development
+%   - History of Attacks
+%   - Modern Theory and Structure
+
 \subsection{Introduction on the Development of Keccak/SHA-3} \label{sec:i:keccak}
@@
 
+% TODO Outline
+%    - Authors Background (keccak team)
+%    - Function design goals
+%    - NIST Contest
+%    - NIST standardization controversy? (sources?)
+
 \subsection{Introduction on the Structure of SHA-3} \label{sec:i:structure}
+
+    SHA-3 consists of two parts: a core permutation function, KECCAK-$f$, and
+a domain extender, the KECCAK sponge function.
+
+% TODO Outline
+%   - Sponge Function
+%   - State Array
+%   - Five functions
+%   - Keccak-$f$
 
@@
 
-\subsubsection{Bijectivity} \label{sec:p:bijectivity}
+% TODO Outline
+%   - Outline following subsections
 
-\subsubsection{Order of the Permutation} \label{sec:p:order}
+\subsection{Bijectivity} \label{sec:p:bijectivity}
 
-\subsubsection{Inverse of the Permutation} \label{sec:p:inverse}
+% TODO Outline
+%   - Reason why bijectivity is important
+%   - Impact on attacks?
 
-\subsubsection{XOR ($\oplus$) Distributivity} \label{sec:p:xor}
+\subsubsection{Of $\theta$} \label{sec:p:b:t}
 
-\subsubsection{Generalizations of $\theta$ and $\chi$} \label{sec:p:generalizations}
+% TODO Outline
+%   - Theta -> XOR with 11 different values; linear equations?
+%   - What sizes are bijective?
+%   - Discussion in literature
 
-\subsubsection{Choice of Parameters for $\theta$ and $\chi$} \label{sec:p:generalizations}
+\subsubsection{Of $\rho$} \label{sec:p:b:r}
+
+% TODO Outline
+%   - Obvious permutation of bit orders
+
+\subsubsection{Of $\pi$} \label{sec:p:b:p}
+
+% TODO Outline
+%   - Obvious: permutation of bit orders
+
+\subsubsection{Of $\chi$} \label{sec:p:b:c}
+
+% TODO Outline
+%   - Harder, discussion in literature.
+
+\subsubsection{Of $\iota$} \label{sec:p:b:i}
+
+% TODO Outline
+%   - Obvious: XOR with fixed value.
+
+\subsection{Order of the Permutation} \label{sec:p:order}
+
+% TODO Outline
+%   - Outline why interesting
+%   - Attack on PRF/XOF -> low cycle size
+%   - Values over 2^(25w) indicate PRF has smaller upper bound for actual
+%       entropy
+%   - T R P C I
+
+\subsection{Evaluation of the Orders of Composition of Permutations}
+
+% TODO Outline
+%   - w=1 evaluation
+%   - Evaluation of different orderings of trpci
+
+\subsection{Inverse of the Permutation} \label{sec:p:inverse}
+
+% TODO Outline
+%   - Outline simplification of f^{-1}
+%   - Closed form for theta, chi
+%   - Give algorithm for any w?
+
+\subsection{XOR ($\oplus$) Distributivity} \label{sec:p:xor}
+
+% TODO Outline
+%   - Theta, Rho, Pi - free
+%   - Chi, Iota - no
+%   - Makes analysis harder, which is good
+
+\subsection{Generalizations of $\theta$ and $\chi$} \label{sec:p:generalizations}
+
+% TODO Outline
+%   - Evaluation of effect?
+%   - Pending eval?
+
+\subsection{Choice of Parameters and Ordering of Composition} \label{sec:p:generalizations}
+% TODO Outline
+%   - Pending evaluation?
 
@@
 
+% TODO Outline
+%   - Overview of why it matters
+
 \subsection{Margin Properties} \label{sec:d:margin}
+
+% TODO Outline
+%   - Overview of why it matters - one round
+%   - Can see effect of rho in w increasing
+%   - Graphics
 
@@
 
+% TODO Outline
+%   - Graphics
+%   - Impact of importance
+%   - Show composition order impact if any?
+
 \subsection{Input Margin Impact on Differential Properties} \label{sec:d:input}
+
+% TODO Outline
+%   - Graphics
+%   - Impact and importance
 
@@
 
+% TODO Outline
+%   - Graphics
+%   - Impact and importance
+
 \section{Exhaustive Collision Searches} \label{sec:search}
+
+% TODO Outline
+%   - Importance
+%   - Finding patterns hard
+%   - No obvious connection for how to extend w=1 -> w=2
+%   - Searches limited
 
@@
 
+% TODO Outline
+%   - Outline subsections
+
 \subsection{Full Fixed Points} \label{sec:f:full}
+
+% TODO Outline
+%   - Probable due to order analysis
+%   - Impact, attack
 
@@
 
+% TODO Outline
+%   - More probable
+%   - Impact, attack
+
 \section{Conclusions and Further Work} \label{sec:conclusion}
+
+% TODO Outline
+%   - SHA-3 Secure, but introduces new attack vectors
+%   - Need more in the way of LC for SHA-3-esq functions
+%_  - Needs permutation-based cryptanalysis
+%   - Differential cryptanalysis not likely to be of much use
 

### 2018-04-10 s2018-math490/thesis/thesis.tex
Updates to thesis; fleshed out bijectivity sections
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-\title{Keccak Thesis}
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage{amsmath,amsthm,amssymb,enumerate,euscript,mathrsfs}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
+
+\title{A Collection of Thoughts on the Keccak Construction}
 
@@
 the resulting
-\href{https://github.com/cipherboy/hash_framework}{hash\_framework} project.
+\href{https://github.com/cipherboy/hash\_framework}{hash\_framework} project.
 Additional artifacts related to this project can be seen in the
@@
     While there have been many applications of mathematics, few are as
-demanding and shrouded in secrecy as that of cryptography. Cryptography exists
+demanding and shrouded in as much secrecy as cryptography. Cryptography exists
 because of the fundamental need of civilizations, governments, and individuals
-to keep secrets secure from devoted adversaries.
+to keep secrets secure from devoted adversaries. While modern cryptography
+combines the disciplines of mathematics, computer science, and computer
+engineering, prior to the turn of the 20th century, cryptography lacked much
+of its modern rigor.
+
+    Prior to the 20th century, cryptography was split into three major
+branches: ciphers, codes, and stenography.
 
@@
 
+\subsection{Introduction to Terminology} \label{sec:i:terminology}
+
+Define $\Sigma = \{0, 1\}$ to be the alphabet, and for
+$w \in \{1, 2, 4, 8, 16, 32, 64\}$, $S_{w} = \Sigma^{25w}$, to be the set of
+all binary strings of length $25w$.
+
 \section{Mathematical Properties of the Five Round Functions} \label{sec:properties}
+
+In the following section, we detail various mathematical properties of the five
+permutation functions which make up the core round function of Keccak. In most
+cases, we seek to provide mathematical proofs of these properties. In all
+cases, we rely on external code and Boolean Satisfiability for computerized
+proofs of these theorems, in ways independent of the ways proved here. In a
+few cases, we provide references into the existing literature.
 
@@
 
-\subsection{Bijectivity} \label{sec:p:bijectivity}
-
-% TODO Outline
-%   - Reason why bijectivity is important
-%   - Impact on attacks?
-
-\subsubsection{Of $\theta$} \label{sec:p:b:t}
+\subsection{Properties of $\theta$} \label{sec:p:t}
 
@@
 
-\subsubsection{Of $\rho$} \label{sec:p:b:r}
+In this section, we show that $\theta$ is bijective, that XOR distributes
+through $\theta$, give a method for finding the inverse of $\theta$, and
+give the order of $\theta$.
+
+Let $w$ be a fixed power of 2. Since $S_{w}$ is of finite size, it suffices to
+show that, $\forall x, y \in S_{w}$,
+$x \neq y \Rightarrow \theta(x) \neq \theta(y)$. Assume the hypothesis: then
+$x \oplus y \neq 0^{25w}$.
+
+\begin{lemma} \label{lem:p:t:1}
+\begin{align*}
+    \theta(a) = 0^{25w} \iff a = 0^{25w}
+\end{align*}
+\end{lemma}
+\begin{proof}
+
+This follows from the definition of $\theta$: note that
+$\theta(0^{25w}) = 0^{25w}$. If $a \neq 0^{25w}$, then there
+exists an index set $I_{1}$ such that $\forall i \in I_{1}$, $a[i] = 1$,
+and $\forall j \not\in I_{1}$, $a[j] = 0$. Then $\theta(a) \neq 0^{25w}$,
+which follows from the construction of $\theta$. (Note that each output
+bit of $\theta$ is composed of the XOR of 11 values in a fixed pattern).
+
+\end{proof}
+
+\begin{lemma} \label{lem:p:t:2}
+$\forall a, b \in S_{w}$,
+\begin{align*}
+    \theta(a \oplus b) = \theta(a) \oplus \theta(b)
+\end{align*}
+\end{lemma}
+\begin{proof}
+
+This follows from the definition of $\theta$: note that $\theta$ is
+composed entirely of XORs and that XOR is commutative and associative.
+
+\end{proof}
+
+Combining Lemma \ref{lem:p:t:1} and Lemma \ref{lem:p:t:2}, we have that:
+
+\begin{align*}
+    x \oplus y = 0^{25w} & \iff \theta(x \oplus y) = \theta(0^{25w}) \\
+        & \iff \theta(x \oplus y) = 0^{25w} \\
+        & \iff \theta(x) \oplus \theta(y) = 0^{25w}
+\end{align*}
+
+and hence $\theta$ is bijective.
+
+To construct the inverse of $\theta$, note that
+$A'[x, y, z] = A[x, y, z] \oplus D[x, z]$; hence,
+$A[x, y, z] = A'[x, y, z] \oplus D'[x, z]$ for some $D' = D$. Since $D[x, z]$
+is composed of several $C[x, z]$, where
+$C[x, z] = \oplus_{y = 0}^{4} A[x, y, z]$, we can similarly define $C'[x, z]$
+to be $C'[x, z] = \oplus_{y = 0}^{4} A'[x, y, z]$. Then, we expect the inverse
+of $\theta$ to be of a similar form. This reduces to a linear algebra problem
+over boolean variables. We know that $D'[x, z] = D[x, z]$ in order to recover
+$A[x, y, z]$. Hence we can represent $D[x, z]$ as a series of bits of length
+$5 \times w$, where bit $i = z' + 5 \times x'$ is $1$ if and only if
+$C[x', z']$ is used in the construction of $D[x, z]$. Further, we can view each
+of the $C'[x, z]$ as being the conjunction of three $C[x', z']$ in $A$, and thus
+can represent these as bit strings where bit $j = z' + 5\times x'$ is 1 if
+and only if $C[x', z']$ is used to construct $C'[x, z]$. (That is,
+since $C'[x, z] = \oplus_{y = 0}^{4} A'[x, y, z]$,
+$C'[x, z] = \oplus_{y = 0}^{4} (A[x, y, z] \oplus D[x, z])$ and hence
+$C'[x, z] = \oplus_{y = 0}^{4} (A[x, y, z] \oplus C[x', z'] \oplus C[x'', z''])$,
+and thus $C'[x, z] = C[x, z] \oplus C[x', z'] \oplus C[x'', z'']$, for some
+$x, z, x', z', x'', z''$ based on the definition of $\theta$. Thus,
+giving each $C'[x, z]$ a constant $c_{x,z}$ for whether it is used in constructing
+$D'[x, z]$, we can form a system of linear equations and solve for the
+constants $c_{x, z}$ in each expression. Since there are $5 \times w$ variables
+and $5 \times w$ equations in each equation for $D'[x, z]$, this can be solved
+easily, yielding the inverse of $\theta$.
+
+Lastly, we have computed the order of the permutation $\theta$ for all $w$.
+We reproduce them here without proof; they were found by direct search, but
+only verified exhaustively for $w = 1$. In general, the order is given by
+the expression $3 \times w$.
+
+\begin{tabular}{c c} \label{tab:p:t:1}
+    w & Order \\
+    1 & 3 \\
+    2 & 6 \\
+    4 & 12 \\
+    8 & 24 \\
+    16 & 48 \\
+    32 & 96 \\
+    64 & 192 \\
+\end{tabular}
+
+\subsubsection{Of $\rho$} \label{sec:p:r}
 
@@
 
-\subsubsection{Of $\pi$} \label{sec:p:b:p}
+Since $\rho$ is a simple permutation of the location of bits, it holds that
+$\rho(a \oplus b) = \rho(a) \oplus \rho(b)$.
+
+It is obvious that the order of the $\rho$ permutation is $w$: this is obvious
+from the construction of $\rho$.
+
+\subsubsection{Of $\pi$} \label{sec:p:p}
 
@@
 
-\subsubsection{Of $\chi$} \label{sec:p:b:c}
+Since $\pi$ is a simple permutation of the location of bits, it holds trivially
+that $\pi(a \oplus b) = \pi(a) \oplus \pi(b)$.
+
+It is obvious that the order of the $\pi$ permutation is $24$: this is obvious
+from the construction of $\pi$.
+
+\subsubsection{Of $\chi$} \label{sec:p:c}
 
@@
 
-\subsubsection{Of $\iota$} \label{sec:p:b:i}
+It is non-trivial that the order of the $\chi$ function is a constant 4.
+However, this can be verified through the use of SAT to prove that $\chi^{4}$
+is the identity function. Hence $\chi^{-1} = \chi^{3}$.
+
+\subsubsection{Of $\iota$} \label{sec:p:i}
 
@@
 
-\subsection{Order of the Permutation} \label{sec:p:order}
+Note that since $\iota$ is an XOR with a fixed value, it is obvious that
+$\iota$ is a bijection: for any $w$, for any $i$, and for all $x \in S_{w}$,
+$\iota(\iota(x, i), i) = x$, since $x \oplus \iota_{i} \oplus \iota_{i} = x$.
+Hence, $\iota$ is its own inverse and hence $\iota$ is bijective since the
+inverse is well defined for all $x \in S_{w}$.
 
-% TODO Outline
-%   - Outline why interesting
-%   - Attack on PRF/XOF -> low cycle size
-%   - Values over 2^(25w) indicate PRF has smaller upper bound for actual
-%       entropy
-%   - T R P C I
+Lastly, note that it is trivial that the order of the $\iota$ permutation is
+2 by construction (due to the XOR).
 
@@
 %   - Evaluation of different orderings of trpci
-
-\subsection{Inverse of the Permutation} \label{sec:p:inverse}
-
-% TODO Outline
-%   - Outline simplification of f^{-1}
-%   - Closed form for theta, chi
-%   - Give algorithm for any w?
-
-\subsection{XOR ($\oplus$) Distributivity} \label{sec:p:xor}
-
-% TODO Outline
-%   - Theta, Rho, Pi - free
-%   - Chi, Iota - no
-%   - Makes analysis harder, which is good
 

### 2018-04-12 s2018-math490/thesis/thesis.tex
Changes to thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-    While there have been many applications of mathematics, few are as
-demanding and shrouded in as much secrecy as cryptography. Cryptography exists
+    While there are many applications of pure mathematics, few are as
+demanding and shrouded in secrecy as cryptography. Cryptography exists
 because of the fundamental need of civilizations, governments, and individuals
@@
 
-    Prior to the 20th century, cryptography was split into three major
+    Prior to the late 19th century, cryptography was split into three major
 branches: ciphers, codes, and stenography.
@@
 
+\subsection{Introduction to Terminology} \label{sec:i:terminology}
+
+This section contains a collection terminology useful for discussing SHA-3
+and Boolean Satisfiability.
+
+We define $\Sigma = \{0, 1\}$ to be the alphabet.
+
+We define the set $W = \{1, 2, 4, 8, 16, 32, 64\}$ to be the powers of two
+typically used in constructing the Keccak widths; $w=64$ is standardized for
+use in SHA-3, though any power of two can be used.
+
+For a given $w \in W$, we define $S_{w} = \Sigma^{25w}$, to be the set of
+all binary strings of length $25w$; these represent the possible states (which
+are binary strings that map onto an indexible array $A$ described later) and each
+round permutation maps $S_{w} \mapsto S_{w}$.
+
+A binary string $s \in S_{w}$ can be indexed directly (in a zero-indexed
+manner, i.e., the starting bit of $s$ is denoted $s[0]$), or via a 3
+dimensional structure of size 5 by 5 by $w$. This is typically done in
+conjunction with an uppercase letter, $A = s$, and then
+$A[x, y, z] = s[w(5y + x) + z]$. This is in accordance with
+FIPS 202 \cite{NIST202}.
+
 \subsection{Introduction on the Structure of SHA-3} \label{sec:i:structure}
@@
     SHA-3 consists of two parts: a core permutation function, KECCAK-$f$, and
-a domain extender, the KECCAK sponge function.
+a domain extender, the KECCAK sponge function. As discussed by the Keccak
+authors, additional security is given by choosing the permutation functions to
+be bijective, though they need not be (the domain need only be $S_{w}$).
 
@@
 %   - Keccak-$f$
-
-\subsection{Introduction to Terminology} \label{sec:i:terminology}
-
-Define $\Sigma = \{0, 1\}$ to be the alphabet, and for
-$w \in \{1, 2, 4, 8, 16, 32, 64\}$, $S_{w} = \Sigma^{25w}$, to be the set of
-all binary strings of length $25w$.
 
@@
 Lastly, we have computed the order of the permutation $\theta$ for all $w$.
-We reproduce them here without proof; they were found by direct search, but
-only verified exhaustively for $w = 1$. In general, the order is given by
+We reproduce them here without proof; they were found by randomized search, and
+verified with SAT for $w = 1, 2, 4$ and $8$. In general, the order is given by
 the expression $3 \times w$.
@@
 
-\subsection{Evaluation of the Orders of Composition of Permutations}
+\subsection{Evaluation of the Orders of Composition of Permutations} \label{sec:p:composition}
+
+In this section, we discuss how the above five permutation functions compose,
+and the orders of the resulting compositions. We limit our discussion to $w=1$
+for the interests of exhaustive search: $2^{25}$ is possible on commodity
+hardware, $2^{50}$ would require additional resources. In general, we find that
+these permutations interact non-trivially, resulting in cycles of mixed size.
+We reproduce these results in the table below:
+
+\begin{tabular}{c c c c} \label{tab:p:c:1}
+    Function & Order & Number of Cycle & List of Cycles \\
+\end{tabular}
 

### 2018-04-12 s2018-math490/thesis/thesis.tex
Add information about group order
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-\title{A Collection of Thoughts on the Keccak Construction}
+\title{A Collection of Thoughts on the Keccak/SHA-3 Construction}
 
@@
 
-\begin{tabular}{c c c c} \label{tab:p:c:1}
-    Function & Order & Number of Cycle & List of Cycles \\
+\begin{tabular}{c c c c c} \label{tab:p:c:1}
+    Function & Order & Fixed Points & Number of Cycles & List of Cycles \\
+    $\theta$ & 3 & 2097152 & 2 & 1, 3 \\
+    $\rho$ & 1 & 33554432 & 1 & 1 \\
+    $\pi$ & 24 & 4 & 8 & 1, 2, 3, 4, 6, 8, 12, 24 \\
+    $\chi$ & 4 & 32 & 3 & 1, 2, 4 \\
+    $\chi \circ \pi$ & 17360392635484575518934418947500880 $\approx 2^{113.741}$ & 3 & 28 & Not Reproduced \\
+    $\pi \circ \rho \circ \theta$ & 24 & 4 & 8 & 1, 2, 3, 4, 6, 8, 12, 24 \\
+    $\chi \circ \pi \circ \rho \circ \theta$ & 418144575651966378899040573720 $\approx 2^{98.399}$ & 3 & 27 & Not Reproduced \\
+    $r_1$ & 320185339723133697023127516600 $\approx 2^{98.014}$ & 0 & 14 & Not Reproduced \\
+    $r_2$ & $\approx 2^{130.726}$ & 0 & 12 & Not Reproduced \\
+    $r_3$ & $\approx 2^{164.242}$ & 1 & 16 & Not Reproduced \\
+    $r_4$ & $\approx 2^{211.609}$ & 0 & 16 & Not Reproduced \\
+    $r_5$ & $\approx 2^{131.878}$ & 2 & 16 & Not Reproduced \\
+    $r_6$ & $\approx 2^{190.743}$ & 3 & 18 & Not Reproduced \\
+    $r_7$ & $\approx 2^{218.017}$ & 0 & 18 & Not Reproduced \\
+    $r_8$ & $\approx 2^{190.137}$ & 0 & 18 & Not Reproduced \\
+    $r_9$ & $\approx 2^{188.483}$ & 0 & 14 & Not Reproduced \\
+    $r_{10}$ & $\approx 2^{154.432}$ & 0 & 19 & Not Reproduced \\
+    $r_{11}$ & $\approx 2^{223.948}$ & 1 & 18 & Not Reproduced \\
+    $r_{12}$ & $\approx 2^{108.579}$ & 0 & 12 & Not Reproduced \\
+    $r_{13}$ & $\approx 2^{209.785}$ & 1 & 18 & Not Reproduced \\
+    $r_{14}$ & $\approx 2^{170.621}$ & 0 & 14 & Not Reproduced \\
+    $r_{15}$ & $\approx 2^{196.507}$ & 2 & 22 & Not Reproduced \\
+    $r_{16}$ & $\approx 2^{172.643}$ & 1 & 16 & Not Reproduced \\
+    $r_{17}$ & $\approx 2^{221.160}$ & 1 & 22 & Not Reproduced \\
+    $r_{18}$ & $\approx 2^{203.510}$ & 0 & 20 & Not Reproduced \\
+    $r_{19}$ & $\approx 2^{250.748}$ & 3 & 20 & Not Reproduced \\
+    $r_{20}$ & $\approx 2^{183.937}$ & 1 & 19 & Not Reproduced \\
+    $r_{21}$ & $\approx 2^{158.852}$ & 2 & 15 & Not Reproduced \\
+    $r_{22}$ & $\approx 2^{111.874}$ & 1 & 12 & Not Reproduced \\
+    $r_{23}$ & $\approx 2^{230.807}$ & 0 & 24 & Not Reproduced \\
+    $r_{24}$ & $\approx 2^{140.355}$ & 1 & 14 & Not Reproduced
 \end{tabular}
+
+Note that, for $w=1$, $\rho$ is the identity function and is thus omitted from
+the tables above. Note that $r_{n}$ denotes the first $n$ rounds of SHA-3.
+
+While on first glance, large orders make it appear that the hash function is
+strong, SHA-3 also defines an XOF (or eXtensible Output Function) construct.
+This allows SHA-3 to act as a pseudo-random number generator: after hashing
+an input, the internal state of SHA-3 has reached some value $S$. Suppose we
+want to request $m$ bits of state. If $k$ exceeds our security margin $k/2$,
+we take $k/2$ bits, permute $S \rightarrow S'$ according to KECCAK-$f$,
+and repeat until all $m$ bits have been retrieved.
+
+An ideal permutation function for an $XOF$ would ensure that there exists one
+random cycle through all possible states (and thus contain a cycle of $2^{25}$,
+in this case). However, the order of the $r_{n}$ rounds of SHA-3 far exceed
+$2^{25}$ and all of them have several cycles. This suggests that the actual
+security margin of the XOF construct is far less than the theoretical value.
+However, this analysis needs to be extended to at least $w=4$ (performing
+$2^{100}$ iterations of the hash function core which is not feasible) to
+fully verify this.
 

### 2018-04-16 sat-competition-2018/conference.tex
Initial sat conference submission
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

%\documentclass[runningheads]{llncs}
\documentclass[final]{ieee}

\usepackage{microtype} %This gives MUCH better PDF results!
%\usepackage[active]{srcltx} %DVI search
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{fnbreak} %warn for split footnotes
\usepackage{url}
%\usepackage{qtree} %for drawing trees
%\usepackage{fancybox} % if we need rounded corners
%\usepackage{pict2e} % large circles can be drawn
%\usepackage{courier} %for using courier in texttt{}
%\usepackage{nth} %allows to \nth{4} to make 1st 2nd, etc.
%\usepackage{subfigure} %allows to have side-by-side figures
%\usepackage{booktabs} %nice tables
%\usepackage{multirow} %allow multiple cells with rows in tabular
\usepackage[utf8]{inputenc} % allows to write Faugere correctly
\usepackage[bookmarks=true, citecolor=black, linkcolor=black, colorlinks=true]{hyperref}
\hypersetup{
pdfauthor = {Alexander Scheel},
pdftitle = {CryptoMiniSat v5.5},
pdfsubject = {SAT Competition 2018},
pdfkeywords = {SAT Benchmarks, Cryptography, Logical Cryptanalysis},
%pdfcreator = {PdfLaTeX with hyperref package},
%pdfproducer = {PdfLaTex}
}
%\usepackage{butterma}

%\usepackage{pstricks}
\usepackage{graphicx,epsfig,xcolor}

\begin{document}
\title{Logical Cryptanalysis Benchmarks for \\ Classical and Modern Hash Functions}
\author{Alexander Scheel, Iowa State University, alexander.m.scheel@gmail.com}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\section{Introduction}
This collection of benchmarks focuses on new techniques in logical
cryptanalysis for analyzing the structure of various hash functions. We
offer as a benchmark a new technique analyzing the security of a hash function
of the Merkle-D{\aa}mgard construction which are broadly applicable to all such
constructions and several benchmarks analyzing the various security properties
of the Keccak hash function. While the latter techniques have not yeilded
useful cryptanalysis results, they pose as a source of variable scalable
benchmarks due to several choices of parameters (the bitwidth, $w$ and the
number of rounds).

\section{Classical Hash Functions}
The MD4 hash function is composed of 48 iterated rounds, each updating one of
four 32-bit state variables \cite{rfc1320}. While authors such as I. Mironov
have applied SAT solvers to finding collisions given an existing differential
path \cite{Mironov2006} and D. Jovanovi{\'{c}} has applied applied SAT solvers
to brute force finding collisions and preimages in hash functions
\cite{LogicalAnalysis}, these techniques either require pre-existing
collisions or are too hard for SAT solvers on a large number of rounds.

One technique that is efficient for even a modest number of rounds ($\leq 28$)
is the notion of a differential family search. In most cases, the differential
path between two blocks $b_1$ and $b_2$ is the simple XOR difference between
the intermediate rounds. For a given path $p$, there are often many such blocks
which create a collision; I. Mironov showed that SAT solvers can find such
blocks in MD4 and MD5 \cite{Mironov2006} relatively quickly. We can extend this
concept to differential paths as well: given a differential path $p$, its
family is the ordered tuple of indices of the rounds with a non-zero
difference. For example, the family of the differential path introduced by
M. Schlaffer (in \cite{Schlaffer2006}) is:
\begin{align*}
    (1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, 15, 16, 19, 20, 35, 36)
\end{align*}
Thus for a given family, $f$, there are multiple possible differential paths
which have a structure described by $f$.

By using this concept of a differential family, we can divide the search space
into multiple different SAT problems: for a given number of rounds, $r$, there
are $2^{r-4}$ possible differential families (versus $2^{32w-32}$ possible
differential paths). For $r \leq 28$ it is possible to exhaustively search
large portions of the family space; for $32 \leq r \leq 36$ it becomes possible
to find in select cases, and for $r > 36$, it remains impossible in nearly all
cases (due to the large time limit). Our benchmarks are a sample of possible
differential families for $r=24$, mixing both SAT and UNSAT results. These
benchmarks can be found in the \texttt{families} folder \cite{Sat2018}.

\section{Modern Hash Functions}

The notion of a differential family is not as useful for Keccak due to
the interaction between the sponge function and the internal permutation
functions causing the collision space of families to be entirely different
between successive rounds. From an algebraic perspective, and to study the
security margin of the XOF consruct \cite{Keccak3}, it is important to show
that the Keccak round functions are bijective and have high permutation order.
Two of our benchmarks (\texttt{bijection} and \texttt{orders}) model these
problems in SAT. In general, these instances are easy, but with a few notable
outliers: for $w \geq 4$, proving the order of $\theta$ is $3w$ is difficult.
Note that this should simplify to showing that $\forall x, x \neq x$ is UNSAT,
and thus should be relatively trivial; however $w=4$ produces runtimes in
excess of one hour. Further, anything later involving $\theta$ (such as
$\rho \circ \theta$, etc.) also becomes difficult for $w\geq4$.

In \texttt{differences}, we study the differential properties of the
permutation functions: for a given parameters $x, y \leq 25w$ and round
function $f$, we seek to find witnesses $a$ and $b$ such that:
\begin{align*}
    & \#(1, a \oplus b) = x \\
    & \#(1, f(a) \oplus f(b)) = y
\end{align*}
That is, find an input with difference $x$ which produces an output with
difference $y$. There are a few interesting outliers in this model: for
$w\geq4$, most round functions cannot be individually analyzed this way.
Further, while $\pi$ is a permutation of the order of the bits (and thus
does not change the values of any bits), certain instances in $w=2$
where $x \neq y$ produce runtimes in excess of an hour. This is surprising
as the model is trivially SAT if and only if $x == y$ (and the model merely
involves changing the orders of variables).

The benchmarks in \texttt{output-margins} consider the effects of the sponge
function. Since all of the round functions are bijective, if two inputs differ,
the interal state of Keccak must also differ. However, to produce a collision,
only the first $y$ bits of the output must be the same. Thus we can create
a model for a given number of differences $x, z \leq 25w$, security margin
$y \leq 25w$, and round functions $f$:
\begin{align*}
    & \#(1, a xor b) = x \\
    & \#(1, (f(a) xor f(b))[0:y]) = z
\end{align*}
If such a witness a and b exist for $z=0$, then the input difference $x$ is
possible of producing collisions at a security margin $m$. Our provided
benchmarks sample the space for small values of $w$ and relatively few
round functions; for $w\geq 16$ and for any set of functions including
$\theta$, the runtimes become exceedingly long.

The benchmarks in \texttt{xof-state} attempt to recreate the internal state
of Keccak given a series of outputs from the XOF (extensible output function
at a given margin). In general, this is possible for either small values of
$w$ or small numbers of rounds (for larger values of $w$). However, care must
be selected in choosing the base seed, otherwise, there can possibly be
multiple satisfying seeds. However, for a set of inputs with unique solution,
these benchmarks can be extended to contain the output from several rounds of
Keccak. After a threshhold dependent on the margin, these are redundant
information and thus test the SAT solvers to work with larger models which
overfit to the solution.

\section{Thanks}
A special thanks to Mate Soos and his CryptoMiniSat solver
\cite{CryptoMiniSat5} for suggesting these problems be submitted as benchmarks
to the conference.

This work was completed as part of an Undergraduate Honors Research project
at Iowa State University under the guidance of Eric W. Davis and as part of
MATH 490: Independent Study of SHA-3 under the guidance of Clifford Bergman.

\bibliographystyle{splncs03}
\bibliography{competition.bib}

\vfill
\pagebreak

\end{document}

### 2018-04-16 sat-competition-2018/conference.tex
Spellcheck
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 constructions and several benchmarks analyzing the various security properties
-of the Keccak hash function. While the latter techniques have not yeilded
+of the Keccak hash function. While the latter techniques have not yielded
 useful cryptanalysis results, they pose as a source of variable scalable
-benchmarks due to several choices of parameters (the bitwidth, $w$ and the
+benchmarks due to several choices of parameters (the bit width, $w$ and the
 number of rounds).
@@
 between successive rounds. From an algebraic perspective, and to study the
-security margin of the XOF consruct \cite{Keccak3}, it is important to show
+security margin of the XOF construct \cite{Keccak3}, it is important to show
 that the Keccak round functions are bijective and have high permutation order.
@@
 Note that this should simplify to showing that $\forall x, x \neq x$ is UNSAT,
-and thus should be relatively trivial; however $w=4$ produces runtimes in
+and thus should be relatively trivial; however $w=4$ produces run times in
 excess of one hour. Further, anything later involving $\theta$ (such as
@@
 does not change the values of any bits), certain instances in $w=2$
-where $x \neq y$ produce runtimes in excess of an hour. This is surprising
+where $x \neq y$ produce run times in excess of an hour. This is surprising
 as the model is trivially SAT if and only if $x == y$ (and the model merely
@@
 function. Since all of the round functions are bijective, if two inputs differ,
-the interal state of Keccak must also differ. However, to produce a collision,
+the internal state of Keccak must also differ. However, to produce a collision,
 only the first $y$ bits of the output must be the same. Thus we can create
@@
 these benchmarks can be extended to contain the output from several rounds of
-Keccak. After a threshhold dependent on the margin, these are redundant
+Keccak. After a threshold dependent on the margin, these are redundant
 information and thus test the SAT solvers to work with larger models which
-overfit to the solution.
+over fit to the solution.
 

### 2018-04-17 s2018-math490/thesis/thesis.tex
Updates to thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-\subsection{Margin Properties} \label{sec:d:margin}
+\subsection{Marginal Properties} \label{sec:d:margin}
 
@@
 
+In this section, we introduce two potential attacks against Keccak/SHA-3 using
+fixed points in the underlying round function. However, neither of these
+attacks have been proven possible with current analysis except for small values
+of $w$ and small numbers of rounds. The first is a possible attack using full
+fixed points in the core round function, while the latter describes a category
+of partial fixed points.
+
 % TODO Outline
@@
 
+One theoretical attack against SHA-3 is by using a fixed point. If there
+existed a state $x$ such that $h(x) = x$, for $h$ the round function, then
+for all additional blocks, $h(x | b) = h(x \oplus b)$. While fixed points can
+and do occur (see 24-rounds in Table \ref{tab:p:c:1}), using a fixed-point
+attack is more subtle in practice: it is unlikely that a fixed point is a valid
+block (that is, $x$ contains a suffix that is $0^{m}$ for the current margin,
+$m$). Thus one must find a series of blocks $b_1$, $b_2$ such that
+$h(b_1) \oplus b_2 = x$; this in turn extends the attack from being
+$h(x | b) = h(x \oplus b)$ to:
+\begin{align*}
+    x = h(h(b_1) \oplus b_2) & \Rightarrow \\
+    h(x | b) = h(b_1 | b_2 | b) & = h(b_1 | h_2 \oplus b) \\
+    & = h(b_1 | h_2 | 0^{25w} | b)
+\end{align*}
+However finding $b_1$, $b_2$ amounts to a preimage attack and is thus unlikely
+to occur.
+
 \subsection{Partial Fixed Points} \label{sec:f:partial}
+
+Another theoretical attack is using a partial fixed point. Suppose there
+exists a block $x$ such that $h(x) = y$ is also a block (that is, $y$ has
+a suffix of $0^{m}$). In this case, the following is a valid collision,
+for all blocks $b$:
+\begin{align*}
+    h(x \oplus b) = h(y | b)
+\end{align*}
+We have verified the existance of these partial fixed points for small
+$w$ and number of rounds. Furthermore, they are more readily found using SAT
+than finding a fixed point is.
 

### 2018-04-25 s2018-math490/thesis/thesis.tex
Updates to thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
+For $w=1, 2$, we have verified that no fixed points occur which are valid blocks
+for $1 \leq r \leq 24$, where $r$ is the number of rounds, at margins of 4 and
+8 bits.
+
 \subsection{Partial Fixed Points} \label{sec:f:partial}
@@
 than finding a fixed point is.
+
+For $w=1$, we reproduce a table of quantities of partial fixed points per round
+below:
+\begin{tabular}{c c c c c} \label{tab:f:p:1}
+    Rounds & Quantity & Input & Output \\
+    1 & 512 & & \\
+    2 & 567 & & \\
+    3 & 527 & & \\
+    4 & 545 & & \\
+    5 & 488 & & \\
+    6 & 533 & & \\
+    7 & 510 & & \\
+\end{tabular}
+Note that the above table is only for an effective margin of 256-bits (4 bits
+when $w=1$). For an effective margin of 512-bits (8 bits), we found no such
+partial fixed points. Further, example inputs and outputs are given as natural
+numbers versus binary strings for compactness.
+
+The following table describes our limits for finding partial fixed points
+across various sizes of $w$, rounds, and whether or not we've been able to
+complete an exhaustive search of all such instances.
+\begin{tabular}{c c c c c} \label{tab:f:p:2}
+    $w$ & Rounds & Exhaustive & Input & Output \\
+\end{tabular}
 

### 2018-04-26 s2018-math490/thesis/thesis.tex
Changes
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

--- 2018-04-17 s2018-math490/thesis/thesis.tex
@@
 
-It is obvious that the order of the $\rho$ permutation is $w$: this is obvious
+It is obvious that the order of the $\rho$ permutation is $w$: this follows
 from the construction of $\rho$.
@@
 
-It is obvious that the order of the $\pi$ permutation is $24$: this is obvious
+It is obvious that the order of the $\pi$ permutation is $24$: this follows
 from the construction of $\pi$.
@@
 However, this can be verified through the use of SAT to prove that $\chi^{4}$
-is the identity function. Hence $\chi^{-1} = \chi^{3}$.
+is the identity function. Hence $\chi^{-1} = \chi^{3}$. However, due to the
+construction of $\chi$, XOR does not distribute over it.
 
@@
 
-\subsection{Generalizations of $\theta$ and $\chi$} \label{sec:p:generalizations}
+% \subsection{Generalizations of $\theta$ and $\chi$} \label{sec:p:generalizations}
 
@@
 
-\subsection{Choice of Parameters and Ordering of Composition} \label{sec:p:generalizations}
+% \subsection{Choice of Parameters and Ordering of Composition} \label{sec:p:generalizations}
 % TODO Outline
@@
 \section{Marginal and Differential Properties of the Five Round Functions} \label{sec:differentials}
+
+This section studies three approaches to extracting information about the five
+permutations functions using SAT. None of these approaches can directly lead
+to a collision in SHA-3, however, by combining them, possible differential
+trails can be approximated. Sadly for $w \geq 4$, several of these attacks take
+too long to run (even for individual permutation functions) and thus do not
+emit a valid attack for full Keccak.
 
@@
 
+By modeling the problem with SAT, we were able to recreate the work of prior
+authors (\cite{Homsirikamol2012}, \cite{cryptoeprint:2010:285}). However, like
+previous authors, our results do not scale to useful results. For $w=1$, we
+have exhaustively searched the collision space; the results of this search are
+reproduced below:
+
+\begin{tabular}{c c c} \label{tab:s:1}
+    $w$ & $r$ & Number of collisions \\
+    1 & 1 & 1024 \\
+    1 & 2 & 502 \\
+    1 & 3 & 543 \\
+    1 & 4 & 525 \\
+    1 & 5 & 488 \\
+    1 & 6 & 518 \\
+    1 & 7 & 532 \\
+    1 & 8 & 498 \\
+    1 & 9 & 506 \\
+    1 & 10 & 522 \\
+    1 & 11 & 506 \\
+    1 & 12 & 503 \\
+    1 & 13 & 495 \\
+    1 & 14 & 522 \\
+    1 & 15 & 485 \\
+    1 & 16 & 540 \\
+    1 & 17 & 467 \\
+    1 & 18 & 490 \\
+    2 & 1 & $ > 2^{20}$ \\
+    2 & 2 & $ > 2^{14}$ \\
+    2 & 3 & $ > 74$ \\
+\end{tabular}
+
+Note that the results for $w=2$ are incomplete; the search was terminated
+after one week. Further, all of these used an effective margin of 512-bits.
+In our limited time, we were unable to find any useful patterns in this data.
+
 % TODO Outline
@@
 %   - Searches limited
+
+\section{Correlation Matrices} \label{sec:correlation}
+
+Another area of study was the resistance of Keccak to correlation attacks.
+That is, given some structured input (in this case, a valid block where the
+remaining bits are all zeros), does SHA-3 emit any useful two bit correlations
+across rounds? Here, we constructed a program to generate matricies of
+correlation for all two-variable boolean functions given some state. We then
+performed Monte-Carlo simulation; for $w=8$ and $r=24$, we found no useful
+correlations under sufficiently large starting states ($>16$ million). Thus,
+Keccak is secure from cross-round first and second order correlations; all
+distinguisher attacks must thus rely on at least third-order correlations. For
+$w=8$, checking all third-order correlations across 24 rounds becomes
+computationally infeasible and thus was not performed, since it is not likely
+to yeild useful results.
 
@@
 than finding a fixed point is.
+
+For $w=1$, we reproduce a table of quantities of partial fixed points per round
+below for the first 8 rounds:
+
+\begin{tabular}{c c c c c} \label{tab:f:p:1}
+    Rounds & Quantity & Input & Output \\
+    1 & 512 & \texttt{FFFFFFFFFFFFFFFFFFFFFFFFF} & \texttt{TFFFFFFFFFFFFFFFFFFFFFFFF} \\
+    2 & 567 & \texttt{FTFTTTTFFTTTFFTFTFFFFFFFF} & \texttt{TTTFTFFTFTTTFTFFTFFFFFFFF} \\
+    3 & 527 & \texttt{FFTFFFFTFTFFFFFFFFFFFFFFF} & \texttt{TTTFTTTTFTFTFTFTFFFFFFFFF} \\
+    4 & 545 & \texttt{TTTFFTTTFFTTFTTFFFFFFFFFF} & \texttt{FFFTTTTFTTFFFTFTFFFFFFFFF} \\
+    5 & 488 & \texttt{FFFFFFFFTFFFTFFFFFFFFFFFF} & \texttt{TFFTFFTFFFFFFTFFTFFFFFFFF} \\
+    6 & 533 & \texttt{FFTFFFFTTFFFTFFFFFFFFFFFF} & \texttt{TTTFFFFTFFFFTTFFTFFFFFFFF} \\
+    7 & 510 & \texttt{TFFFTTTFTTTTFFTFFFFFFFFFF} & \texttt{FTFFFFFFFTFTTTFTTFFFFFFFF} \\
+\end{tabular}
+
+Note that the above table is only for an effective margin of 256-bits (4 bits
+when $w=1$). For an effective margin of 512-bits (8 bits), we found no such
+partial fixed points.
+
+However, brute forcing a partial fixed point for larger values of $w$ and $r$
+quickly grows impossible. A similar technique to the aforementioned
+differential matrices could extend these techniques to larger values of $w$
+and $r$. Instead of building a model to check for a specific number of
+differences, instead build a model to check for a specific number of zeros in
+the inputs and outputs. This could provide a reduction in number of search
+paths for the SAT solver, and allow for more explicit parallelism.
 

### 2018-04-29 s2018-math490/thesis/thesis.tex
Clean up discussion of chi's order
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 
-\subsection{Properties of $\theta$} \label{sec:p:t}
+\subsection{Of $\theta$} \label{sec:p:t}
 
@@
 
-It is non-trivial that the order of the $\chi$ function is a constant 4.
-However, this can be verified through the use of SAT to prove that $\chi^{4}$
-is the identity function. Hence $\chi^{-1} = \chi^{3}$. However, due to the
-construction of $\chi$, XOR does not distribute over it.
+The order of the $\chi$ function is 4, and is independent of the width
+$w$. This is because $\chi$ is independent of both $y$ and $z$ coordinate.
+We will show algebraically that the order of $\chi$ is 4, however, it has also
+been verified with SAT. We first present three basic lemmas without proof:
+
+\begin{lemma} \label{lem:p:c:1}
+\begin{align*}
+    \lnot(a \oplus b) = \lnot a \oplus b
+\end{align*}
+\end{lemma}
+
+\begin{lemma} \label{lem:p:c:2}
+\begin{align*}
+    (a \oplus b) \land (c \oplus d) = (a \land c) \oplus (a \land d) \oplus (b \land c) \oplus (b \land d)
+\end{align*}
+\end{lemma}
+
+\begin{lemma} \label{lem:p:c:3}
+\begin{align*}
+    \lnot(a \oplus b) \land (c \oplus d) = (\lnot a \oplus b) \land (c \oplus d)
+\end{align*}
+\end{lemma}
+
+We define a row of the state cube, $A$ to have elements $a_1$, $a_2$, $a_3$,
+$a_4$, and $a_5$. After applying $\chi$ to this row, we define the resulting
+output row to be $b_1$, $b_2$, $b_3$, $b_4$, $b_5$, where:
+\begin{align*}
+    b_1 & = a_1 \oplus (\lnot a_2 \land a_3) \\
+    b_2 & = a_2 \oplus (\lnot a_3 \land a_4) \\
+    b_3 & = a_3 \oplus (\lnot a_4 \land a_5) \\
+    b_4 & = a_4 \oplus (\lnot a_5 \land a_1) \\
+    b_5 & = a_5 \oplus (\lnot a_1 \land a_2) \\
+\end{align*}
+And similarly for $c_1$ ... $c_5$,  $d_1$ ... $d_5$, $e_1$ ... $e_5$. Thus, to
+show $\chi$ has order 4, it suffices to show that $e_1 = a_1$, (since $\chi$
+has rotational symmetry; if $e_1 = a_1$, then $e_2 = a_2$ ... $e_5 = a_5$).
+
+\begin{lemma} \label{lem:p:c:4}
+\begin{align*}
+    c_1 = & b_1 \oplus (\lnot b_2 \land b_3) \\
+        = & (a_1 \oplus (\lnot a_2 \land a_3)) \oplus (\lnot (a_2 \oplus (\lnot a_3 \land a_4)) \land (a_3 \oplus (\lnot a_4 \land a_5))) \\
+        = & a_1 \oplus (\lnot a_2 \land a_3) \oplus (\lnot a_2 \land a_3) \oplus (\lnot a_2 \land \lnot a_4 \land a_5) \\
+          & \oplus (a_3 \land \lnot a_3 \land a_4) \oplus (a_3 \land a_4 \land \lnot a_4 \land a_5) \text{ (by Lemma \ref{lem:p:c:3} )} \\
+        = & a_1 \oplus (\lnot a_2 \land \lnot a_4 \land a_5) \\
+\end{align*}
+And:
+\begin{align*}
+    c_2 = & a_2 \oplus (\lnot a_3 \land \lnot a_5 \land a_1) \\
+    c_3 = & a_3 \oplus (\lnot a_4 \land \lnot a_1 \land a_2) \\
+    c_4 = & a_4 \oplus (\lnot a_5 \land \lnot a_2 \land a_3) \\
+    c_5 = & a_5 \oplus (\lnot a_1 \land \lnot a_3 \land a_4) \\
+\end{align*}
+\end{lemma}
+
+Thus, to show $e_1 = a_1$:
+\begin{align*}
+    e_1 = & d_1 \oplus (\lnot d_2 \land d_3) \\
+        = & c_1 \oplus (\lnot c_2 \land \lnot c_4 \land c_5) \\
+        = & (a_1 \oplus (\lnot a_2 \land \lnot a_4 \land a_5)) \oplus (\lnot (a_2 \oplus (\lnot a_3 \land \lnot a_5 \land a_1)) \\
+          & \land \lnot (a_4 \oplus (\lnot a_5 \land \lnot a_2 \land a_3)) \\
+          & \land (a_5 \oplus (\lnot a_1 \land \lnot a_3 \land a_4))) \\
+        = & a_1
+\end{align*}
+
+Hence $\chi$ has order 4 and $\chi^{-1} = \chi^{3}$. The Keccak reference states:
+``We refer to [20, Section 6.6.2] for an algorithm for computing the inverse of $\chi$.''
+\cite{Keccak3}; this points a reference in J. Daemen's thesis \cite{DaemenThesis}.
+However, given $\chi$, $\chi^{3}$ is easy to compute and thus $\chi^{-1}$ is also
+easy to compute.
+
+Lastly, note that XOR does not distribute over $\chi$ due to the
+introduction of the and.
 
@@
 and $r$. Instead of building a model to check for a specific number of
-differences, instead build a model to check for a specific number of zeros in
+differences, build a model to check for a specific number of zeros in
 the inputs and outputs. This could provide a reduction in number of search
@@
 \section{Conclusions and Further Work} \label{sec:conclusion}
+
+Overall, Keccak/SHA-3 remains secure to many attacks from SAT solvers.
 

### 2018-04-29 s2018-math490/thesis/thesis.tex
Additional inverse
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 cases, we rely on external code and Boolean Satisfiability for computerized
-proofs of these theorems, in ways independent of the ways proved here. In a
-few cases, we provide references into the existing literature.
+proofs where proofs are not directly provided. We do note the introduction of
+a cleaner form of $\chi^{-1}$ than in existing literature.
 
@@
 However, given $\chi$, $\chi^{3}$ is easy to compute and thus $\chi^{-1}$ is also
-easy to compute.
+easy to compute. Alternatively, using the above, it is easy to see that:
+\begin{align*}
+    d_1 = & a_1 \oplus (\lnot a_2 \land a_3) \oplus (\lnot a_2 \land \lnot a_4 \land a_5)
+\end{align*}
 

### 2018-04-29 s2018-math490/thesis/thesis.tex
Update thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 over boolean variables. We know that $D'[x, z] = D[x, z]$ in order to recover
-$A[x, y, z]$. Hence we can represent $D[x, z]$ as a series of bits of length
+$A[x, y, z]$. Hence we can represent $D[x, z]$ as strings of length
 $5 \times w$, where bit $i = z' + 5 \times x'$ is $1$ if and only if
 $C[x', z']$ is used in the construction of $D[x, z]$. Further, we can view each
-of the $C'[x, z]$ as being the conjunction of three $C[x', z']$ in $A$, and thus
-can represent these as bit strings where bit $j = z' + 5\times x'$ is 1 if
-and only if $C[x', z']$ is used to construct $C'[x, z]$. (That is,
-since $C'[x, z] = \oplus_{y = 0}^{4} A'[x, y, z]$,
+of the $C'[x, z]$ as being the conjunction of three $C[x', z']$ in $A$; thus
+these are strings where bit $j = z' + 5\times x'$ is 1 if and only if
+$C[x', z']$ is used to construct $C'[x, z]$.
+
+(That is, since $C'[x, z] = \oplus_{y = 0}^{4} A'[x, y, z]$,
 $C'[x, z] = \oplus_{y = 0}^{4} (A[x, y, z] \oplus D[x, z])$ and hence
@@
 and thus $C'[x, z] = C[x, z] \oplus C[x', z'] \oplus C[x'', z'']$, for some
-$x, z, x', z', x'', z''$ based on the definition of $\theta$. Thus,
-giving each $C'[x, z]$ a constant $c_{x,z}$ for whether it is used in constructing
-$D'[x, z]$, we can form a system of linear equations and solve for the
-constants $c_{x, z}$ in each expression. Since there are $5 \times w$ variables
-and $5 \times w$ equations in each equation for $D'[x, z]$, this can be solved
-easily, yielding the inverse of $\theta$.
+$x, z, x', z', x'', z''$ based on the definition of $\theta$).
+
+Thus, giving each $C'[x, z]$ a constant $c_{x,z}$ for whether it is used in
+constructing $D'[x, z]$, we can form a system of linear equations and solve for
+the constants $c_{x, z}$ in each expression. Since there are $5 \times w$
+variables and $5 \times w$ equations in each equation for $D'[x, z]$, this can
+be solved easily, yielding the inverse of $\theta$.
 
@@
 
-Overall, Keccak/SHA-3 remains secure to many attacks from SAT solvers.
+Overall, Keccak/SHA-3 remains secure to many attacks from SAT solvers. The
+majority of the techniques discussed here do not scale to the full $w=64$,
+rendering them ineffective for attacking SHA-3 in general. More efficient
+techniques are thus necessary for finding collisions in SHA-3; this likely
+involves finding ways in which structure in small values of $w$ translate to
+structure in larger values of $w$.
+
+Future work includes scaling the partial fixed point attacks via intermediate
+models, evaluating weaknesses in alternate constructions, and improving
+the run time of marginal and differential analysis. However, without
+significant breakthroughs in logical cryptanalysis, most of these attacks are
+unlikely. Further, it is not immediately obvious what effect low permutation
+order has on the possibility of collisions; introducing permutation-based
+cryptanalysis could result in new techniques applicable to SHA-3.
 

### 2018-04-29 s2018-math490/thesis/thesis.tex
thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
     Within this section are a series of introductions which provide necessary
-background on the topics of cryptography, hash functions, the development of
+background on the topics of cryptographic hash functions, the development of
 Keccak/SHA-3, and its structure. While certain sections can be skipped if the
@@
 % More information about structure of analysis section.
-
-\subsection{Introduction on the Topics of Cryptography} \label{sec:i:crypto}
-
-    While there are many applications of pure mathematics, few are as
-demanding and shrouded in secrecy as cryptography. Cryptography exists
-because of the fundamental need of civilizations, governments, and individuals
-to keep secrets secure from devoted adversaries. While modern cryptography
-combines the disciplines of mathematics, computer science, and computer
-engineering, prior to the turn of the 20th century, cryptography lacked much
-of its modern rigor.
-
-    Prior to the late 19th century, cryptography was split into three major
-branches: ciphers, codes, and stenography.
 
@@
 
-\subsection{Introduction on the Topics of Hash Functions} \label{sec:i:hash}
+\subsection{Introduction on the Topics of Cryptographic Hash Functions} \label{sec:i:hash}
 
-    Within cryptography's collection of algorithms, few are are as useful as
+    While there are many applications of pure mathematics, few are as
+demanding and shrouded in secrecy as cryptography. Cryptography exists
+because of the fundamental need of civilizations, governments, and individuals
+to keep secrets secure from devoted adversaries. While modern cryptography
+combines the disciplines of mathematics, computer science, and computer
+engineering, prior to the turn of the 20th century, cryptography lacked much
+of its modern rigor.
+
+    Within cryptography's collection of algorithms, few are as useful as
 hash functions have proven to be to cryptographers and non-cryptographers
 alike. A hash function maps arbitrary length binary strings to binary strings
-of a fixed length.
+of a fixed length. Because the inputs are of arbitrary size, by the pigeonhole
+principle, there must exist at least one collision. However, to be considered
+cryptographically secure, finding such a collision must be hard (on the order
+of $2^{\frac{b}{2}}$ for $b$ the number of bits in the output).
+
+    Further, a cryptographic hash function must be resistant finding a
+preimage: for a given output value $v$, finding an input $x$ such that
+$f(x) = v$ should be roughly on the order of $2^{b}$.
 

### 2018-05-01 s2018-math490/thesis/thesis.tex
Finished thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
     % TODO
-    TODO
 \end{abstract}
@@
 \href{https://github.com/cipherboy/keccak-attacks}{keccak-attacks} repository.
+A permanent location for this document is in the
+\href{https://github.com/cipherboy/papers}{papers} repository.
 
@@
 evaluation and further work.
+
+    The remaining sections outside of the introductions discuss 1) various
+mathematical properties of the core round functions, 2) exhaustive collision
+searches on small instances, 3) correlation matrices, and 4) fixed point
+attacks.
 
@@
 preimage: for a given output value $v$, finding an input $x$ such that
-$f(x) = v$ should be roughly on the order of $2^{b}$.
+$f(x) = v$ should be roughly on the order of $2^{b}$. Given such an attack,
+it is possible to find a collision: choose a random input, compute its hash,
+then use the preimage attack to find a different input for that hash; these two
+inputs form a collision pair. Thus a preimage attack is stronger than a
+collision attack.
+
+    However, if the attacker is allowed to choose only one message (that is,
+the challenger chooses the other), this forms a second preimage attack. More
+formally, for a given fixed message $x$, the attacker seeks a $y \neq x$ such
+that $h(x) = h(y)$. While a collision attack is only valid for some such
+messages, $x$, a preimage attack has to be valid for all such $x$; thus it is
+a stronger attack than collision attacks, but weaker than a preimage attack
+as it presupposes an existing message.
+
+    This independent study focused on a single hash function, Keccak, which
+was chosen by NIST to be the next standardized algorithm, SHA-3 \cite{NIST202}.
+This hash function differs from previously standardized algorithms in two major
+ways: one, it is not based on the Merkel-Damg{\aa}rd construct and is rather
+based on the Sponge construction, and two, the internal round function is not a
+compression function and is instead a permutation. Outside of a few theorems
+done by hand and verifying the lack of low-order correlations, we focused on
+using automated tools to learn more about the structure of SHA-3.
+
+    In the late 1990s, F. Massacci started analyzing cryptographic
+constructions using Boolean Satisfability \cite{CryptoSAT}; however, applying
+this to hash functions is a largely understudied field. E. Homsirikamol tried
+brute forcing collisions in SHA-3 to limited results \cite{Homsirikamol2012},
+and P. Morawiecki performed similar work for preimages \cite{cryptoeprint:2010:285}.
+However, neither contributed useful, general cryptanalysis techniques to scale
+attacks against reduced rounds (and reduced sizes) to larger numbers of rounds.
+Our work begins exploring new techniques and several consequences in the
+following sections.
 
@@
 %   - Modern Theory and Structure
-
-\subsection{Introduction on the Development of Keccak/SHA-3} \label{sec:i:keccak}
-
-    FIPS 202 \cite{NIST202}.
-
-% TODO Outline
-%    - Authors Background (keccak team)
-%    - Function design goals
-%    - NIST Contest
-%    - NIST standardization controversy? (sources?)
 
@@
 
-\subsection{Introduction on the Structure of SHA-3} \label{sec:i:structure}
+\subsection{Introduction on the Development of Keccak/SHA-3} \label{sec:i:keccak}
 
-    SHA-3 consists of two parts: a core permutation function, KECCAK-$f$, and
-a domain extender, the KECCAK sponge function. As discussed by the Keccak
-authors, additional security is given by choosing the permutation functions to
-be bijective, though they need not be (the domain need only be $S_{w}$).
+    NIST standardized Keccak as SHA-3 in FIPS 202 \cite{NIST202}; Keccak was
+chosen as the winner of the corresponding hash function design competition
+started in 2009. Designed by the Keccak Team (G. Bertoni, J. Daemen, M.
+Peeters, G. Van Assche, and R. Van Keer \cite{KeccakTeam}), Keccak features a
+novel construction.
+
+    Keccak is built around the concept of a state cube: a 5 by 5 by $w$ cube
+of bits (where $w$ is a power of 2). By increasing $w$, a higher security
+margin can be given, at the expense of increasing the computational intensity.
+However, because internal round operations on this state cube are bijective,
+it can be implemented efficiently in hardware with minimal gate and propagation
+delays. Typically this state is called $A$ when represented as a cube, and $s$
+when represented as a (flat) bit string. There are five major functions which
+compose to form a single round: $\theta$, $\rho$, $\pi$, $\chi$, and $\iota$;
+each maps bijectively over the entire state space (i.e., elements of
+$2^{25w}$).
+
+    $\theta$ is an eleven way XOR and is thus a linear map. For all
+$0 \leq x \leq 4$ and $0 \leq z < w$, define:
+\begin{align*}
+    C[x, z] & := \oplus_{0 \leq y \leq 4} A[x, y, z] \\
+    D[x, z] & := C[(x-1) \text{ mod } 5, z] \oplus C[(x+1) \text{ mod } 5, (z - 1) \text{ mod } w] \\
+\end{align*}
+Then:
+\begin{align*}
+    A'[x, y, z] & := A[x, y, z] \oplus D[x, z]
+\end{align*}
+
+    $\rho$ is a pure permutation function, permuting the locations of bits in
+a lane (the $z$ component) of the state cube. For all
+$0 \leq z < w: A'[0, 0, z] := A[0, 0, z]$. Then let $x, y = (0, 0)$, and for
+$0 \leq t \leq 23$:
+\begin{align*}
+    \forall 0 \leq z < w: A'[x, y, z] & := A[x, y, (z - \frac{(t+1)(t+2)}{2}) \text{ mod } w] \\
+    let x, y & = (y, (2x + 3y) \text{ mod } 5)
+\end{align*}
+
+    $\pi$ is another pure permutation function, permuting locations along the
+face of the cube. For all $0 \leq x \leq 4, 0 \leq y \leq 4, 0 \leq z < w$:
+\begin{align*}
+    A'[x, y, z] := A[(x + 3y) \text{ mod } 5, x, z].
+\end{align*}
+
+    $\chi$ is the only non-linear function and is of degree one (with inverse
+of degree three). For all $0 \leq x \leq 4, 0 \leq y \leq 4, 0 \leq z < w$:
+\begin{align*}
+    A'[x, y, z] := A[x, y, z] \oplus (\lnot A[(x + 1) \text{ mod } 5, y, z] \land A[(x + 2) \text{ mod } 5, y, z]).
+\end{align*}
+
+    $\iota$ is a fixed-value XOR useful for preventing trivial fixed points,
+and is the only function dependent on the current round number. Refer to
+page 16 of FIPS 202 (\cite{NIST202}) for its specification; it is not discussed
+extensively in this paper.
+
+    The sponge construction can be described as follows. Fix $w$ as a power of
+two. To have a security margin of $m$ bits, the output needs to have $m$ bits;
+however, if the attacker has complete control of all $25w$ bits of state,
+they would be able to choose the output easily because Keccak is bijective.
+Thus, restrict the input to $25w - 2m$ bits and set the remaining $2m$ bits to
+zero. To find a preimage for a $m$-bit output, the attacker would have to find
+a $25w - m$ bit suffix such that the preimage of that suffix has $2m$ zeros.
+Similarly, for a collision attack, an attacker has to find two $25w-2m$
+messages such that their hashes have the same prefix of length $m$. Under an
+ideal round permutation, clearly these are non-trivial and hard to find in
+general.
+
+    For a multiple block message the procedure is thus as follows: call the
+composition of 24 rounds to be \texttt{KECCAK-$f$}. Start with $s=0^{25w}$.
+Splitting the message into chunks of $25w-2m$ (and padding according to
+the scheme defined in FIPS 202 \cite{NIST202}), for each block $b$, extend
+$b$ to length $25w$ with zeros, then compute $s = f(s \oplus b)$. From $s$,
+take the top $m$ bits as the output of the hash function.
+
+% TODO Outline
+%    - Authors Background (keccak team)
+%    - Function design goals
+%    - NIST Contest
+%    - NIST standardization controversy? (sources?)
+
+\subsection{Introduction to the Remaining Sections}
+
+%\subsection{Introduction on the Structure of SHA-3} \label{sec:i:structure}
+
+%    SHA-3 consists of two parts: a core permutation function, KECCAK-$f$, and
+%a domain extender, the KECCAK sponge function. As discussed by the Keccak
+%uthors, additional security is given by choosing the permutation functions to
+%be bijective, though they need not be (the domain need only be $S_{w}$).
 
@@
 
-\section{Marginal and Differential Properties of the Five Round Functions} \label{sec:differentials}
+%\section{Marginal and Differential Properties of the Five Round Functions} \label{sec:differentials}
 
-This section studies three approaches to extracting information about the five
-permutations functions using SAT. None of these approaches can directly lead
-to a collision in SHA-3, however, by combining them, possible differential
-trails can be approximated. Sadly for $w \geq 4$, several of these attacks take
-too long to run (even for individual permutation functions) and thus do not
-emit a valid attack for full Keccak.
+%This section studies three approaches to extracting information about the five
+%permutations functions using SAT. None of these approaches can directly lead
+%to a collision in SHA-3, however, by combining them, possible differential
+%trails can be approximated. Sadly for $w \geq 4$, several of these attacks take
+%too long to run (even for individual permutation functions) and thus do not
+%emit a valid attack for full Keccak.
 
@@
 
-\subsection{Marginal Properties} \label{sec:d:margin}
+%\subsection{Marginal Properties} \label{sec:d:margin}
 
@@
 
-\subsection{Differential Properties} \label{sec:d:diff}
+%\subsection{Differential Properties} \label{sec:d:diff}
 
@@
 
-\subsection{Input Margin Impact on Differential Properties} \label{sec:d:input}
+%\subsection{Input Margin Impact on Differential Properties} \label{sec:d:input}
 
@@
 
-\subsection{Output Margin Impact on Differential Properties} \label{sec:d:output}
+%\subsection{Output Margin Impact on Differential Properties} \label{sec:d:output}
 
@@
 order has on the possibility of collisions; introducing permutation-based
-cryptanalysis could result in new techniques applicable to SHA-3.
+cryptanalysis could result in new techniques applicable to SHA-3. Additional
+work could include automatic lemmaizing of found constraints and including
+them in future searches.
 

### 2018-05-07 s2018-math490/thesis/thesis.tex
Last changes
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>
Updates to pdf

@@
 \begin{abstract}
-    % TODO
+    In this paper, we discuss new techniques for the cryptanalysis of
+    Keccak/SHA-3, though none of the resulting attacks scale to full versions
+    of the hash function. We show an explicit construction for the inverses of
+    the core round functions, the permutation orders of the round functions,
+    and an attack which exploits low-order cycles. Further, we introduce new
+    logical cryptanalysis techniques and show the existence of partial fixed
+    points to produce 2-1 block collisions.
 \end{abstract}
@@
 Bergman at Iowa State University of Science and Technology. This work was an
-extension of the author's Honors Project under Dr. Eric Bergman and utilized
+extension of the author's Honors Project under Dr. Eric Davis and utilized
 the resulting
-\href{https://github.com/cipherboy/hash\_framework}{hash\_framework} project.
+\href{https://github.com/cipherboy/hash\_framework}{hash\_framework} project
+(viewable under the \texttt{cipherboy} account on GitHub.com).
 Additional artifacts related to this project can be seen in the
@@
 alike. A hash function maps arbitrary length binary strings to binary strings
-of a fixed length. Because the inputs are of arbitrary size, by the pigeonhole
-principle, there must exist at least one collision. However, to be considered
-cryptographically secure, finding such a collision must be hard (on the order
-of $2^{\frac{b}{2}}$ for $b$ the number of bits in the output).
+of a fixed length; typically we denote such a function $h$. Because the inputs
+are of arbitrary size, by the pigeonhole principle, there must exist at least
+one collision: two blocks $x \neq y$ such that $h(x) = h(y)$. However, to be
+considered cryptographically secure, finding such a collision must be hard
+(on the order of $2^{\frac{b}{2}}$ for $b$ the number of bits in the output).
 
@@
 preimage: for a given output value $v$, finding an input $x$ such that
-$f(x) = v$ should be roughly on the order of $2^{b}$. Given such an attack,
+$h(x) = v$ should be roughly on the order of $2^{b}$. Given such an attack,
 it is possible to find a collision: choose a random input, compute its hash,
@@
 as it presupposes an existing message.
+
+\subsection{Introduction on the Usages of Hash Functions}
+
+    Studying the integrity of hash function is important because hash functions
+are a fundamental building block of both cryptographic and non-cryptographic
+systems. While a body of literature developed around attacking the previous
+hash functions (of MD4, MD5, SHA-1 and SHA-2), many of these attacks do not
+directly translate to Keccak/SHA-3 because its construction differs
+substantially. Further, many of these attacks started only as generic attacks
+with low utility, and only with substantial work, were they translated to
+highly specialized attacks on specific use cases. Below is a discussion of
+various places where cryptographic hash functions are used, and what types of
+attacks they are most susceptible to.
+
+    The most common use case for hash functions are in signatures. Often times
+the cryptography underlying asymmetric signatures (where one party can sign a
+message and any party can verify that the signature and message is authentic)
+can only sign fixed length messages, often only in the thousands of bytes.
+However, longer data frequently must be signed. Thus, most implementations of
+signature schemes use a cryptographically secure hash function and sign the
+digest instead of the actual message. This can be seen in PGP, TLS, and PKI.
+Often, these implementations are most vulnerable to a second preimage attack,
+as one message is already provided. However, when the attacker has control
+over both messages, a collision attack will suffice.
+
+    Another common use case for hash functions are in message integrity
+verification. In this case, a hash of the message is sent along side of the
+message to ensure that the message hasn't been tampered with. While often this
+signature should be signed, it is frequently used in cases where the
+authenticity of the hash isn't questioned. This includes verifying contents of
+files on disk or in version control systems such as Git. In both of these
+cases, second preimage and collision attacks are the most useful for an
+attacker.
+
+\subsection{Introduction on the Scope of the Independent Study}
 
@@
 \begin{align*}
-    C[x, z] & := \oplus_{0 \leq y \leq 4} A[x, y, z] \\
+    C[x, z] & := \bigoplus_{0 \leq y \leq 4} A[x, y, z] \\
     D[x, z] & := C[(x-1) \text{ mod } 5, z] \oplus C[(x+1) \text{ mod } 5, (z - 1) \text{ mod } w] \\
@@
 a lane (the $z$ component) of the state cube. For all
-$0 \leq z < w: A'[0, 0, z] := A[0, 0, z]$. Then let $x, y = (0, 0)$, and for
+$0 \leq z < w: A'[0, 0, z] := A[0, 0, z]$. Then let $(x, y) = (1, 0)$, and for
 $0 \leq t \leq 23$:
@@
     \forall 0 \leq z < w: A'[x, y, z] & := A[x, y, (z - \frac{(t+1)(t+2)}{2}) \text{ mod } w] \\
-    let x, y & = (y, (2x + 3y) \text{ mod } 5)
+    \text{ let } (x, y) & = (y, (2x + 3y) \text{ mod } 5)
 \end{align*}
@@
 
-    $\chi$ is the only non-linear function and is of degree one (with inverse
+    $\chi$ is the only non-linear function and is of degree two (with inverse
 of degree three). For all $0 \leq x \leq 4, 0 \leq y \leq 4, 0 \leq z < w$:
@@
 
-\subsection{Introduction to the Remaining Sections}
+%\subsection{Introduction to the Remaining Sections}
 
@@
 is composed of several $C[x, z]$, where
-$C[x, z] = \oplus_{y = 0}^{4} A[x, y, z]$, we can similarly define $C'[x, z]$
-to be $C'[x, z] = \oplus_{y = 0}^{4} A'[x, y, z]$. Then, we expect the inverse
+$C[x, z] = \bigoplus_{y = 0}^{4} A[x, y, z]$, we can similarly define $C'[x, z]$
+to be $C'[x, z] = \bigoplus_{y = 0}^{4} A'[x, y, z]$. Then, we expect the inverse
 of $\theta$ to be of a similar form. This reduces to a linear algebra problem
@@
 
-(That is, since $C'[x, z] = \oplus_{y = 0}^{4} A'[x, y, z]$,
-$C'[x, z] = \oplus_{y = 0}^{4} (A[x, y, z] \oplus D[x, z])$ and hence
-$C'[x, z] = \oplus_{y = 0}^{4} (A[x, y, z] \oplus C[x', z'] \oplus C[x'', z''])$,
+(That is, since $C'[x, z] = \bigoplus_{y = 0}^{4} A'[x, y, z]$,
+$C'[x, z] = \bigoplus_{y = 0}^{4} (A[x, y, z] \oplus D[x, z])$ and hence
+$C'[x, z] = \bigoplus_{y = 0}^{4} (A[x, y, z] \oplus C[x', z'] \oplus C[x'', z''])$,
 and thus $C'[x, z] = C[x, z] \oplus C[x', z'] \oplus C[x'', z'']$, for some

### 2018-05-07 s2018-math490/thesis/thesis.tex
Updates
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

@@
 page 16 of FIPS 202 (\cite{NIST202}) for its specification; it is not discussed
-extensively in this paper.
+extensively in this paper. It takes an additional parameter, $c$, the current
+round number.
+
+    The a single round of the core round function function, $r$ is defined as
+follows:
+\begin{align*}
+    r(A, c) = \iota(\chi(\pi(\rho(\theta(A)))), c)
+\end{align*}
+And the entire core round function, \texttt{KECCAK-$f$}, is thus:
+\begin{align*}
+    \text{\texttt{KECCAK-$f$}}(A) = r(r( ...  r(A, 0) ... , 22), 23)
+\end{align*}
 
     The sponge construction can be described as follows. Fix $w$ as a power of
-two. To have a security margin of $m$ bits, the output needs to have $m$ bits;
+two in $W$. Then define a state, $s$, as above to be a binary string of length
+$25w$. Fix a security margin $m$ (usually 228, 256, 384, or 512 bits in the
+SHA-3 standard). Define $i_{m} = 25w - 2m$ be the input margin, and
+$o_{m} = 2m$ be the output margin. For a message $M$, assume it is padded
+to a multiple of $i_{m}$ bits according to the scheme in FIPS 202
+(\texttt{pad10*1}). Start with initial state $s_0 = 0^{25w}$. For each block
+$M_{i}$, XOR $M_{i}$ into the first $|M_{i}|$ bits of state $s_i$ and define:
+\begin{align*}
+    s_{i+1} = \text{\texttt{KECCAK-$f$}}(s_{i} \oplus M_{i})
+\end{align*}
+Then if $s_{k}$ is the final state, take the first $m$ bits ($s_{k}[0..m]$)
+as the output digest. Additionally if more than $2m$ bits are required, take
+the first $2m$ bits, then compute $s_{k+1} = \text{\texttt{KECCAK-$f$}}(s_{k})$
+and continue taking at most the first $2m$ bits from $s_{k+1}$, repeating as
+necessary. This is demonstrated in Figure \ref{f:s}.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=0.7\textwidth]{figs/Sponge-150.png}
+\caption{Graphical description of the sponge construction. Courtesy of the
+\href{https://keccak.team/images/Sponge-150.png}{Keccak Team}.}
+\label{f:s}
+\end{center}
+\end{figure*}
+
+To have a security margin of $m$ bits, the output needs to have $m$ bits;
 however, if the attacker has complete control of all $25w$ bits of state,
-they would be able to choose the output easily because Keccak is bijective.
-Thus, restrict the input to $25w - 2m$ bits and set the remaining $2m$ bits to
-zero. To find a preimage for a $m$-bit output, the attacker would have to find
-a $25w - m$ bit suffix such that the preimage of that suffix has $2m$ zeros.
-Similarly, for a collision attack, an attacker has to find two $25w-2m$
-messages such that their hashes have the same prefix of length $m$. Under an
-ideal round permutation, clearly these are non-trivial and hard to find in
-general.
+they would be able to choose an output with prefix $m$ easily because
+Keccak is bijective with a known, computable inverse. Thus, by restricting
+the initial state to have a suffix of $2m$ bits of zeros, the attacker must
+not only restrict the output to have the correct prefix, but also the input
+to have the correct suffix. In this way, the authors claim that Keccak/SHA-3
+has collision resistance equal to preimage resistance; finding a collision
+is of equivalent difficulty as finding a preimage, and thus Keccak/SHA-3 is
+more secure than necessary \cite{Keccak3}.
 
-    For a multiple block message the procedure is thus as follows: call the
-composition of 24 rounds to be \texttt{KECCAK-$f$}. Start with $s=0^{25w}$.
-Splitting the message into chunks of $25w-2m$ (and padding according to
-the scheme defined in FIPS 202 \cite{NIST202}), for each block $b$, extend
-$b$ to length $25w$ with zeros, then compute $s = f(s \oplus b)$. From $s$,
-take the top $m$ bits as the output of the hash function.
+Further, SHA-3 is secure from trivial multiple-block collisions due to the
+restriction of allowing $i_{m}$-bits of input state to be changed directly
+via a block. Thus, for a multiple-block collision attack to be possible, the
+remaining $25w - i_{m} = 2m$ bits must be the same between the internal state
+of a collision pair, allowing two different blocks $b$ and $b'$ to cancel
+their initial prefix differences. All of these attacks are constrained
+input/output problems and thus reduce to SAT; thus, considerable work
+is required to successfully perform an attack.
 
@@
 
-This follows from the definition of $\theta$: note that
-$\theta(0^{25w}) = 0^{25w}$. If $a \neq 0^{25w}$, then there
-exists an index set $I_{1}$ such that $\forall i \in I_{1}$, $a[i] = 1$,
-and $\forall j \not\in I_{1}$, $a[j] = 0$. Then $\theta(a) \neq 0^{25w}$,
-which follows from the construction of $\theta$. (Note that each output
-bit of $\theta$ is composed of the XOR of 11 values in a fixed pattern).
+First, note that $\theta(0^{25w}) = 0^{25w}$ due to construction of $\theta$
+having an even number of XORs. Suppose that there existed some other input, x,
+to $\theta$ such that $\theta(x) = 0^{25w}$. Note that, $x$ with odd number of
+1s (that is, $|\{ i : x_{i} = 1 \}| \equiv 1 \text { mod } 2$), $\theta(x)$
+likewise has an odd number of ones due to $\theta$ having an even number of
+XORs. Similarly, for any $x$ with has an even but non-zero number of ones,
+then there exists index $(i, j, k)$ such that $\theta(x)[i, j, k] = 1$. This is
+because, for any pair of indices $(i, j, k)$ and $(i', j', k')$ such that
+$x[i, j, k] = 1 = x[i', j', k']$, there exists an index $(a, b, c)$ such that
+$\theta(x)[a, b, c]$ is dependent on only one of $(i, j, k)$ or $(i', j', k')$,
+by construction of $\theta$. Thus the kernel of $\theta is 0^{25w}$.
 
@@
 This follows from the definition of $\theta$: note that $\theta$ is
-composed entirely of XORs and that XOR is commutative and associative.
-
+composed entirely of XORs and that XOR is commutative and associative. For any
+index $(i, j, k)$, we want to show that
+\begin{align*}
+    \theta(a \oplus b)[i, j, k] = \theta(a)[i, j, k] \oplus \theta(b)[i, j, k]
+\end{align*}
+Note that:
+\begin{align*}
+    \theta(a \oplus b)[i, j, k] & = (a[i, j, k] \oplus b[i, j, k]) \\
+        & \oplus (\bigoplus_{0 \leq y \leq 4} (a[(i - 1) \text{ mod } 5, j, k] \oplus b[(i - 1) \text{ mod } 5, j, k])) \\
+        & \oplus (\bigoplus_{0 \leq y \leq 4} (a[(i + 1) \text{ mod } 5, j, (k - 1) \text{ mod } w] \oplus b[(i + 1) \text{ mod } 5, j, (k - 1) \text{ mod } w])) \\
+        & = (a[i, j, k ] \oplus (\bigoplus_{0 \leq y \leq 4} a[(i - 1) \text{ mod } 5, j, k]) \oplus (\bigoplus_{0 \leq y \leq 4} a[(i + 1) \text{ mod } 5, j, (k - 1) \text{ mod } w])) \\
+        & \oplus (b[i, j, k ] \oplus (\bigoplus_{0 \leq y \leq 4} b[(i - 1) \text{ mod } 5, j, k]) \oplus (\bigoplus_{0 \leq y \leq 4} b[(i + 1) \text{ mod } 5, j, (k - 1) \text{ mod } w])) \\
+        & = \theta(a)[i, j, k] \oplus \theta(b)[i, j, k]
+\end{align*}
 \end{proof}
@@
 
-To construct the inverse of $\theta$, note that
-$A'[x, y, z] = A[x, y, z] \oplus D[x, z]$; hence,
-$A[x, y, z] = A'[x, y, z] \oplus D'[x, z]$ for some $D' = D$. Since $D[x, z]$
-is composed of several $C[x, z]$, where
-$C[x, z] = \bigoplus_{y = 0}^{4} A[x, y, z]$, we can similarly define $C'[x, z]$
-to be $C'[x, z] = \bigoplus_{y = 0}^{4} A'[x, y, z]$. Then, we expect the inverse
-of $\theta$ to be of a similar form. This reduces to a linear algebra problem
-over boolean variables. We know that $D'[x, z] = D[x, z]$ in order to recover
-$A[x, y, z]$. Hence we can represent $D[x, z]$ as strings of length
-$5 \times w$, where bit $i = z' + 5 \times x'$ is $1$ if and only if
-$C[x', z']$ is used in the construction of $D[x, z]$. Further, we can view each
-of the $C'[x, z]$ as being the conjunction of three $C[x', z']$ in $A$; thus
-these are strings where bit $j = z' + 5\times x'$ is 1 if and only if
-$C[x', z']$ is used to construct $C'[x, z]$.
-
-(That is, since $C'[x, z] = \bigoplus_{y = 0}^{4} A'[x, y, z]$,
-$C'[x, z] = \bigoplus_{y = 0}^{4} (A[x, y, z] \oplus D[x, z])$ and hence
-$C'[x, z] = \bigoplus_{y = 0}^{4} (A[x, y, z] \oplus C[x', z'] \oplus C[x'', z''])$,
-and thus $C'[x, z] = C[x, z] \oplus C[x', z'] \oplus C[x'', z'']$, for some
-$x, z, x', z', x'', z''$ based on the definition of $\theta$).
-
-Thus, giving each $C'[x, z]$ a constant $c_{x,z}$ for whether it is used in
-constructing $D'[x, z]$, we can form a system of linear equations and solve for
-the constants $c_{x, z}$ in each expression. Since there are $5 \times w$
-variables and $5 \times w$ equations in each equation for $D'[x, z]$, this can
-be solved easily, yielding the inverse of $\theta$.
+To construct the inverse of $\theta$, note that since $\theta$ is a linear map,
+there exists a matrix $M$ such that $\theta(x) = Mx$, when $x$ is treated as a
+column vector of bits. $M$ is described by a square matrix of size $25w$ by
+$25w$, where for a given row $i$:
+\begin{align*}
+    M_{i, j} = 1 \iff \theta(x)[i] \text{ depends on } x[j]
+\end{align*}
+Since we know that $\theta$ is invertible, $M$ must be invertible and thus
+$M^{-1}$ describes the inverse of $\theta$.
 

### 2018-05-07 s2018-math490/thesis/thesis.tex
Finished

@@
 to $\theta$ such that $\theta(x) = 0^{25w}$. Note that, $x$ with odd number of
-1s (that is, $|\{ i : x_{i} = 1 \}| \equiv 1 \text { mod } 2$), $\theta(x)$
+$1$s (that is, $|\{ i : x_{i} = 1 \}| \equiv 1 \text { mod } 2$), $\theta(x)$
 likewise has an odd number of ones due to $\theta$ having an even number of
@@
 $\theta(x)[a, b, c]$ is dependent on only one of $(i, j, k)$ or $(i', j', k')$,
-by construction of $\theta$. Thus the kernel of $\theta is 0^{25w}$.
+by construction of $\theta$. Thus the kernel of $\theta is 0^{25w}$. (This is
+also extensible to an odd-sized subset of a even-sized set of ones; that is,
+for any four indices for which $x[i] = 1$, there exists a subset of either
+one or three indices for which the other indices don't appear, up to a set of
+size 10. Since this odd subset exists, then there are an odd number of $1$s,
+and hence the output has at least one $1$ in it, hence $\theta(x) \neq 0^{25w}$).
 

### 2018-05-07 s2018-math490/jan-status/status-jan.tex
WIP: Artifacts from Spring 2018, MATH 490: Independent Study (#4)
* Status report for January.

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Added margin renderings

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Correlation analysis

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Additional analysis

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Add revised correlation images

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Added margins for steps

Signed-off-by: Alexander Scheel <alexander.m.sche

\documentclass[10pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasmathematics} % Choose template
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Status Update: End of January \\ MATH 490 - Independent Study}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Alexander Scheel}

\affil[a]{Iowa State University}

% Please give the surname of the lead author for the running footer
\leadauthor{Scheel}

% Please add here a significance statement to explain the relevance of your work

% Please include corresponding author, author contribution and author declaration information

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:

\begin{abstract}
    End of January status update for MATH 490: Independent Study at Iowa State
    University. Proposes to study SHA-3 and outlines current and future areas
    of study and completed work.
\end{abstract}

\keywords{MATH 490 $|$ Status Report $|$ SHA-3 $|$ hash\_framework }

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
% \thispagestyle{firststyle}
% \ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% Bibliography
% \bibliography{pnas-sample}

\section{Introduction}

    The purpose of this independent study is to study cryptography from a
mathematical background. The target object is the Keccak hash function
as standardized by NIST as SHA-3 and its collision resistance. A cryptographic
hash function maps arbitrary length input strings to a fixed-length digest;
finding two elements of the input domain which map to the same output should
require on the order of $2^{\frac{l}{2}}$ amount of work, where $l$ is the
output length of the hash function. SHA-3 provides an excellent study due to
its relatively simple construction as compared to previous standardized hash
functions. The goal of this independent study is to better understand
cryptanalysis techniques such as differential cryptanalysis and computer-aided
proof techniques.

    The rest of this status report is organized as follows: section
\ref{sec:overview} details the construction of SHA-3, section
\ref{sec:literature} details some existing literature in the field, section
\ref{sec:completed} details the completed work, section \ref{sec:current}
details current tasks, and section \ref{sec:future} concludes the status
report with a list of future tasks and deliverables.

\section{Overview of SHA-3}\label{sec:overview}

    SHA-3 is the latest generation of hash function standardized by NIST in
FIPS-202, published in 2015 \cite{NIST202}. The team behind this function
includes Guido Bertoni, Joan Daemen, Micha{\"e}l Peeters, Gilles Van Assche,
and Ronny Van Keer \cite{KeccakTeam}, a European team. This hash function
differs significantly from the previously accepted construction of MD4, MD5,
SHA-1, and SHA-2: while the previous hash functions relied on the
Merkel-Damg{\aa}rd construction, SHA-3 uses a sponge-function design.

    The Merkel-Damg{\aa}rd construction for cryptographic hash functions
follows what has been termed the ARX construction: arithmetic-rotation-xor.
This construction transforms an initial state to an ouput state according to
a predefined series of round functions. Each round function is typically
different from the next, and each operates on a different portion of the block.
Assume a 32-bit hash function. Typically there are between 4 and 8 state
variables, in this case, 32-bit unsigned integers. The input block is split
into 32-bit chunks and expanded according to the defined block schedule.
This block schedule can either be simple (in the case of MD4 and MD5, simply
permutations of the ordering of the chunks) or complex (in the case of SHA-1
and SHA-2, which involve xor-ing various chunks together to produce additional
chunks).

    Then, starting from the initial state, each round applies the corresponding
round function to the current state and the current block chunk, producing a
new state variable. The round function is made up of several additions,
at least one rotation and an xor; hence the term ARX construction.
This state variable is then used to replace an element of
the current state (directly in the case of MD4 and MD5, else in some fixed
pattern in SHA-1 and SHA-2). Lastly, after the last block chunk has been
processed, the initial state and the final state are added together to produce
the output state. Multiple blocks are then handled by assuming this output
state as input to the next block, and some suitable padding scheme is applied
to the input before splitting into blocks. In the case of MD4, MD5, SHA-1,
SHA-2 256 and SHA-2 512, all of the state variables are directly outputted.
Therefore, an extension attack is possible: given any input $B$, which
produced output $O$, it holds that all $C$ produce $h(O, C) = h(IV, B || C)$,
where $IV$ is the default initial state value.

    In contrast, Keccak uses a construction they call a sponge construction.
Rather than several independent state variables, Keccak uses a single
array as state, of size $25 \times 2^{l}$ for $0 \leq l \leq 6$. This is
conceptualized as a $5 \times 5 \times 2^{l}$ cuboid. Blocks are mixed in
once, via a simple xor, and the initial state is the all zero array. Then,
a round function is permuted 24 times. This round function is composed of
five functions: $\theta, \rho, \pi, \chi, \iota$. Each function is a bijection
on the set of all state arrays. Thus, after 24 rounds are applied, a portion
of the resulting state is specified as output. If multiple blocks are given,
the entire state is used as input, with the next block xored in, before
applying the round function.

    SHA-3 standardized $l=6$, producing a state array of 1600 bits. Further,
since SHA-2 standardized several different security margins, (224, 256,  384,
and 512), SHA-3 did as well. Take the security margin to be $b$ bits.
Then the input block size is $1600 - 2\times b$ and $b$ bits of the resulting
state is taken as output. While this prevents length extension attacks on
every block, this opens the possibility for a specialized extension attack
with lower possibility. Suppose $S_0$ is the initial state after mixing
in a block. If $\text{keccak-}p(S_0)$ has the same lower $2 \times b$ bits
as $S_0$, then $h(S_0 \oplus C) = h(S_0 || C)$, for all messages $C$.

\section{Review of Literature}\label{sec:literature}

    The canonical reference for SHA-3 is FIPS-202 \cite{NIST202}. However,
there are several useful supplementary materials to use in conjunction with
FIPS-202. The first is the core Keccak Reference: published by the Keccak team,
this contains extended information underlying their choice of functions,
properties of these functions, documentation on inverse functions, and external
references \cite{Keccak3}. Supplementing this is the Keccak Code Package,
which contains many implementations of Keccak and test cases for several bit
widths \cite{KeccakCodePackage}. Lastly, the PhD thesis of J. Daemen seems
to have heavily influenced the construction of Keccak, and provides extensive
reference material \cite{DaemenThesis}.

    Of the attacks I've looked at, the work of E. Homsirikamol, et. al,
is most relevant to my existing work \cite{cryptoeprint:2012:421}. They
directly applied SAT solvers to find collisions in Keccak. However, they used
only the most simple model and were only able to find a two-round collision
within 30 hours. Thus it serves more as a benchmark than as useful
cryptanalysis. The work of P. Morawiecki and M. Srebrny also looks at applying
SAT solvers to Keccak, but explicitly focused on preimage attacks
\cite{cryptoeprint:2010:285}. In this instance, it appears the authors only
considered a direct preimage attack with little other work towards splitting
the work between multiple servers or doing advanced exploration of possible
paths. However, they did appear to build a toolkit, CryptLogVer, for studying
this problem.

\section{Completed Work} \label{sec:completed}

    Completed is an initial survey of the literature on the subject of the
construction of SHA-3 and attacks against it. Further, an implementation
of SHA-3 has been added to my hash\_framework package \cite{hf}. This has been
tested against the vectors in the Keccak Code Project \cite{KeccakCodePackage},
and is shown correct.

    Several initial attack vectors have been considered as well. Towards
studying the specialized extension attack, an initial pass trying to bound the
error in the remaining bits. However, up to an input difference of 5, no
alternate paths have been found producing a net-zero difference in the tail.
This has yet to be studied on reduced width versions of Keccak.

Towards studying collision attacks, a similar error-bounding property has been
studied on front prefixes. Given that there are $k$ bits of difference in the
input, and given an input margin of $m$ (where the remaining 1600 - m bits
are all zero), what is the largest $M$ such that, after applying
only the first round, the difference between the two blocks is zero in the
first $M$ bits. This has been exhaustively studied on the $l=3$ and $l=4$
bit-width versions of Keccak, and partially on the full width ($l=6$). The
results are startling: for small, even $k$, $M$ quickly grows large, whereas
for odd, small $k$, $M$ remains fixed at a relatively small number. Thus
there is the potential to exploit this. However, as $k$ grows large, the
odd and even values of $k$ converge in structure. This can seen in the links
attached to this email.

\section{Current Work} \label{sec:current}

    Currently, I am rewriting my job distribution mechanism to work more
efficiently on larger deployments such as an HPC cluster or AWS cloud
instances. This will add several desirable features: better fault tolerance,
handling job distribution automatically and prioritizing between several
tasks, and devoting a set number of cores per task. Each task will then have
many jobs and statistics about them can be collected and centralized (such as
number of variables, time to solve, etc.). This would provide the basis for
more closely analyzing the performance of the entire system.

\section{Future Work} \label{sec:future}

    After this rewrite, I'll look at exploring a few areas: analyzing the
impact of each of the sub-round functions ($\theta, \rho, \pi, \chi, \iota$)
on various differential paths, continuing the above work on analyzing
input and output margins on $k$-bit differences, and looking at total, given
$k$ bits of input difference, what possible combinations of output differences
are possible.

    Further, I'll implement HPC-specific workloads. The most obvious starting
point is to analyzing small versions of Keccak ($l=4$) for two-variable
correlations across each of the several rounds. Further, studying the
per-round correlation of initial input differential on resulting intermediate
round differentials would be an interesting result as well, for small $l$.

    Once these properties have been more closely explored, there is potential
for proving theorems based on the results. However, I think it is important to
get some initial results from these various areas to see how they all fit
together and what avenues are most likely to help attack Keccak.

\section{Bibliography} \label{sec:bibliography}

\bibliography{thesis}

\end{document}
